### 2018-11-28 papers/chromatography (for ethics course)/paper.tex
added paper for ethics course homework

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Comanda, analiza și prelucrarea datelor a unui cromatograf de proces
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Toma Becea}
\IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
\textit{Technical University of Cluj-Napoca}\\
Cluj-Napoca, Romania \\
tomabecea@pm.me}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Honoriu Vălean}
\IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
\textit{Technical University of Cluj-Napoca}\\
Cluj-Napoca, Romania \\
Honoriu.Valean@aut.utcluj.ro}
}

\maketitle

\begin{abstract}
Realizarea unei aplicații de achiziții a datelor provenite dintr-un cromatograf de proces. 
Prelucrarea datelor se face printr-o funcție de liniarizare specifică fiecărui aparat, printr-o funcție de calibrare iar apoi,
rezultatele sunt trecute într-o bază de date special proiectată pentru aplicații de urmărire geologică. 
Totodată, în funcție de performanțele constructive ale aparatului, se face comanda lui prin porturi digitale.
\end{abstract}

\begin{IEEEkeywords}
cromatografie, , style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
%\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2018-11-28 papers/chromatography (for ethics course)/paper.tex
Abstract and introduction written on chroma paper

@@
 
-\title{Comanda, analiza și prelucrarea datelor a unui cromatograf de proces
+\title{Command, analyse and data mining of a chromatograph during drilling operations
 }
@@
 \begin{abstract}
-Realizarea unei aplicații de achiziții a datelor provenite dintr-un cromatograf de proces. 
-Prelucrarea datelor se face printr-o funcție de liniarizare specifică fiecărui aparat, printr-o funcție de calibrare iar apoi,
-rezultatele sunt trecute într-o bază de date special proiectată pentru aplicații de urmărire geologică. 
-Totodată, în funcție de performanțele constructive ale aparatului, se face comanda lui prin porturi digitale.
+    Chromatograph analyses need a great attention to the way results are generated and how they are calibrated.
+    At the same time the construction and maintenance need to be simple and quick such that any operation which
+    needs to be made on a chromatograph doesn't stop the entire drilling crew.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-cromatografie, , style, styling, insert
+chromatography, geology, mud-logging
 \end{IEEEkeywords}
@@
 \section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+    During drilling operations, while seeking gas or oil, certain gas analyses must be taken.
+    Those analyses are performed by a special device called chromatograph. 
+    Present paper presents a novel way of generating the mathematical model of a chromatograph during its construction
+    and then using this model to ensure the results accuracy.
+\section{chromatograph}
 
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
+\subsection{Construction}
 
@@
 
-\section{Prepare Your Paper Before Styling}
+\subsection{Maintenance}
+
+\section{Mathematical model}
 Before you begin to format your paper, first write and save the content as a 

### 2018-11-30 papers/chromatography (for ethics course)/paper.tex
Chromatography paper: introduction and first part with hardware added

@@
 
-\title{Command, analyse and data mining of a chromatograph during drilling operations
-}
+\title{Analyse and data mining of a chromatograph during drilling operations}
 
@@
 
+\subsection{Chromatography concepts}
+
+    Gas chromatography means to separate components such that they can be analyzed independently. To achive this a
+    separation column is used. A probe of gas, taken from drilling installation, is pushed through the column which
+    means that all components enter into it at the same time. The column is filled with carefully choosen particles,
+    whose type and size are competitive commercial advantages (and because this is not the object of the current paper,
+    their details will not be disclosed). The probe is pushed through the column using a bearing gas, usually hydrogen.
+    The probe has to make its way through the column's particles and because of the friction between them different
+    components will have different speeds with which they advance. Thus, the components are reaching the end of the 
+    column at different times. The time which is needed for a component to traverse the column is called retention time.
+    Figure 1 contains a representation of a column and how two components, A and B, enter into the column at the same
+    time but they exit at different times.
+    During the usual drilling operations geologist engineers are interested in certain components so that they can analyze
+    and predict the behavior of the gas or petrol reservoir. Those components are: methane (C1), ethane (C2), propane (C3),
+    n-butane (nC4), isobutane (iC4), neopentane (nC5), isopentane (iC5). The sum of all those components yields a number
+    called theoretical total gas. The probe is also analysed as a whole to find out the total gas. The ratio of those two
+    numbers is of paramount importance.
+    Among the details which can be find out using the analyses of those components are the quality of the gas or petrol
+    which is in the reservoir, the proximity of the reservoir or certain predictors important for drilling engineers, for
+    example, the danger of the entire reservoir to erupt.
+
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.5\textwidth]{column.png}
+        \caption{Gas separation column}
+    \end{figure}
+
+\subsection{Flame Ionisation Detector}
+
+    The components are slowly (or fast, depending on the performance of the chromatograph) pouring out of the column
+    into a detector called Flame Ionisation Detector or FID. Its role is to transform over the analyses time the quantity
+    of each component into voltage. This voltage is read by ADC boards and transmitted to a computer for further calculations.
+    Figure 2 shows the schematics of such a FID. Item 1 represents the separation column, item 2 represents suplimentary
+    gas supply, 3 represents hydrogen supply, 4 represents the lighter, 5 represents the collector electrode, 6 is the
+    polarizing electrode, 7 is the burner and 8 is the air supply.
+
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.5\textwidth]{fid.png}
+        \caption{Flame Ionisation Detector}
+    \end{figure}
+
 \subsection{Construction}
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+Following description will be brief one and it focuses only on what is important for this paper. The construction of a 
+chromatograph implies certain components to be glued together.
 

### 2018-12-02 papers/chromatography (for ethics course)/paper.tex
more writing in chroma paper

@@
     time but they exit at different times.
+
     During the usual drilling operations geologist engineers are interested in certain components so that they can analyze
@@
     numbers is of paramount importance.
+
     Among the details which can be find out using the analyses of those components are the quality of the gas or petrol
@@
 
-Following description will be brief one and it focuses only on what is important for this paper. The construction of a 
-chromatograph implies certain components to be glued together.
+    Following description will be a brief one and it focuses only on what is important for this paper. The construction of a 
+    chromatograph implies certain components to be glued together. We have already seen the separation columns. Those are 
+    two or three wired through special designed pneumatics components. Then there might be two or three FIDs where the
+    output of the columns is going. The entire process also needs a source of compressed air and a source of hydrogen. For
+    easier manipulation the front panel may contain analog pressure indicators and few knobs for pressure and hydrogen
+    regulation.
+
+    The FIDs have two electrodes which are wired to a ADC converter through a special interface that has the output between
+    0 Volts and 10 Volts. The ADC converts this data into a digital format and sends it to a computer over a USB connection.
+
+    Following the mechanical construction of such a chromatograph the engineer who built it needs to calibrate it. For this
+    a calibrating gas is used and injected via a face mounted pneumatic orifice. The chromatograph will analyze the gase received
+    from this input, rather than the gas received from the usual output, i.e. site drilling rig.
+
+    Therefore, an easy approach should be considered on delivering a chromatograph with ready-to-use software.
 
@@
 
+    A drilling site is filled with dust, the temperatures might drop below freezing for extended periods of time, the mud is
+    a constant and the mechanical load during transport from one site to another is a burden. Then there always might extended
+    periods with no activity.
+
+    Therefore, the chromatograph must be easy to setup again, easy to start and easy to calibrate again because in time its 
+    FID characteristic tend to skew.
+
+    There are also periodic calibration of the chromatograph for better results and for verifying that certain results, which
+    might seem odd at first, are actually real.
+
+    Both the construction side of a chromatograph and its maintenance makes the designer of this device to consciously design
+    it for those situations.
+
 \section{Mathematical model}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-%\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.5\textwidth]{fidgraph.png}
+        \caption{FID Graph}
     \end{figure}
 
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
 \begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\bibitem{b1} G. Becea, ``First steps in chromatography'' Internal document of S.C. Daflog S.R.L., 2007.
+\bibitem{b2} T. Beu ``Numerical analyses in Turbo Pascal'', 1992.
+\bibitem{b3} I. Ciucanu, ``Gas chromatography with capilar columns'' 1990.
 \end{thebibliography}
 \vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
 

### 2018-12-02 papers/chromatography (for ethics course)/paper.tex
Finished chroma paper for ethics course.

@@
     column at different times. The time which is needed for a component to traverse the column is called retention time.
-    Figure 1 contains a representation of a column and how two components, A and B, enter into the column at the same
+    Figure \ref{fig:column} contains a representation of a column and how two components, A and B, enter into the column at the same
     time but they exit at different times.
@@
         \caption{Gas separation column}
+        \label{fig:column}
     \end{figure}
@@
     of each component into voltage. This voltage is read by ADC boards and transmitted to a computer for further calculations.
-    Figure 2 shows the schematics of such a FID. Item 1 represents the separation column, item 2 represents suplimentary
+    Figure \ref{fig:fid} shows the schematics of such a FID. Item 1 represents the separation column, item 2 represents suplimentary
     gas supply, 3 represents hydrogen supply, 4 represents the lighter, 5 represents the collector electrode, 6 is the
@@
         \caption{Flame Ionisation Detector}
+        \label{fig:fid}
     \end{figure}
@@
 
-\section{Mathematical model}
-
     \begin{figure}
@@
         \caption{FID Graph}
+        \label{fig:fidgraph}
     \end{figure}
+
+\section{Mathematical model}
+
+    The mathematical model is so simple, one could wonder if it works at all: it's based on the least squares equation.
+    From figure \ref{fig:fidgraph} it can be seen that during one minute of analyses the first two FIDs have some components present.
+    Using different rules for establishing a baseline the area of each componente (each "hill") is computed. This area is 
+    proportional with the amount of the component from the last probe analysed. The engineer who builds the chromatograph and then
+    the engineer who maintain it will introduce, in consecutive analyses, different gases with a precisely known composition, using
+    different concentrations. Thus, a series of analyses will be made, knowing for each one the calibrating gas composition and the
+    results. A look tabel is born from this operation.
+
+    During normal operations the look up tabel will be used to analyze gases from drilling site. To be able to give results for any
+    given concentration, for any given component, an interpolation is needed between two known points, using first two points or
+    using last two points. This kind of interpolation is very similar to an average between two numbers. Equation \ref{eq:average}
+    represents such an average: \(K_n\) and \(K_{n+1}\) are numerical integrations results and  \(C_n\) and \(C_{n+1}\) are
+    corresponding component's concentrations from the lookup table. \(R\) will be the result of the current analyze.
+
+    \begin{equation}
+        R = \frac{C_n + C_{n+1}}{K_n + K_{n+1}}, K_{n} < R < K_{n+1}
+        \label{eq:average}
+    \end{equation}
+
+    Another method for obtaining results using the data from the lookup table is to use the least squares method. Equation \ref{eq:leastSquares}
+    shows the gist of this method. \(x_{1}\) through \(x_{n}\) are the known concentrations from calibration operation. \(y\) terms are
+    the numerical integration results for a component, for each concentration. \(b\) terms are the coefficients which are found by 
+    solving the equation \ref{eq:leastSquares}. \(b\) will form a polynom which will be used during normal analyze mode to computed
+    the real concentration of a component from a probe taken from the drilling site.
+
+    \begin{equation}
+        \begin{bmatrix}
+             x_0 \\ \vdots \\ x_{n} 
+         \end{bmatrix}
+             =
+         \begin{pmatrix}
+            y_{00} & \hdots & y_{0(m-1)} & y_{0m}\\
+            \vdots & & \vdots & \vdots \\
+            y_{n0} & \hdots & y_{n(m-1)} & y_{nm}
+            \end{pmatrix}
+            \times
+            \begin{pmatrix}
+            b_0 \\ \vdots \\ b_m
+        \end{pmatrix}
+        \label{eq:leastSquares}
+    \end{equation}
 

### 2018-12-02 papers/chromatography (for ethics course)/paper.tex
Small corrections on chroma paper.

@@
     A drilling site is filled with dust, the temperatures might drop below freezing for extended periods of time, the mud is
-    a constant and the mechanical load during transport from one site to another is a burden. Then there always might extended
+    a constant and the mechanical load during transport from one site to another is a burden. Then there always might be extended
     periods with no activity.
@@
     From figure \ref{fig:fidgraph} it can be seen that during one minute of analyses the first two FIDs have some components present.
-    Using different rules for establishing a baseline the area of each componente (each "hill") is computed. This area is 
+    Using different rules for establishing a baseline the area of each component (each "hill") is computed. This area is 
     proportional with the amount of the component from the last probe analysed. The engineer who builds the chromatograph and then
@@
     \begin{equation}
-        R = \frac{C_n + C_{n+1}}{K_n + K_{n+1}}, K_{n} < R < K_{n+1}
+        R = \frac{C_n + C_{n+1}}{K_n + K_{n+1}}, K_{n} \leq R \leq K_{n+1}
         \label{eq:average}
@@
     the numerical integration results for a component, for each concentration. \(b\) terms are the coefficients which are found by 
-    solving the equation \ref{eq:leastSquares}. \(b\) will form a polynom which will be used during normal analyze mode to computed
+    solving the equation \ref{eq:leastSquares}. \(b\) terms will form a polynom which will be used during normal analyze mode to computed
     the real concentration of a component from a probe taken from the drilling site.

### 2019-01-10 papers/paper reviews (for ethics course)/rec_1.tex
Added blind paper reviews.

\documentclass{article}
% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{geometry}
    \geometry{
        a4paper,
        total={170mm,257mm},
        left=20mm,
        top=10mm,
        }

\begin{document}

\title{Peer single blind review of "Web-Controlled Robot Based on Raspberry Pi" paper}

\author{}

\maketitle

\section{Plagiarism assesment}

The tool which was used: www.copyleaks.com. Results:

\begin{itemize}

    \item 23\% percent of content is similar to other sources
    \item 519 words copied
    \item less than 1 \% minor changes
    \item less than 1 \% of words with related meanings

\end{itemize}

\section{Abstract and introduction}

\begin{itemize}

    \item Provide a short description of the proposed improvement (solution) in the abstract so that subsequent sections are easier to understand within a context
    \item Expand the solution in the description instead of the lenghty Raspberry PI history

\end{itemize}

\section{Research specifics}

\begin{itemize}

    \item Add reference for "Getting Started" documentation of PiCamera Python library
    \item On Section II "Related work" describe the actual work done to connect the PiCamera to a Raspberry and how its Python library was used to interface with it
    \item Describe the novelty which the proposed solution brings besides interconneting various Raspberry compatible kits with a Raspberry computer.
    \item Provide actual data of how the research went during the control of the robot (communication round trip time, lag, time needed for IP registration, connection loss mitigation, bandwith benchmarks, resolution fallback on low bandwith, etc.)

\end{itemize}

\section{Results specifics}

\begin{itemize}

    \item Provide data interpretation of the tests run on the robot and its connectivity platform
    \item Provide and document the novelty of the proposed solution, together with its significance

\end{itemize}

\section{Recommended decision}

Fundamental changes are required for this paper, therefore the paper is rejected.

\end{document}

### 2019-01-10 papers/paper reviews (for ethics course)/rec_2.tex
Added blind paper reviews.

--- 2019-01-10 papers/paper reviews (for ethics course)/rec_1.tex
@@
 
-\title{Peer single blind review of "Web-Controlled Robot Based on Raspberry Pi" paper}
+\title{Peer single blind review of "Control loops in automotive systems" paper}
 
@@
 
-The tool which was used: www.copyleaks.com. Results:
-
-\begin{itemize}
-
-    \item 23\% percent of content is similar to other sources
-    \item 519 words copied
-    \item less than 1 \% minor changes
-    \item less than 1 \% of words with related meanings
-
-\end{itemize}
-
-\section{Abstract and introduction}
-
-\begin{itemize}
-
-    \item Provide a short description of the proposed improvement (solution) in the abstract so that subsequent sections are easier to understand within a context
-    \item Expand the solution in the description instead of the lenghty Raspberry PI history
-
-\end{itemize}
+The tool which was used: www.quetext.com Results: no plagiarism detected.
 
@@
 
-    \item Add reference for "Getting Started" documentation of PiCamera Python library
-    \item On Section II "Related work" describe the actual work done to connect the PiCamera to a Raspberry and how its Python library was used to interface with it
-    \item Describe the novelty which the proposed solution brings besides interconneting various Raspberry compatible kits with a Raspberry computer.
-    \item Provide actual data of how the research went during the control of the robot (communication round trip time, lag, time needed for IP registration, connection loss mitigation, bandwith benchmarks, resolution fallback on low bandwith, etc.)
+    \item State clearly the novelty of the control loops used
+    \item State clearly the novelty of some of the tools and techniques used for verifying the control loops: Unity, UDP, unit tests, systems testing.
 
@@
 
-    \item Provide data interpretation of the tests run on the robot and its connectivity platform
-    \item Provide and document the novelty of the proposed solution, together with its significance
+    \item Provide data interpretation of the control loop results, using their time diagrams
+    \item Discuss the performance of each control loop
+    \item Add and briefly discuss future improvements which can be made to this system
 
@@
 
-Fundamental changes are required for this paper, therefore the paper is rejected.
+Minor revisions are required for this paper.
 

### 2019-10-22 papers/traffic simulation/paper.tex
Start to edit the Traffic Simulation paper.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{A microservices based approach for traffic simulation}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
Simulating traffic in a city is tedious task.
\end{abstract}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
% \centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2019-10-23 papers/traffic simulation/paper.tex
Outlined the city traffic format.

@@
 
-\title{A microservices based approach for traffic simulation}
+\title{A microservices based approach for city traffic simulation}
 
@@
 \begin{abstract}
-Simulating traffic in a city is tedious task.
+Simulating traffic in a city is a difficult task.
+Current paper is proposing a novel way using state of the art in distributed  microservices: orchestrating microservices.
+The desing is centered around two types of actors: a city simulation which keeps track of streets and cars densities.
+The other actor is the care which travels across the city.
 \end{abstract}
@@
 \section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
 
-\section{Ease of Use}
+\section{Microservices}
 
-\subsection{Maintaining the Integrity of the Specifications}
+\subsection{Inception}
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+\subsection{Current state}
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+\section{Orchestration}
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
+\section{Design}
 
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
+\subsection{Tools and programming languages}
 \begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+\item Go programming language
+\item Docker
+\item Docker compose
+\item OpenSourceMap
 \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+\subsection{Choices and rationales}
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+\subsection{Alternatives}
 
-\subsection{\LaTeX-Specific Advice}
+\subsection{Communication design}
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+\subsection{Orchestration design}
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
+\section{Results}
 
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
+\subsection{Visual}
 
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
+\subsection{Performance}
 
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-% \centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+\subsection{}
 
@@
 \vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
 

### 2019-10-23 papers/traffic simulation/paper.tex
Outlines and notes completed.

@@
 \documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
@@
 Simulating traffic in a city is a difficult task.
-Current paper is proposing a novel way using state of the art in distributed  microservices: orchestrating microservices.
-The desing is centered around two types of actors: a city simulation which keeps track of streets and cars densities.
-The other actor is the care which travels across the city.
+Current paper is proposing a novel way using state of the art in distributed systems: orchestrating microservices.
+The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation.
+The other actor type (multiple instances) is the car which travels across the city.
 \end{abstract}
@@
 \section{Introduction}
+Simulation of traffic in a given city has many benefits and applications.
+
+(Terms definition)
 
@@
 
-\subsection{Inception}
+\subsection{Current state}
+
+\subsection{Implementation}
+
+\section{Actors}
+
+\subsection{City simulation actor}
+(Reason)
+
+(Types of data held: lines and densities)
+
+(Future enhancements: integration into car actor for distributed consensus and data exchange)
+
+\subsection{Car actor}
+(Routing logic)
+
+\section{Orchestration}
 
@@
 
-\section{Orchestration}
+\subsection{Implementation}
+
+\subsection{Kubernetes}
 
@@
 \item OpenSourceMap
+\item Graphhopper
+\item Web page choices (web page will become an actor)
 \end{itemize}
@@
 \subsection{Alternatives}
+(Actor model and consacrated frameworks)
 
-\subsection{Communication design}
+\subsection{Networking design}
 
@@
 
-\subsection{}
+\section{Enhancements}
+(Distributed database for multiple city simulators)
+
+(More granular logic on actor behavior)
 

### 2019-10-27 papers/traffic simulation/paper.tex
Created separate wiki.bib and advanced on the paper writing.

@@
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
+\usepackage{url,apacite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \usepackage{xcolor}
+\bibliographystyle{apacite}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 Simulating traffic in a city is a difficult task.
-Current paper is proposing a novel way using state of the art in distributed systems: orchestrating microservices.
+Current paper is proposing a novel way using state of the art in distributed systems: microservices orchestration.
 The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation.
@@
 \section{Introduction}
-Simulation of traffic in a given city has many benefits and applications.
+A solution which simulates car, pedestrian, etc.
+traffic in a given city may reap many benefits.
+It can help understand patterns of traffic and its flow.
+It can help unerstand the particularities and pecularities of a city's streets arrangements.
+It can help identify bottlenecks.
+It can help find solutions to rush problems and explore them.
+But for those areas to be tackled, appropiate methods of simulating traffic must be found.
+We are proposing a novel way of simulating traffic based on the microservices orchestration concept.
 
-(Terms definition)
+We define a \textbf{(city) actor} as being a independent entity which chose to move between two geographical points within a city.
+It can be a car, a pedestrian or a bike.
+We also define the \textbf{city simulator} as being a single entity (subject to distributed and load balacing services) which keeps data about the city (e.g.
+streets with city actors on them).
+
+\section{Existing solutions and ideas}
+(Summary of other papers with traffic simulation)
 
@@
 \subsection{Current state}
+Microservices are not a new concept.
+The idea behind them has existed since Linux kernel has started to be enriched with a concept called namespaces \cite{wiki:linuxns}.
+This allows one set of processes to see one set of resources while another set of processes see another set of resources, where resources might be, but not limited to, process IDs, file names and network resources.
+Those linux kernel abilities form the base of containers.
 
-\subsection{Implementation}
+Thus, a container is a small set of processes which run in isolation.
+They allow packaging a linux distro, a set of libraries and skd and on top of those custom code.
+This forms an image, which is essentially a tar gzipped file.
+Once the build process of an image is finished, it can be spinned up in one or more running containers.
+The custom code written and embedded in the image is running in parallel in each container.
 
-\section{Actors}
+To go to solution for building, manipulating and running images is Docker \cite{docker}.
+It allows easy software installation, it works cross platform and it offers a smooth experience most of the time.
 
-\subsection{City simulation actor}
-(Reason)
+\section{Implementation}
 
-(Types of data held: lines and densities)
+The entities which are participating into a traffic simulation are called actors.
+They are two: the city actor and the city simulator.
+The supporting containers are not themselves part of traffic simulation but they do have an important and supporting role.
+
+\subsection{City simulator}
+The most common and easy solution to share data across all the city actors is to have a centralized store to keep it.
+City simulator acts as a centralized store for all other city actors.
+In the current implementation the city simulator keeps a set of data which can be described as a list of pairs, each pair having a line and a real number, called density.
+
+The density is defined as the number of actors (cars) which are at a given moment present on a given segment of street.
+If the city simulator has no entry of a street segment then it will consider the density as being 0, i.e.
+there is no actor on that street.
+The density is modeled as a unsigned integer.
 
@@
 \subsection{Car actor}
+The city actor represents a moving actor within a city.
+It can be a car which moves across the city, it can be a bike or it can be a pedestrian.
+Its naming suggests that it can be any entity or living being which moves within a city and interacts with the other entities or affects the other entities in some manner.
+A pedestrian would directly interact with other pedestrians but not with cars.
+A pedestrian would indirectly affect other cars by willing to
 (Routing logic)
+
+\subsection{Interactions}
 
@@
 
+\subsection{Docker compose}
+
 \subsection{Kubernetes}
 
-\section{Design}
+\section{Design choices}
 
@@
 
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
+(More types of "junctions": classic junction for cars, pedestrian crossings, subways access stairs and elevators, bus stations, etc.)
+
+(Contained actors: pedestrians in a bus, pedestrian in a subway)
+
+\bibliography{wiki}
 \vspace{12pt}

### 2019-11-03 papers/traffic simulation/paper.tex
Advance in writing the traffic simulation paper.

@@
 \usepackage{xcolor}
+\graphicspath{ {.} }
 \bibliographystyle{apacite}
@@
 T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\usepackage{listings}
+\usepackage{xcolor}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
+
 \begin{document}
@@
 \begin{abstract}
+
 Simulating traffic in a city is a difficult task.
@@
 The other actor type (multiple instances) is the car which travels across the city.
+
 \end{abstract}
@@
 \section{Introduction}
+
 A solution which simulates car, pedestrian, etc.
@@
 
-\section{Existing solutions and ideas}
+\section{Existing solutions}
+
 (Summary of other papers with traffic simulation)
@@
 \subsection{Current state}
+
 Microservices are not a new concept.
@@
 \subsection{City simulator}
+
 The most common and easy solution to share data across all the city actors is to have a centralized store to keep it.
@@
 there is no actor on that street.
-The density is modeled as a unsigned integer.
+The density is modeled as an unsigned integer.
 
-(Future enhancements: integration into car actor for distributed consensus and data exchange)
+Figure \ref{fig:citysimrelations} depicts the city simulator in relation with the other entities.
+A notable exception is the web page.
+Its purpose is not to participate in the same information exchange the other have but to offer a visual interpretation of the way the city and the other actors are interacting with one another.
+
+\begin{figure}
+    \includegraphics[width=8cm]{CitySimulator.png}
+    \centering
+    \caption{The relations of city simulator with other entities}
+    \label{fig:citysimrelations}
+\end{figure}
+
+As various actors send data about their location, the city simulator needs to keep location and density data in its store.
+It does not know of Open Street Map maps and its corresponding map data sets but instead it requires any actor to send its current set of coordinates which it is crossing or which it left.
+If any other actor send the same set of coordinates, or a subset of it then the first set of coordinates will have its density incremented.
+Listing \ref{lst:goactorstruct} shows the Go struct which is used by an actor to send its report to the city simulator and is used by the city simulator to decode a message received from an actor.
+Apart from this line there are two more details to complete the report picture.
+Listing \ref{lst:goactorreports} contains the type of communications between actors and the citysimulator.
+
+\begin{lstlisting}[caption=Go struct for actor's report, label=lst:goactorstruct]
+// Report is the base type for reporting
+// status and vectors to a city entity
+type Report struct {
+	CurrentLine  [][]float64
+	ReportDetail int
+}
+\end{lstlisting}
+
+As part of this proposal the city simulator is made to be a standalone entity (container) which communicates with actors.
+This might not be the case for other types of communications, as part of other architectures.
+One example can be the integration or the unifying of the city simulator with the city actor, both becoming one entity.
+In this case the communication between entities is subject to an entire panoply of choices.
 
 \subsection{Car actor}
+
 The city actor represents a moving actor within a city.
@@
 A pedestrian would directly interact with other pedestrians but not with cars.
-A pedestrian would indirectly affect other cars by willing to
-(Routing logic)
+A pedestrian would indirectly affect other cars by willing to walk over a crossing.
+
+The implementation proposed here is aiming to model a single type of actor: a car which moves between two points across a city.
+Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \cite{openstreetmap} and GraphHopper \cite{graphhopper}.
+Open Street Map is an open source licensed map of the entire world.
+Graphhoper is an open source service which offers directions APIs and route planning.
+It can use Open Street Map as an underlying map provider.
+They provider free tiers and paid subscription for accessing an API and get routes.
+However, a city actor is not using any public api but a special crafted container which contains GraphHopper and an open street map embedded into it.
+Thus, this container is running in parallel with the other city actors and provides them with the routing API they need.
+
+A first way to introduce randomness into the entire simulation of a city is to choose a set of two coordinates inside the given city, a set for each city actor.
+Then the actor proceeds to ask GraphHopper service for a route between those two points.
+Because each city actor 'lives' only while its moving across its route and 'dies' as soon as it reached the finish point, the entire city simulation which takes place is made of independent, random and always new actors.
 
 \subsection{Interactions}
+
+An actor interacts, for the time being, with the city simulator by using a Go struct and a Go enum. They can be seen in listings \ref{lst:goactorstruct} and \ref{lst:goactorreports}.
+
+\begin{lstlisting}[caption=Go struct for actor's report, label=lst:goactorreports]
+const (   
+    // ReportOnTheLine is the report sent by one agent to
+    // notify the city that he is currently advancing
+    // through one line.
+    ReportOnTheLine = iota
+
+    // ReportOffFromLine is the report sent by one agent to
+    // notify the city that he has finished advancing through
+    // one line and has departed from it.
+    ReportOffFromLine = iota
+)
+
+const (
+    // SendReport is a message passed from an actor to the city
+    // indicating its status (e.g. location).
+    SendReport = iota
+
+    // AskForLine is a message passed from an actor to the city.
+    // A response is awaited.
+    AskForLine = iota
+
+    // RespondWithLine is a message passed from the city to
+    // an actor and it contains line data.
+    RespondWithLine = iota
+)
+\end{lstlisting}
 
@@
 \item Docker compose
-\item OpenSourceMap
-\item Graphhopper
 \item Web page choices (web page will become an actor)

### 2019-11-03 papers/traffic simulation/paper.tex
A little more writings.

@@
 A notable exception is the web page.
-Its purpose is not to participate in the same information exchange the other have but to offer a visual interpretation of the way the city and the other actors are interacting with one another.
+Its purpose is not to participate in the same information exchange the other entities have but to offer a visual interpretation of the way the city and the other actors are interacting with one another.
 
@@
 A pedestrian would directly interact with other pedestrians but not with cars.
-A pedestrian would indirectly affect other cars by willing to walk over a crossing.
+A pedestrian would indirectly affect other cars by willing to walk over a street crossing.
 
@@
 
-An actor interacts, for the time being, with the city simulator by using a Go struct and a Go enum. They can be seen in listings \ref{lst:goactorstruct} and \ref{lst:goactorreports}.
+An actor interacts, for the time being, only with the city simulator by using a Go struct and a Go enum.
+They can be seen in listings \ref{lst:goactorstruct} and \ref{lst:goactorreports}.
+The message sent from one side to another has a general structure called \textit{Envelope}.
+It is meant to offer a top level message which can be serialized (or encoded, the way to do this in Golgang, if not gRPC, is the gob pacakge it offers out of the box) and passed around, enabling any party involved to understand what this message is about and how to decode it.
 
-\begin{lstlisting}[caption=Go struct for actor's report, label=lst:goactorreports]
+Listing \ref{lst:goenvelope} show the top level message.
+It contains a message type which instructs the reader what kind of message it has to deal with and the possible types are the second part of listing \ref{lst:goactorstruct}: \textit{SendReport}, \textit{AskForLine}, \textit{RespondWithLine}.
+Let's take them one by one.
+\textit{SendReport} represents a message sent from an actor to the city simulator and the \textit{Payload} contains the report seen in listing \ref{lst:goactorreports}.
+
+\textit{AskForLine} is a message sent from an actor to the city simulator in which the actor asks about the density of any line.
+This way any actor can take conscious decisions on which route to go on when based on what lies ahead in terms of street densities.
+When a first route is known, between two desired points, the actor can ask about each line it has on that route.
+The city simulator will respond with the known density of it.
+If the actor desires, it can try to find another route by asking GraphHopper service to compute a new one but with an additional rule: avoid a certain point (street, junction, etc.).
+\textit{RespondWithLine} is the type of message which the city simulator sends back to an actor after it received an \textit{AskForLine} message.
+
+\begin{lstlisting}[caption=Go enumerations for messaging, label=lst:goactorreports]
 const (   
@@
 
+\begin{lstlisting}[caption=Go top level struct (envelope), label=lst:goenvelope]
+// Envelope is the container for different messages sent back
+// and forth between an actor and a city
+type Envelope struct {
+    MessageType int
+    Payload     interface{}
+}
+\end{lstlisting}
+
 \section{Orchestration}
+
+The term orchestration means the handling of containers and microservices in order to bring coherence into their interaction.
+They have to run on multiple nodes (computers or virtual machines) and at the same time to communicate with one another.
+They need to be updated in place and without service disruption.
+Whenever one of them crash they have to restart as quickly as possible.
+They have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc.
+They have to
 

### 2019-11-10 papers/traffic simulation/paper.tex
Wrote another bunch of things into the traffic simulation paper.

@@
 \usepackage{xcolor}
+\usepackage{natbib}
 \graphicspath{ {.} }
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\author{\IEEEauthorblockN{1\textsuperscript{st} Toma Becea}
+    \IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
+    \textit{Technical University of Cluj-Napoca}\\
+            Cluj-Napoca, Romania \\
+            tomabecea@pm.me}
     \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+    \IEEEauthorblockN{2\textsuperscript{nd} Honoriu V{\~a}lean}
+    \IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
+    \textit{Technical University of Cluj-Napoca}\\
+            Cluj-Napoca, Romania \\
+            Honoriu.Valean@aut.utcluj.ro}
 }
@@
 
+This paper is structured in few sections.
+Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulatino.
+Section \ref{sec:implementation} discuss the core approach of the current paper in architecting a software solution based on microservices for simulating traffic in a city.
+Section \ref{sec:orchestration} discuss the architectural aspect of orchestrating microservices using popular frameworks and solutions.
+Section \ref{sec:results} presents the results obtained after the implementation of some of the core concepts from the previous sections.
+
 \section{Existing solutions}
+\label{sec:existingsolutions}
 
@@
 
-\section{Microservices}
+\section{Implementation}
+\label{sec:implementation}
 
-\subsection{Current state}
+The entities which are participating into a traffic simulation are called actors.
+They are two: the city actor and the city simulator.
+The supporting containers are not themselves part of traffic simulation but they do have an important and supporting role.
+All of them are modeled as microservices and a short introductory is needed altought section \ref{sec:orchestration} and paragraph \ref{subsec:designchoices} offers a broader perspective around them.
+
+\subsection{Microservices}
 
 Microservices are not a new concept.
-The idea behind them has existed since Linux kernel has started to be enriched with a concept called namespaces \cite{wiki:linuxns}.
+The idea behind them has existed since Linux kernel has started to be enriched with a concept called namespaces \citep{wiki:linuxns}.
 This allows one set of processes to see one set of resources while another set of processes see another set of resources, where resources might be, but not limited to, process IDs, file names and network resources.
@@
 
-To go to solution for building, manipulating and running images is Docker \cite{docker}.
+To go to solution for building, manipulating and running images is Docker \citep{docker}.
 It allows easy software installation, it works cross platform and it offers a smooth experience most of the time.
-
-\section{Implementation}
-
-The entities which are participating into a traffic simulation are called actors.
-They are two: the city actor and the city simulator.
-The supporting containers are not themselves part of traffic simulation but they do have an important and supporting role.
 
@@
 The implementation proposed here is aiming to model a single type of actor: a car which moves between two points across a city.
-Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \cite{openstreetmap} and GraphHopper \cite{graphhopper}.
+Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \citep{openstreetmap} and GraphHopper \citep{graphhopper}.
 Open Street Map is an open source licensed map of the entire world.
@@
 
-\section{Orchestration}
+\subsection{Design choices}
+\label{subsec:designchoices}
 
-The term orchestration means the handling of containers and microservices in order to bring coherence into their interaction.
-They have to run on multiple nodes (computers or virtual machines) and at the same time to communicate with one another.
-They need to be updated in place and without service disruption.
-Whenever one of them crash they have to restart as quickly as possible.
-They have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc.
-They have to
+As with every software project started from scratch there are a number of choices to make when choosing software stacks, programming languages, networking models, etc.
+This section is aiming to explore the rationale behind some of those decisions and how they influenced the final application of building such an application.
 
-\subsection{Current state}
-
-\subsection{Implementation}
-
-\subsection{Docker compose}
-
-\subsection{Kubernetes}
-
-\section{Design choices}
-
-\subsection{Tools and programming languages}
 \begin{itemize}
 \item Go programming language
-\item Docker
-\item Docker compose
 \item Web page choices (web page will become an actor)
 \end{itemize}
-
-\subsection{Choices and rationales}
 
@@
 
+\section{Orchestration}
+\label{sec:orchestration}
+
+The term orchestration means the handling of containers and microservices in order to bring coherence into their interaction and an unifying experience to the end user of a service or of a product.
+They migth run on multiple nodes (computers or virtual machines) and at the same time they need to communicate with one another.
+They need to be updated in place and without service disruption.
+Whenever one of them crash they have to restart as quickly as possible.
+They have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc.
+Those are just few concerns around orchestration and the current paper is not aiming to provide a comprehensive view of what it means and what can be achieved with it but rather to set a basis of understanding enough context for the subject of simulating a city with its traffic.
+
+\cite{7922500} offers a comprehensive view of the current state of orchestration of cloud providers. (to add details)
+
+\subsection{Current state}
+
+As noted above the standard in easiness of developing and working with containers is Docker \citep{docker}.
+While it has an offering of orchestrating containers, called \textit{docker-compose}, which is simple to start with, its functionality is limited in comparison with other offerings, the most notable one being Kubernetes \citep{Kubernetes}.
+Those are not the single tools available and many more can be found but they offer a starting point (especially Docker) and Kubernetes, altough it has a steep learning curve, it does offer a comprehensive and complex panoply of details around orchestration.
+
+\subsection{Implementation}
+
+The city simulator is, currently, a single service which means it runs as a single instance.
+The city actor container runs in multiple instances and it has the need to do so as part of the entire simulation workload.
+The other two instances which run in the simulation are the routing service, a Graphhopper instance, and the front end web erver which displays a web page for offering visual clues around the simulation.
+While they run as such there should be an easy way, without friction and additional compute logic to "discover" a certain service.
+A city actor will need to create connect itself to the routing service and to the city simulator.
+At the same time the front end instance need to connect itself to the city simulator to source its data.
+
+\subsection{Docker compose}
+
+The \textit{docker-compose} tool gives the ability to manipulate more containers and services, with a simple file written in yaml format.
+Listing \ref{lst:dockercompose} shows the docker-compose file in its brevity and briefness.
+Let's dissect it.
+There are four services: \textit{graphhopper}, \textit{citysim}, \textit{cityactor} and \textit{cityfront}.
+Each of them need an image to run and this image can be specified in two ways: either as an image already compiled and hosted on an container registry or as a local folder which contains an file to build one (defined by the presence of a file called \textit{dockerfile}).
+Because Graphhoper, once compiled with desired maps and settings do not need any more development work, it is taken as an image from a docker registry.
+The other three services are the places where the most development efforts take place therefore they need to be compiled or recompiled each time the entire traffic simulation application starts.
+
+The containers which are running may need to have different properties.
+First and most important is the container port which needs to be published.
+The port from inside the container, where a certain process expect a TCP communication is forwarded to the local host port and thus is accessible from outside the Docker internal network.
+Another detail of the container is its dependency.
+For example \textit{cityactor} cannot run without textit{Graphhopper} because it doesn't have any place to obtain a route.
+Therefore it depends on \textit{Graphhopper}.
+And on \textit{citysim}, of course.
+The last bit of detail which can be seen here is the policy of restarting a container.
+By default the policy is set to "No" which means that the container will not be restarted if there is any failure within it and it stops.
+However, for simulation purposes, as we need a constant stream of city actors to swarm through the city, the policy is set to "Always" which will restart the container after if closes itself, i.e.
+it finishes its travel.
+
+\begin{lstlisting}[caption=Docker-compose file, label=lst:dockercompose]
+services:
+  graphhopper:
+    image: tomabecea/graphhopper:latest
+    ports:
+      - "8989:8989"
+  citysim:
+    build: ./city/citysim
+    ports:
+      - "9000:9000"
+  cityactor:
+    build: ./city/cityactor
+    depends_on: 
+      - "graphhopper"
+    restart: always
+  cityfront:
+    build: ./cityfront
+    ports:
+      - "80:80"
+    depends_on:
+      - "citysim"
+\end{lstlisting}
+
+Finally, to run this docker compose there is a simple command to bring everything to life: \textit{\textbf{docker-compose up}}.
+This will take everything which is in the \textit{docker-compose.yaml} file, will build the container if it has not been built before or if any source code file has been changed and then will run all of them in the order specified by the dependency graph.
+
+The other detail of running this set of containers is to scale up a specific service.
+In our case we would like to have more than one city actor.
+Therefore the command to run is \textit{\textbf{docker-compose up --scale cityactor=1001}}.
+This way one thousand and one city actors will be created.
+In combination with the option of always restarting a container which exits, the entire simulation will run virtually forever.
+
+\subsection{Kubernetes}
+
+While the docker compose offers a basic set of functionalities to bootstrap our application, for a large number of city actors a single computer might not suffice.
+Here comes Kubernetes in play.
+Kubernetes \citep{Kubernetes} is an open-source system for automating deployment, scaling and management of containerized applications.
+It originates from Google and is their third approach on orchestrating microservices, as described in \cite{burns2016borg}.
+
+While Docker-Compose runs on a single machine, Kubernetes is made to run on multiple machines called nodes.
+It is by default enriched with certain abilities needed to run in this configuration.
+Made initially to run with LXC and Docker containers \citep{7036275}, it is now a more open system where one can choose the container runtime interface, the container network interface, storage interface, service meshes, etc.
+from a broad range of vendors.
+On top of those base offerings, which provides only the backbone of Kubernetes, a distributed application must be built.
+The desired architecture of a Kubernetes application may consists of many concepts and building bricks.
+Using \cite{hightower2017kubernetes} as a starting point we can see that there are the most simple and basic units, called pods, which may encompass one or more containers (usually there is one main container and the others, e.g.
+service mesh supporting side car, are called side cars).
+There are deployments which gather together pods and rollout or rollback control.
+There are services which makes pods universally available into the cluster and makes them discoverable, regardless of what node are they running onto and abstracting away failures and upgrades.
+And then there are replica sets, daemon sets, secrets, config maps and many other resources.
+On top of those one is able to deploy own custom resource definitions as well as custom controllers.
+Put together, all those concepts have a rather intimdating allure.
+
+For running the entire traffic simulation onto a Kubernetes cluster we have the basic needs: all services are Docker images.
+Once each service is correctly described using a yaml format, each consisting of a service and deployment the entire application can be deployed via \textit{kubectl} command.
+The scaling problem is solved using a similar approach with \textit{docker-compose}.
+The command will be \textit{\textbf{kubectl scale --replicas=1001 deployment/cityactor}}.
+As before, with \textit{docker-compose}, it is to be noted the simple approach towards scaling: a core tenet of distributed systems, a property which, as it stands for our use case is quite simple to model.
+But for other applications, a more sensible approach is needed, as noted by \cite{jogalekar2000evaluating}.
+
 \subsection{Networking design}
+\label{subsec:networking}
 
-\subsection{Orchestration design}
+A distributed system has to carefully design the way in which its services communicates between them.
+As the nature of a distributed system is to run in multiple nodes or machines, it is obvious that the communication medium between them has to be one based on TCP/IP stack.
+Therefore the entire modeling of networking relies on this assumption.
+
+A simple and basic idea, noted here only to help on constructing the final proposed solution, is to have each service always located at a certain IP address and a certain port within a network.
+This means, for example, the city simulator will always be located at 192.168.0.101:7450.
+The other services which need to be accessed will listen on similar IP addresses and sockets.
+While this offers a convenient way to have them unified, they are not suitable to run in other environments than a development PC.
+In this case \textit{docker-compse} will run them such that they are accessible on local host address, i.e.
+127.0.0.1:7450.
+As soon as there are multiple nodes involved this model is not suitable anymore.
+
+Enter the DNS (Domain Name System), the backbone of the entire internet.
+In its simplest description, avoiding many inherent and nitpicking details, it is a dictionary which keeps track of every registered and easy to memorize name, e.g.
+en.wikipedia.com and the IP addresses where it is located.
+Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtain IP address.
+The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \citep{risley2001domain}.
+
+When many resources are needed, to serve a great number of users or to support a great number of city actors, a certain service might be so busy with serving data that the compute and memory resources it needs are greater than the hardware is able to give.
+Therefore it might be located at few addresses at the same time and any client which wants to access it should access the address where there are available compute and memory resources to serve its queries.
+Ideal would be to abstract or to decouple this information from the client and make it transparent for it.
+To do this the client needs to know only the service name and a certain "networking" entity should route its request to the appropiate available service.
+Such a entity could be a DNS authoritative server which, based on the load, availability and latency of the services will route the request to the appropiate node \citep{swildens2006scalable}.
+
+Docker and Kubernetes are doing a similar job.
+They offer a DNS service and according to the load of each node where a service run, a request is rooted to a node which will be able to respond to it.
+This logic is completely decoupled from the clients.
+Listings \ref{lst:gocityclientdial} and \ref{lst:cityfrontclientdial} shows how both the city front and a city actor calls the city simulator.
+All what they do is to use the address which was specified in the docker-compose file (listing \ref{lst:dockercompose}).
+Docker or Kubernetes will do the actual job to route the request to the appropiate node of the service.
+
+\begin{lstlisting}[caption=City client dialing City simulator, label=lst:gocityclientdial]
+conn, err := net.Dial("tcp", "citysim:7450")
+if err != nil {
+    fmt.Println("Error on dialing", err)
+    break
+}
+defer conn.Close()
+\end{lstlisting}
+
+\begin{lstlisting}[caption=City front dialing City simulator, label=lst:cityfrontclientdial]
+var startWebsocket = function (callback) {
+
+    var ws = new WebSocket("ws://citysim:9000/city")
+
+    ws.onopen = function(evt) {
+        console.log("OPEN");
+        ws.send("just sent some messageeeee")
+    }
+    ws.onclose = function(evt) {
+        console.log("CLOSE");
+        ws = null;
+    }
+    ws.onmessage = function(evt) {
+        callback(evt.data);
+    }
+    ws.onerror = function(evt) {
+        console.log("ERROR: " + evt.data);
+    }
+}
+\end{lstlisting}
 
 \section{Results}
+\label{sec:results}
 
@@
 \section{Enhancements}
+\label{sec:enhancements}
 (Distributed database for multiple city simulators)

### 2019-11-15 papers/traffic simulation/paper.tex
More writing on the traffic simulation paper.

@@
 \subsection{City simulator}
+\label{subsection:citysimulator}
 
@@
 
+\begin{figure*}
+  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface.png}
+  \centering
+  \caption{Screenshot of the web interface showing the running simulation}
+  \label{fig:ScreenshotWebInterface}
+\end{figure*}
+
+\begin{figure*}
+  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.png}
+  \centering
+  \caption{Screenshot of the web interface showing the running simulation with a larger number of actors}
+  \label{fig:ScreenshotWebInterface2}
+\end{figure*}
+
 \section{Orchestration}
@@
 \subsection{Kubernetes}
+\label{subsec:kubernetes}
 
@@
 Docker or Kubernetes will do the actual job to route the request to the appropiate node of the service.
+One thing to note is the difference between the two.
+A city client will communicate via TCP/IP using a Go language Glob format (add referenc to the design choices paragraph) while the city front use a WebSocket communication technology (add reference to the design choices paragraph) but both of them are transparently routed by Docker.
 
-\begin{lstlisting}[caption=City client dialing City simulator, label=lst:gocityclientdial]
+\begin{lstlisting}[caption=City actor dialing City simulator, label=lst:gocityclientdial]
 conn, err := net.Dial("tcp", "citysim:7450")
@@
 
-\subsection{Visual}
+Figure \ref{fig:ScreenshotWebInterface} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a simulation with less than 20 actors.
+The web interface is a simple, httml/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}.
+The city simulator will send a notification the web page server each time the density changes for a certain line.
+It can be that a new or first actor entered a certain line (a list of coordinates which represent a street) or it can be that one or the last actor left a line.
+The web page will draw any line with a density greater than 0 on top of the map.
 
-\subsection{Performance}
+The map is sourced from Open Street Map \citep{openstreetmap} and is showing the map zoomed to a specific city, same used by the city actors in their walkings.
+Thus, whenever a city actor reports that it is traveling across a line and this information arrives at the web interface through the city simulator, the respective line will be immediately drawn into the view.
+
+A laptop used for simulating (MacBook Pro, 16 GB RAM, i5 3.1 GHz) is able to run easily 50 city actors.
+Above a certain threshold the bottleneck, surprisingly, is not the memory or the cpu but the way docker handles network interfaces.
+For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}).
+Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time through the city.
 
@@
 \label{sec:enhancements}
-(Distributed database for multiple city simulators)
 
-(More granular logic on actor behavior)
+As with all technicl projects, there are a number of enhacements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
+This section will briefly propose few of those details in the following paragraphs.
 
-(More types of "junctions": classic junction for cars, pedestrian crossings, subways access stairs and elevators, bus stations, etc.)
+As the number of the actors are growing a first bottlneck in the current architecture will be the city simulator.
+It is a single entity and it represents a single source of truth for the actors.
+To keep it as single source of truth but at the same time to scale it horizontally means that we need to introduce few additional concepts.
+In \cite{6847479} a first idea might be of using multiple cloud providers for taking advantage of their available compute resources and also for combining them.
+This would serve both to the city simulator as well to the swarm of city actors.
+The other part which remains is to keep the city simulator as a single source of truth by taking advantage of a distributed database.
+Something similar can be seen in \cite{8509417} where a NoSQL database is used for horizontally scaling.
+To take advantage of the Kubernetes cluster there are similar horizontally scalable databases which can be built upon.
+One option is the distribute key-value store used by Kubernetes itself: Etcd datbase \citep{etcd}.
 
-(Contained actors: pedestrians in a bus, pedestrian in a subway)
+A second idea for enhancing the simulation is to make the actors to have a more granular and diverse logic of traversing the city.
+The solution proposed into this paper is a basic one: an actor has a route between two random points on the map and then it goes on to travel across that route.
+It can also be made to look for street densities in advance and act accordingly.
+If a street is too crowded the actor might choose to avoid it and compute a new route around it.
+
+A third idea is to add more interactions between actors.
+It has been discussed that a first interaction is and indirect one: the number of actors on a certain street, number called density.
+But there are many more points of interactions.
+Let's take junctions.
+The city simulator can model any junction and then it can permit actors to pass or not through it.
+Then there are pedestrian crossings where different types of actors can meet each other.
+Or subways access, elevators for them, bus stations, etc.
+
+A fourth idea is to have contained actors.
+A bus is an actor.
+A person is an actor.
+If a person is waiting on a bus station and then it takes a bus, there are two actors, but they are tied and are not independent anymore until the person choose to get out of the bus.
+This means the densities across sidewalkings and bus stations need to be computed accordingly.
+
+A fifth idea of enhancement is not tied to the way simulation is working but to how it is presented.
+The web page can display different colors based on the density of a line, offering a visual clue on it.
+Also it can easily show more information when the cursor is over a point of interest (a street, a junction, etc.).
 

### 2019-11-17 papers/traffic simulation/paper.tex
Finished the traffic simulation paper.

@@
 
-\title{A microservices based approach for city traffic simulation}
+\title{A Microservices Based Approach for City Traffic Simulation}
 
@@
 But for those areas to be tackled, appropiate methods of simulating traffic must be found.
-We are proposing a novel way of simulating traffic based on the microservices orchestration concept.
+We are proposing a novel way of simulating traffic based on the microservices orchestration concept and using a discrete microscopic logic.
 
@@
 
-(Summary of other papers with traffic simulation)
+Existing solutions have a general distinction of being on a spectrum: macroscopic, mesoscopic or microscopic.
+This distinction is caused by how the traffic is modeled: on a macroscopic scale like streets or sections of a highway or on microscopic level, focusing on each car and its relation with neighbours.
+The relation between a car and its neighbours, can also be modeled as microscopic or nanoscopic.
+
+\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with own personality and decisions.
+This idea is an interesting one especially because it is proposed as an enhacement to the solution presented by the current paper.
+In the same area of microscopic simulation we can see another proposal from \cite{7004985}.
+Of a great interest is their ability to feed back in the simulation the data gathered from loop detectors across the portion of highway they are particularly interested in.
+
+\cite{6205472} is probably the most viable approach as it is a complete simulation software.
+Based on the Ngel-Schreckenberg \citep{nagel-schreckenberg} automaton model of simulating a freeway traffic, P.
+Gora is advancing this idea into a fully fledged software simulation framework.
+
+In \cite{8569576} the broad idea is to us a discrete event architecture where there are logical processes for executing a number of agents.
+For a computation intensive those processes can be clustered into agent clusters with dependencies between them.
+The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.
+
+Finally, \cite{6244609} propose optimization computation for intersections (junctions) by using a microscopic approach towards data analysis.
+The proposal use a discrete and stochastic time step of simulation model, combined with Synchro, a signal timing optimization software.
 
@@
 Thus, a container is a small set of processes which run in isolation.
-They allow packaging a linux distro, a set of libraries and skd and on top of those custom code.
+They allow packaging a linux distro, a set of libraries and sdk and on top of those custom code.
 This forms an image, which is essentially a tar gzipped file.
@@
 As with every software project started from scratch there are a number of choices to make when choosing software stacks, programming languages, networking models, etc.
-This section is aiming to explore the rationale behind some of those decisions and how they influenced the final application of building such an application.
+This section is aiming to explore the rationale behind some of those decisions and how they influenced the building of  such an application.
 
-\begin{itemize}
-\item Go programming language
-\item Web page choices (web page will become an actor)
-\end{itemize}
+A first choice is the programming language to write both the city simulator and the city actor.
+Go programming langauge is born out of Google and it resembles the philosophy they were trying to embedded in it for taming the complexity of their systems \citep{donovan2015go}.
+Today it has gained a lot of popularity and tools like Kubernetes are wrritten entirely in Go, making it the default language of the cloud technologies.
+It is a "C-like" language and it has a familiar look but with few traits which makes it different.
+Out of those, few are notable, not only because they facilitate programming endeavors but also because both city simulator and city actor are modeled around them in their communications.
 
-\subsection{Alternatives}
-(Actor model and consacrated frameworks)
+\begin{lstlisting}[caption=Go routine from city actor, label=lst:goroutine]
+go func() {
+	for {
+		select {
+		case <-ticker.C:
+			advance(city, reportChan, lineChan)
+		case _ = <-lineChan:
+			//fmt.Println("Received answer with line", j)
+		}
+	}
+}()
+\end{lstlisting}
+
+Go routines are a lightweight thread of executing.
+Being lightweight it means they can be easily started, with no overhead and especially without the ceremony of dealing with threads which other popular languages (e.g.
+Java) have.
+Listing \ref{lst:goroutine} shows such a routine.
+It is a part of logic where the city actor responds to either two events: the timer has expired and a decision has to be made or an information about the line is currently traveling on has been received from the city simulator.
+Another Go trait is the channels, an indexed communication pipe where there are asynchronous writers and readers, used to decouple two routines.
+Listing \ref{lst:goroutine} also shows couple of channels.
+Variables \textit{ticker} and \textit{lineChan} are two channels.
+They are declared using a type and after that they are used to write and read objects (or structs) of the respective type.
+Combined with the \textit{select} keyword, it becomes easy to write code which takes advantage of multithreading paradigms.
 

### 2019-12-08 papers/traffic simulation/paper.tex
Some corrections.

@@
 Simulating traffic in a city is a difficult task.
-Current paper is proposing a novel way using state of the art in distributed systems: microservices orchestration.
+Current paper is proposing a novel way using the state of the art in today's distributed systems: microservices orchestration.
 The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation.
@@
 It can help understand patterns of traffic and its flow.
-It can help unerstand the particularities and pecularities of a city's streets arrangements.
+It can help understand the particularities and pecularities of a city's streets arrangements, together with their junctions.
 It can help identify bottlenecks.
 It can help find solutions to rush problems and explore them.
-But for those areas to be tackled, appropiate methods of simulating traffic must be found.
-We are proposing a novel way of simulating traffic based on the microservices orchestration concept and using a discrete microscopic logic.
+But for those areas to be tackled, appropiate methods of simulating traffic must be found and explored.
+We are proposing a novel way of simulating traffic based on the microservices orchestration concept and using a discrete microscopic logic for the behavior of each actor.
 
-We define a \textbf{(city) actor} as being a independent entity which chose to move between two geographical points within a city.
+We define a \textbf{(city) actor} as being an independent entity which chose to move between two geographical points within a city.
 It can be a car, a pedestrian or a bike.
@@
 This paper is structured in few sections.
-Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulatino.
+Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulation.
 Section \ref{sec:implementation} discuss the core approach of the current paper in architecting a software solution based on microservices for simulating traffic in a city.
@@
 
-\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with own personality and decisions.
-This idea is an interesting one especially because it is proposed as an enhacement to the solution presented by the current paper.
+\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with its own personality and decisions.
+This idea is an interesting one especially because it is proposed as an enhancement to the solution presented by the current paper.
 In the same area of microscopic simulation we can see another proposal from \cite{7004985}.
@@
 
-In \cite{8569576} the broad idea is to us a discrete event architecture where there are logical processes for executing a number of agents.
-For a computation intensive those processes can be clustered into agent clusters with dependencies between them.
+In \cite{8569576} the broad idea is to use a discrete event architecture, in which there are logical processes for executing and simulating a number of agents.
+For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them.
 The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.
@@
 Thus, a container is a small set of processes which run in isolation.
-They allow packaging a linux distro, a set of libraries and sdk and on top of those custom code.
-This forms an image, which is essentially a tar gzipped file.
+They allow packaging a linux distro, a set of libraries and a software development kit and on top of those the custom code of an application.
+Taken together, those form an image, which is essentially a tar gzipped file.
 Once the build process of an image is finished, it can be spinned up in one or more running containers.
@@
 In the current implementation the city simulator keeps a set of data which can be described as a list of pairs, each pair having a line and a real number, called density.
+The line is a series of coordinates and in the proposed implementation their meaning is a street which a city actor is reportedly traveling on it.
 
@@
 A notable exception is the web page.
-Its purpose is not to participate in the same information exchange the other entities have but to offer a visual interpretation of the way the city and the other actors are interacting with one another.
+Its purpose is not to participate in the same information exchange the other entities have but to offer a visual interpretation about the way the city and the other actors are interacting with one another.
 
@@
 Its naming suggests that it can be any entity or living being which moves within a city and interacts with the other entities or affects the other entities in some manner.
-A pedestrian would directly interact with other pedestrians but not with cars.
+A pedestrian would directly interact with other pedestrians but not with cars unless it crosses the street on red or on unmarked places.
 A pedestrian would indirectly affect other cars by willing to walk over a street crossing.
@@
 It can use Open Street Map as an underlying map provider.
-They provider free tiers and paid subscription for accessing an API and get routes.
+They provide free tiers and paid subscription for accessing an API and compute various routes.
 However, a city actor is not using any public api but a special crafted container which contains GraphHopper and an open street map embedded into it.
@@
 The message sent from one side to another has a general structure called \textit{Envelope}.
-It is meant to offer a top level message which can be serialized (or encoded, the way to do this in Golgang, if not gRPC, is the gob pacakge it offers out of the box) and passed around, enabling any party involved to understand what this message is about and how to decode it.
+It is meant to offer a top level message which can be serialized (or encoded, the way to do this in Golgang, if not gRPC, is the gob pacakge offered out of the box as a base package within the lanugage) and passed around, enabling any party involved to understand what this message is about and how to decode it.
 
@@
 \textit{AskForLine} is a message sent from an actor to the city simulator in which the actor asks about the density of any line.
-This way any actor can take conscious decisions on which route to go on when based on what lies ahead in terms of street densities.
-When a first route is known, between two desired points, the actor can ask about each line it has on that route.
+This way any actor can take conscious decisions on which route to go on, based on what lies ahead in terms of street densities.
+When a first route is chosen, between two desired points, the actor can ask about each line which is part of that route.
 The city simulator will respond with the known density of it.
-If the actor desires, it can try to find another route by asking GraphHopper service to compute a new one but with an additional rule: avoid a certain point (street, junction, etc.).
+If the actor desires, it can try to find another route by asking GraphHopper service to compute a new route but with an additional rule: avoid a certain point (street, junction, etc.).
 \textit{RespondWithLine} is the type of message which the city simulator sends back to an actor after it received an \textit{AskForLine} message.
@@
 As with every software project started from scratch there are a number of choices to make when choosing software stacks, programming languages, networking models, etc.
-This section is aiming to explore the rationale behind some of those decisions and how they influenced the building of  such an application.
+This section is aiming to explore the rationales behind some of those decisions and how they influenced the building of the current prototype.
 
@@
 Listing \ref{lst:goroutine} shows such a routine.
-It is a part of logic where the city actor responds to either two events: the timer has expired and a decision has to be made or an information about the line is currently traveling on has been received from the city simulator.
+It is a part of logic where the city actor responds to either two events: the timer has expired and a decision needs to be made or an information about the line is currently traveling on has been received from the city simulator.
 Another Go trait is the channels, an indexed communication pipe where there are asynchronous writers and readers, used to decouple two routines.
@@
 They are declared using a type and after that they are used to write and read objects (or structs) of the respective type.
-Combined with the \textit{select} keyword, it becomes easy to write code which takes advantage of multithreading paradigms.
+The \textit{select} keyword acts like a switch, not on variables and their values but on channels.
+It will execute the block of code for the channel which has something new to deliver.
+All those three concepts, \textit{goroutines}, \textit{channels} and \textit{select} switches, combined together makes it easy to write code which takes advantage of multithreading paradigms and which has to run in a heavy networking environment.
 
@@
 Whenever one of them crash they have to restart as quickly as possible.
-They have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc.
+The services have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc.
 Those are just few concerns around orchestration and the current paper is not aiming to provide a comprehensive view of what it means and what can be achieved with it but rather to set a basis of understanding enough context for the subject of simulating a city with its traffic.
@@
 
-The city simulator is, currently, a single service which means it runs as a single instance.
+The city simulator is, currently, a single service which means it runs as a single instance as viewed by a city actor.
 The city actor container runs in multiple instances and it has the need to do so as part of the entire simulation workload.
-The other two instances which run in the simulation are the routing service, a Graphhopper instance, and the front end web erver which displays a web page for offering visual clues around the simulation.
+The other two instances which run in the simulation are the routing service, a Graphhopper instance, and the front end web erver which displays a web page for offering visual clues about how the simulation is running.
 While they run as such there should be an easy way, without friction and additional compute logic to "discover" a certain service.
-A city actor will need to create connect itself to the routing service and to the city simulator.
+A city actor will need to connect itself to the routing service and to the city simulator.
 At the same time the front end instance need to connect itself to the city simulator to source its data.
@@
 There are four services: \textit{graphhopper}, \textit{citysim}, \textit{cityactor} and \textit{cityfront}.
-Each of them need an image to run and this image can be specified in two ways: either as an image already compiled and hosted on an container registry or as a local folder which contains an file to build one (defined by the presence of a file called \textit{dockerfile}).
-Because Graphhoper, once compiled with desired maps and settings do not need any more development work, it is taken as an image from a docker registry.
+Each of them need an image to run and this image can be specified in two ways: either as an image already compiled and hosted on an container registry or as a local folder which contains a file to build one (usually named \textit{dockerfile}).
+Because Graphhoper, once compiled with desired maps and settings do not need any more development work, it is uploaded to a personal docker registry and taken from there, whenever needed.
 The other three services are the places where the most development efforts take place therefore they need to be compiled or recompiled each time the entire traffic simulation application starts.
@@
 For example \textit{cityactor} cannot run without textit{Graphhopper} because it doesn't have any place to obtain a route.
-Therefore it depends on \textit{Graphhopper}.
+Therefore it depends on \textit{Graphhopper}, i.e.
+it waits for GraphHopper to start first.
 And on \textit{citysim}, of course.
@@
 In its simplest description, avoiding many inherent and nitpicking details, it is a dictionary which keeps track of every registered and easy to memorize name, e.g.
-en.wikipedia.com and the IP addresses where it is located.
-Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtain IP address.
+\textit{en.wikipedia.com} and the IP addresses where it is located.
+Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtained IP address.
 The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \citep{risley2001domain}.
 
-When many resources are needed, to serve a great number of users or to support a great number of city actors, a certain service might be so busy with serving data that the compute and memory resources it needs are greater than the hardware is able to give.
+When many resources are needed, to serve a great number of users or to support a great number of city actors, a certain service might be so busy with serving data that the compute and memory resources it needs are greater than the underlying hardware is able to support.
 Therefore it might be located at few addresses at the same time and any client which wants to access it should access the address where there are available compute and memory resources to serve its queries.
@@
 Listings \ref{lst:gocityclientdial} and \ref{lst:cityfrontclientdial} shows how both the city front and a city actor calls the city simulator.
-All what they do is to use the address which was specified in the docker-compose file (listing \ref{lst:dockercompose}).
+All what they do is to use the address, the service name, which was specified in the docker-compose file (listing \ref{lst:dockercompose}).
 Docker or Kubernetes will do the actual job to route the request to the appropiate node of the service.
 One thing to note is the difference between the two.
-A city client will communicate via TCP/IP using a Go language Glob format (add referenc to the design choices paragraph) while the city front use a WebSocket communication technology (add reference to the design choices paragraph) but both of them are transparently routed by Docker.
+A city client will communicate via TCP/IP using the Go default serialization package while the city front use a WebSocket communication technology (add reference to the design choices paragraph) but both of them are transparently routed by Docker.
 
@@
 Figure \ref{fig:ScreenshotWebInterface} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a simulation with less than 20 actors.
-The web interface is a simple, httml/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}.
-The city simulator will send a notification the web page server each time the density changes for a certain line.
+The web interface is a simple: a html/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}.
+The city simulator will send a notification to the web page server each time the density changes for a certain line.
 It can be that a new or first actor entered a certain line (a list of coordinates which represent a street) or it can be that one or the last actor left a line.
@@
 For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}).
-Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time through the city.
+Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time throughout the city.
 
@@
 
-As with all technicl projects, there are a number of enhacements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
+As with all technicl projects, there are a number of enhancements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
 This section will briefly propose few of those details in the following paragraphs.
@@
 If a street is too crowded the actor might choose to avoid it and compute a new route around it.
+A set of actors can be made to always run between same points (to simulate, for example, buses).
 
 A third idea is to add more interactions between actors.
-It has been discussed that a first interaction is and indirect one: the number of actors on a certain street, number called density.
+It has been discussed that a first interaction is an indirect one: the number of actors on a certain street, a number called density.
 But there are many more points of interactions.
@@
 A person is an actor.
-If a person is waiting on a bus station and then it takes a bus, there are two actors, but they are tied and are not independent anymore until the person choose to get out of the bus.
+If a person is waiting on a bus station and then it takes a bus, there are two actors, but they are tied together, or contained and are not independent anymore until the actor-person choose to get out of the actor-bus.
 This means the densities across sidewalkings and bus stations need to be computed accordingly.
@@
 Also it can easily show more information when the cursor is over a point of interest (a street, a junction, etc.).
+Or it can speed up a past simulation and it can display an animation.
 

### 2019-12-25 thesis/report1/paper.tex
Added files for report nr 1.

\documentclass[a4paper,12pt,twoside]{book}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}

\pagestyle{myheadings}

\title{Opportunities on }
\author{Toma Becea}

\begin{document}

\maketitle

\chapter{Using different page styles}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\chapter{Sample Chapter}
\thispagestyle{empty}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\clearpage

\section{New section}

\Blindtext

\end{document}

### 2019-12-29 thesis/report1/paper.tex
Introduction done. Abstract almost done.

@@
 
-\title{Opportunities on }
+\title{Analyzing opportunities arround city traffic optimization }
 \author{Toma Becea}
@@
 
-\chapter{Using different page styles}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter{Sample Chapter}
-\thispagestyle{empty}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
 \clearpage
 
-\section{New section}
+\section{Introduction}
 
-\Blindtext
+The traffic in a city is a very complex phenomenon.
+The parties involved are of many types: cars, pedestrians, buses, trams, taxi cabs (they are arugably cars, too but there might be small differences, e.g.
+they are allowed to drive on public transport lanes), bicycles, e-scooters (again, they are arguably same as bicycles), etc.
+Then there are rules on how they move: traffic lights for cars in junctions, across roads for pedestrian crossings, different traffic lights for buses and trams, etc.
+Within this rule-bounded context is the free will of every intelligent decision maker (it might be a human or it might be a self driving car with algorithms which we can deem as being intelligent from the perspective of behaving and moving within the traffic), the individual randomness unified into a collective orderness.
+
+But the devil is in the details: the orderness we see when looking across the entire city uncovers many opportunities.
+The architecture of many old cities is not meant for current traffic needs.
+The behavior of residents and dwellers of a city is not optimized for collective good but for personal comfort.
+The heterogenous and continous development of a city might add pain to the picture.
+Highly-accelerated developing areas or neighborhoods pose problems which city administrations are slow to understand and tackle with.
+Those, so far, are problems on the macro scale of a spectrum.
+On the other side of the spectrum, called micro, we can find another set of opportunities which can be modeled and analyzed through computer based simulations.
+Junction design, traffic lights programming and synchronization, pedestrian crossings and their influence on the surrounding driving lanes, bus stops and their influence, side walks design, lanes formation and one way drives, etc.
+Those are localized problems, focused on small geographical areas.
+
+The current chapter aim to find and define technical means on exploring intricacies within the traffic of a city.
+Software simulation is a broad term which encompasses any idea or way of using a computer to mimic as closely as possible a real phenomenon, with or without the help of a mathematical apparatus.
+This entire chapter is centered around various tehniques of using a computer for traffic simulation.
+There are a number of existing solutions to simulating traffic using software abilites, on various places on the macro-micro spectrum, as well as on other spectrums like discrete event vs continous simulation.
+Those will be explored and briefly categorized.
+
+\section{Abstract}
+
+The current chapter is aiming to propose a novel solution for simulating traffic in a city and exploring the opportunities which this solution makes possible to address.
+To describe it in few words, before going in full details on it on section \ref{sec:microservicesbasedapproachtrafficsim}, it is a discrete event based simulation, using a distributed microservices architecture where actors (defined as independent and intelligent participants in a traffic) are moving across a city and interacting with one another.
+
+The second area of this chapter is to analyze opportunities for optimizing the traffic within a city using the aforementioned simulation solution.
+Those ideas will revolve around the core concepts which form the foundation of the microservices base approach for traffic simulation: discrete and independent actors, decoupled in their logic and choice, interacting via discrete events based on their interest.
+
+\section{Current state of traffic simulation}
+
+\section{A microservices based approach for traffic simulation}
+\label{sec:microservicesbasedapproachtrafficsim}
 

### 2019-12-31 thesis/report1/paper.tex
Update paper.tex
Few more ideas on report nr 1.

@@
 The current chapter is aiming to propose a novel solution for simulating traffic in a city and exploring the opportunities which this solution makes possible to address.
-To describe it in few words, before going in full details on it on section \ref{sec:microservicesbasedapproachtrafficsim}, it is a discrete event based simulation, using a distributed microservices architecture where actors (defined as independent and intelligent participants in a traffic) are moving across a city and interacting with one another.
+To describe it in few words, before going in full details on it during section \ref{sec:microservicesbasedapproachtrafficsim}, it is a discretes event based simulation, using a distributed microservices architecture where actors (defined as independent and intelligent participants in a traffic) are moving across a city and interacting with one another.
+They are a number of different instances running in isolation as containers.
+Their interaction is made via a single instance, called city simulator, which keep various statistics about the actors which are live and moving.
+On top of those there are few other supporting utilities, not essential for the current subject but mandatory for the simulation to run.
+They are a map and routing service which offers a route between two points of a city to those actors which need it and a service which renders a web page with the map of the city where the simulation takes place and the gradual occupation of its streets.
 
 The second area of this chapter is to analyze opportunities for optimizing the traffic within a city using the aforementioned simulation solution.
-Those ideas will revolve around the core concepts which form the foundation of the microservices base approach for traffic simulation: discrete and independent actors, decoupled in their logic and choice, interacting via discrete events based on their interest.
+Those ideas will revolve around the core concepts which form the foundation of the microservices based approach for traffic simulation: discrete and independent actors (also randomness is included into their behavior), decoupled in their logic and choice, interacting via discrete events based on their selfish interest (their behiavor which this simulation is modeled for is their interest in navigating their route as fast as they can, using the shortest path between two points, departure and arrival ones).
+
+Following concepts will be discussed in order to explore the opportunities for optimizing the traffic in a city.
+They are to be found in section \ref{sec:citytrafficoptimizations} and a brief list of them is presented here:
+
+\begin{itemize}
+    \item Modeling various actors with very different behaviors. For example a pedestrian and a city bus. Or a bicycle and a garbage truck.
+    \item Modeling various relations between different actors. An example of such a relation can be contained actors: a pedestrian can take a tram.
+    \item Introducing exception actors and their influence in the surrounding traffic. Firefighters truck is one example. Cars stopped on the road is another.
+    \item Localize hot areas or bottlenecks of a city: junctions, streets, cars stopped on the road.
+    \item Computation of various numbers and statistics. Few examples: average speed on street intervals, cars density per area and/or time, 
+\end{itemize}
 
@@
 
+\section{City traffic optimizations}
+\label{sec:citytrafficoptimizations}
+
 \end{document}

### 2020-01-01 thesis/report1.tex
Added wiki.bib and moved the report one folder up.

--- 2019-12-31 thesis/report1/paper.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{apacite}
+\usepackage{natbib}
 
+\bibliographystyle{apacite}
 \pagestyle{myheadings}
@@
     \item Modeling various relations between different actors. An example of such a relation can be contained actors: a pedestrian can take a tram.
-    \item Introducing exception actors and their influence in the surrounding traffic. Firefighters truck is one example. Cars stopped on the road is another.
-    \item Localize hot areas or bottlenecks of a city: junctions, streets, cars stopped on the road.
-    \item Computation of various numbers and statistics. Few examples: average speed on street intervals, cars density per area and/or time, 
+    \item Introducing exception actors and their influence in the surrounding traffic. Firefighters truck moving throughout the city is one example. Cars stopped on the road is another.
+    \item Localize hot areas or bottlenecks of a city: junctions, streets, bridges.
+    \item Computation of various numbers and statistics. Few examples: average speed on street intervals, cars density per area and/or time.
+    \item Introduce geographical differences in simulation and study their influence: a roundabout instead of a junction, a new lane on a street.
 \end{itemize}
@@
 
+\subsection{Prerequisites}
+\label{subsec:prerequisites}
+In order to discuss how a discrete-based, microservices architectured simulation works, few prerequisites are needed to set the base.
+
+First definition we need is for setting the fact that we are using geographical coordinate system.
+It doesn't matter which one, what it matters is that all actors are using the same set of geographical coordinates and same projection (or to know if there should be a conversion).
+For example, they could be EPSG:4326 or EPSG:3857.
+To define this in an appropiate way, equation \ref{eq:coordinates} will be used.
+Its loosness allow any implementation to treat coordinates using a built in language data type (e.g.
+\textit{double} in C\#).
+
+\begin{equation}
+\label{eq:coordinates}
+    C = \{ c | c \in \Re, c \geq 0 \}
+\end{equation}
+
+Simulation wise, actors do not have any \textit{intent}.
+They are just \textit{dumb} programs which need to be programmed to have any small \textit{intention}.
+To give them a \textit{sense} of selfishness they are programmed to navigate between two random points.
+Therefore points need to be defined as a pair of coordinates.
+The implementation of this will have to take care of their order or of knowing which one represents latitude and which one longitude.
+Equation \ref{eq:coordinatespair} defines such a pair of coordinates.
+Furthermore we define a way which introduce randomness into an actor's behavior: the random choice of two points.
+The actor needs to have two points predefined, $Pd_1$ and $Pd_2$.
+Those are treated as an aproximative rectangular area, out of which other two points will be randomly chosen: $Pr_1$ and $Pr_2$.
+Equation \ref{eq:funcrandom} defines this function, without entering into the details of how such a random function would work.
+
+\begin{equation}
+\label{eq:coordinatespair}
+    P = \{ (p1, p2) | p1 \in C, p2 \in C \}
+\end{equation}
+
+\begin{equation}
+\label{eq:funcrandom}
+\begin{split}
+    & f_{rand}: P \rightarrow P, \\
+    & f_{rand}(Pd_1, Pd_2) = rand (Pd_1, Pd_2) \Rightarrow (Pr_1, Pr_2)
+\end{split}
+\end{equation}
+
+So far an actor has two random points and now it needs a route.
+For this the final function we need to define as a prerquisite is in equation \ref{eq:funcroute}.
+Essentially this function takes two points, in this case the two random points obtained earlier, in equation \ref{eq:funcrandom}: $Pr_1$ and $Pr_2$ and outputs a series or a list of points where first point is the first random point, $Pr_1$ and last point is the second random point, $Pr_2$.
+Between them a number of points represents the path which the actor should follow in order to travel from its first (randomly choosen) point to its second (randomly choosen) point.
+Translated into a more intuitive description, those points are the instruction which any maps application outputs when asked for a route.
+For example OpenStreetMap, \cite{openstreetmap}.
+
+\begin{equation}
+\label{eq:funcroute}
+\begin{split}
+    & f_{route}: P \rightarrow [P], \\
+    & f_{route}(Pr_1, Pr_2) = route (Pr_1, Pr_2) \Rightarrow [Pr_1, Pc_1, Pc_2, ..., Pc_n, Pr_2]
+\end{split}
+\end{equation}
+
+\subsection{Microservices - city actor}
+
+Having defined the prerequisites from section \ref{subsec:prerequisites} we move on to present the first type of microservice from our architecture: the (city) actor.
+A city actor is meant to represent a specific behavior of an entity which moves across the city.
+To create such an actor it must be infused with specific behaviors based on the reality it wants to mimic.
+The prevalent example of this chapter will be a car.
+Therefore the prerequisites from section \ref{subsec:prerequisites} will suit a car's purpose to obtain a route using a predefined area where the car is allowed to wander.
+
+\subsection{Microservices - city simulator}
+
 \section{City traffic optimizations}
@@
 
+\bibliography{wiki}
+
 \end{document}

### 2020-01-10 thesis/report1.tex
Report 1 advancement.

@@
 \usepackage{natbib}
+\usepackage{listings}
 
@@
 Translated into a more intuitive description, those points are the instruction which any maps application outputs when asked for a route.
-For example OpenStreetMap, \cite{openstreetmap}.
+For example OpenStreetMap, \cite{openstreetmap} is a map provider and Graphhopper, \cite{graphhopper} is a route provider which uses OpenStreetMap as the default map.
 
@@
 Therefore the prerequisites from section \ref{subsec:prerequisites} will suit a car's purpose to obtain a route using a predefined area where the car is allowed to wander.
+To further isolate its movements and decisions an algorithm is supposed to be used at its core.
+Listing \ref{lst:caractoralgorithm} presents it.
+
+\begin{lstlisting}[caption=Basic algorithm for a car actor, label=lst:caractoralgorithm]
+\\ Get two random points
+(pr1, pr2) = rand (pd1, pd2)
+\\ Get the route
+R = route (pr1, pr2)
+\\ Set the position to the first point
+P my_position = pr1
+int index = 0
+while (index < R.length) {
+    int speed_coeff = get_info(R[index])
+    my_position = advance(speed_coeff, R, index)
+    report(my_position)
+}
+\\ Report that the route has been finished
+report(no_position)
+\end{lstlisting}
+
+Few remarks about the core algorithm showed in listing \ref{lst:caractoralgorithm}.
+There are few pieces which are missing from the prerequisites but are discussed into section \ref{subsec:citysimulator} because they are related to the communications between a car actor and the city simulator.
+Those pieces are the calls to following methods: \textit{get\_info} and \textit{report}.
+The third one, \textit{advance} is the place where each actor is going to be personalized with own rules and behavior.
+Those aspects will be discussed into the subsection \ref{subsec:actordiffbehaviors}.
+
+The core algorithm needs to be run on continuosly basis, with a standard frequency (e.g.
+500 ms).
+Varying this number, also called simulation step can be an option for speeding or slowing the simulation running time.
+The \textit{advance} method is supposed to compute where to go next and it is responsible for incorporating the \textit{speed\_coeff} coefficient obtained from the city simulator.
+This coefficient is meant to express the density of cars on the street (if the actor which needs it is a car.
+In the case where it is a pedestrian the coefficient will be different.
+This way it is intuitively described the fact that a pedestrian walking on the sidewalk has nothing to do with the cars which are driving on near lanes and viceversa, a car driving on some lane doesn't care how many pedestrians are on the sidewalk near it) and it varies from expressing a totally free road to a road blocked by the numbers of cars waiting to pass through a junction.
+
+A written, more expansive description of this algorithm is as follows.
+For easier following of the simulation and for the current paper all actors are bounded to a certain area, expressed as two defined points $Pd_1$ and $Pd_2$.
+Those points are treated as the opposing corners of a rectangle and another two random points are chosen such that they are wihtin the rectangle: $Pr_1$ and $Pr_2$.
+The actor will then proceed to obtain a route, from another service, in order to know how to travel from $Pr_1$ to $Pr_2$.
+The route obtained is a series of another points, $Pc_1$ tp $Pc_n$, starting from $Pr_1$, called departure point, right to the finish point: $Pr_2$.
+They can be organized into subsets.
+A subset is a number of consecutive points which are treated as one instruction.
+Those instructions are meant to offer an easier guide, vocal or visual, for a human traveler.
+However, there are no human travelers into the simulation so the instructions have found another useful role.
+Usually an instruction is a limited length of a street during which the traveler doesn't need to do any change of direction.
+As soon as a junction must be traveled, there will be another instruction.
+Therefore this subset of the entire vector of points will become the interval which the actor reports as being traveling on to the city simulator.
+The city simulator, having such reports from all actors, will respond with the a number called speed coefficient, which reflects how many other actors are on the same interval and it will influence the traveling speed.
+The $advance$ method is meant to compute how the actor will advance and is the main place, though not the only one, where the behavior of different types of actors will be implemented.
+This method also takes care of handling the speed coefficient.
+Once the method computes the next position where the actor will be, that position will be reported back to the city simulator.
+This logic repeats itself as long as the actor has points and instructions which haven't been traveled yet.
+At last, the actor will report that it has been finished its route so that the city simulator will take into account one less actor.
 
 \subsection{Microservices - city simulator}
+\label{subsec:citysimulator}
+
+In this first chapter the city simulator is a separate entity which takes care of all the reports the actors send to it and also respond with every query about a certain route. Thus, it will need a
+
+Speed coefficient
+
+\subsection{Implementation}
 
@@
 
+\subsection{Actor with different behaviors}
+\label{subsec:actordiffbehaviors}
+
 \bibliography{wiki}

### 2020-01-12 thesis/report1.tex
Writtern more about the city simulator into the refferal 1.

@@
 
-In this first chapter the city simulator is a separate entity which takes care of all the reports the actors send to it and also respond with every query about a certain route. Thus, it will need a
+In this first chapter the city simulator is a separate entity which takes care of all the reports the actors send to it and also respond with every query about a certain route.
+Thus, it will need a way to keep this data.
+There are few assumptions around this which need to be taken into consideration:
 
-Speed coefficient
+\begin{itemize}
+    \item The set of points $P$ represents routes whose interpretation are portions of streets;
+    \item A report sent by an actor contains a set of points and an indication if the actor is currently travelling on that route or if it went off it;
+    \item A set of points which was received from an actor can be a subset of another set of points received earlier from another actor;
+    \item A set of points which was received from an actor can be a superset of another set of points received earlier from another actor;
+    \item An actor can ask various information for any set of points it needs;
+    \item Any information which will be conveyed to an actor must take into account any subset and superset which might already exist;
+    \item An actor will send in each report an piece of information representing its type (car, person, etc.);
+\end{itemize}
+
+To illustrate above points the speed coefficient will be discussed.
+The first and uttermost information which an actor needs is representing the concentration of other actors of its type which around it or on a certain piece of road.
+For example, a car needs to know how many cars are around it, to know how fast it can advance.
+Or it might need to know how many cars are on a certain piece of road to asses if it can find a faster route.
+While the latter example is an intuitive one the former needs a little more explanation.
+Being a simulation, an actor doesn't know how many other actors are around it.
+In a real scenario the driver just have to look around but here there are no drivers and no looking around.
+One solution is for an actor to ask the city simulator how many other actors of its type are on the same street it's currently traveling on.
+Therefore, this information will need to be baked in the each actor type.
+For example a car will need to adjust its nominal speed with the coefficient which it received from the city simulator.
 

### 2020-01-14 thesis/report1.tex
Some more words written on the referral 1, implementation subsection.

@@
 
+The following implementation discuss the solution chosen by the authors in \cite{microservicestrafficsimulation}.
+
+First, a brief and high level view for few of the terms which will be used henceforth.
+A \textbf{microservice} is small and contained application, embedded into a tar file which contains everything it needs to run by itself.
+Moving tar files (called images) and running an image multiple times on the same machine gives a greater flexibility in developing and deploying application where the nature of distributin and scaling matters.
+A \textbf{container} is a running image.
+\textbf{Docker} \citep{docker} is a commercial (with free tier offers) for building, manipulating and running containers with limited functionality on the running side.
+
+Each actor type and the city simulators are an image.
+When they run, they become containers.
+Thus a simulation with hundreds of actors of car type needs only one image, that of a car actor.
+On top of this the second image comes into play: this is different and its called city simulator.
+It will run only as one instance or one container.
+The running environment, called Docker, will take care of running those two images and any other which are needed.
+Docker also offers a running environment where the containers can see each other through assigned names.
+When defining a running instance of type city simulator it can be specified that its name is \textit{citysimulator}.
+Docker will take care of providing a DNS-like experience for finding where this instance is running.
+Any other container which wants to talk to the \textit{citysimulator} will just use its name into the URL it wants to access (provided that the )
+
 \section{City traffic optimizations}
@@
 
-\subsection{Actor with different behaviors}
+\subsection{Actors with different behaviors}
 \label{subsec:actordiffbehaviors}
+
+\subsection{Analyzers}
 

### 2020-01-16 thesis/report1.tex
Progress on referral 1.

@@
 \subsection{Microservices - city actor}
+\label{subsec:cityactor}
 
@@
 
-The following implementation discuss the solution chosen by the authors in \cite{microservicestrafficsimulation}.
+The following implementation briefly presents the solution chosen by the authors in \cite{microservicestrafficsimulation}.
 
@@
 Docker will take care of providing a DNS-like experience for finding where this instance is running.
-Any other container which wants to talk to the \textit{citysimulator} will just use its name into the URL it wants to access (provided that the )
+Any other container which wants to talk to the \textit{citysimulator} will just use its name into the URL it wants to access (provided that the type of communications are REST APIs).
+Thus the easiness in use for container to container communications removes many frictions and smooths activities such as architecting a software application or developing it.
+
+Beyond the two images, those of a city car actor and a city simulator, two others are needed for the simulation to actually take place.
+One of them is a routing engine.
+It uses an underlying map, taken from Open Street Map \citep{openstreetmap}.
+The routing engine itself is called Graphhoper \citep{graphhopper} and offers routing services like finding the best (shortest) route between two points while at the same time providing useful options like computing the routing with the avoidance of a certain point (useful for actors which would like to avoid a too crowded street).
+The other image used is a web server which sources its data from the city simulator instance and presents a map where a human user can see the live simulation of the city.
 
@@
 
+One good opportunity for studying the traffic in a city is to model different actor types, each with different behaviors and travel rules.
+They all share same common goal: travel a route in as little time as possible.
+But the means of traveling (car, bus, pedestrian, etc.) and the rules (junctions, sidewalks, stop signs, bus lanes, bicycle lanes, etc.) are different for each of them.
+Therefore the interaction between them start to be complex in modeling via macro-mathematical means.
+Three examples are discussed in more detail:
+
+\paragraph{Cars} They are the default example of the current chapter.
+Their behavior is described in subsection \ref{subsec:citysimulator} and is presented here only as a comparison member for the other examples.
+\paragraph{Pedestrians} \label{par:pedestrians} They differ drastically from the cars in few aspects.
+First is the fact that they do not share same space of a street.
+From the perspective of a city simulator this means that it needs to know the actor type and it needs to maintain a separate zone of density information.
+The second aspect is that, altough, the pedestrian are by default walking on bare foots along streets, they might take public transport or taxis to travel between their points of interests.
+This has to be reflected into their routing algorithms.
+The other part of this aspect is discussed in subsection \ref{subsec:relationsbetweenactors}.
+\paragraph{Buses} Those are the same as cars but they have a fixed travel route to move across.
+Also they might share same lanes with cars or they might have different lanes.
+Thus the city simulator needs to know this actor type too and it needs to know where on the map they have special public transport lanes, to keep appropiate density numbers for them.
+\paragraph{Bicycles} The interesting particularity of those are the fact that they can limit the speed of other actors (cars or buses) to their speed on the places where overtaking is not possible.
+Again, this is one more burden for the city simulator because it needs to know where bicycles are and it needs to know which streets are too narrow.
+
+\subsection{Relations between actors}
+\label{subsec:relationsbetweenactors}
+
+\paragraph{Driving rules} Those are probably the most common interactions between actors.
+They interact with one another, regardless of their type, through the way a city works.
+Such a relation might be modeled by a junction or roundabout.
+Or it might be a crosswalking.
+Therefore the city simulator, once again, need to take into account those details.
+Modeling a junction might not be an easy feat and most probably here lies one of the places where a discrete-computed approach is more difficult to define and implement than an analogic aproximation.
+The difference, altough stated in loosely defined terms is that the definition of density can be a function of the number of actors on a given street.
+This can be computed at any time (or, better yet, it is computed on every event which occurs.
+i.e.
+an actor reporting its state on or off a street).
+A junction, to the contrary, might be too dynamic to be easy modeled the same way.
+An aproximation of it is a coefficient based upon an actor is allowed to go through that junction and continue its journey.
+Whenever an actor reaches a junction its algorithm will have to compute its advance speed based not only on the density (or speed coefficient) number but also based on the junction coefficient.
+And, of course, the city simulator might choose to vary that very junction coefficient based on a number of factors such as how many cars are awaiting on a joined street (intuitively, as more and more cars are piling up around a junction, harder it is to pass through it).
+
+\paragraph{Contained actors}
+The paragraph \ref{subsec:actordiffbehaviors} discussed a type of actor called pedestrians.
+They might independent but also they might choose to take a bus from one point to another.
+We call this a contained actor.
+Thus its reporting will be enriched with one more detail: if is walking by itself or if its inside a bus or a subway.
+The city simulator will have to compute the densities accordingly for the street or sidewalk but also it might need to know how many pedestrians is a bus allowed to carry.
+
+\paragraph{Close-follow actors}
+The bicycle which pedals onto a bus lane might be holding up a bus behind it.
+Or a number of cars if it's on a common lane.
+Therefore the speed they have might not be the one which they have without a bike.
+
+\subsection{Exception situations}
+Various situations, altough exceptional, might have a big influence on the surround traffic.
+If a portion of a street is going under few heavy modifications then the traffic might not be allowed to flow through there.
+Or if a car is stopped on the lane with hazard lights on.
+Therefore the city simulator will have the opportunity to modify the speed coefficient.
+It might choose to make it difficult for traffic to pass through that area (if out of 4 lanes, two for each direction, only 2 are used, one for each direction) or it might choose to completely block the portion.
+It would be of great interest to see, in this case, how the traffic adapts itself to flow through other streets and how much pressure it puts on them, on a different situation than the usual one.
+The situation of a completely blocked street might also pose the idea of making it a pedestrian-only street and study the effect it would have on the car traffic.
+
 \subsection{Analyzers}
+
+The final part of this traffic simulation chapter is about the possibilities of analyzing pieces of the big picture.
+It is the supreme purpose of a traffic simulation, to be able to analyze how it works, what specific behaviors it exhibits, in order to find new means and ways of improving it.
+
+\paragraph{Hot areas}
+
+\paragraph{Statistics numbers}
+
+\paragraph{Geographical differences}
 

### 2020-01-19 thesis/report1.tex
Referral 1: added a figure and some words on analyzers.

@@
 \usepackage{listings}
+\usepackage{graphicx}
 
@@
 
+\begin{figure*}
+    \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.jpg}
+    \centering
+    \caption{Screenshot of the web interface showing the running simulation}
+    \label{fig:ScreenshotWebInterface2}
+  \end{figure*}
+
 \section{City traffic optimizations}
@@
 
+The central piece of those analyzers is the city simulator.
+It contains all the centralized data, from all its actors and it has a minimal API (Application Programming Interface) to serve this data.
+Actors are the first to consume it as they might need various pieces of information to wisely choose their track.
+A separate microservice may be the one which takes this data and analyzes it further.
+The reason for a separate microservice is the separation of concerns.
+The city simulator is responsible only for keeping together the data and serving it.
+Any other duty of transform, extrapolate, etc.
+is not its task to do.
+
+Therefore few ways of taking advantage of this data are described below. They are by no means exhaustive and represent only few places to start.
+
 \paragraph{Hot areas}
+
+Hot areas are defined as places or streets where traffic is pilling up more than in other areas. Intuitively
 

### 2020-01-19 thesis/report1.tex
Referral 1: few more words on the hot areas analyzer.

@@
 
-Hot areas are defined as places or streets where traffic is pilling up more than in other areas. Intuitively
+Hot areas are defined as places or streets where traffic is pilling up more than in other areas.
+Figure \ref{fig:ScreenshotWebInterface2} \citep{microservicestrafficsimulation} shows such a analyze using visual cues: streets are colored with red where there cars on them.
+A slight implementation of this would be to color them gradually and proportional with how many cars are on a specific street.
+Therefore the human eye will have a better understanding.
+Other places which could be analyzed are junctions.
+Figure \ref{fig:ScreenshotWebInterface2} doesn't show how much load are on junctions or roundabouts but there is an opportunity to take adjacent numbers from the city simulator and compute how much cars are able to pass a certain junction.
 

### 2020-01-19 thesis/report1.tex
Almost finished the referral number 1.

@@
 
-\section{Current state of traffic simulation}
+%\section{Current state of traffic simulation}
 
@@
 Other places which could be analyzed are junctions.
-Figure \ref{fig:ScreenshotWebInterface2} doesn't show how much load are on junctions or roundabouts but there is an opportunity to take adjacent numbers from the city simulator and compute how much cars are able to pass a certain junction.
+Figure \ref{fig:ScreenshotWebInterface2} doesn't show how much load are on junctions or roundabouts but there is an opportunity to take adjacent numbers from the city simulator and compute how much cars are able to pass a certain junction in a given time frame.
 
@@
 
+Sourcing data from the city simulator can meet other necessities.
+For example the average speed on a street interval, within a given time interval during the day.
+This could show how much the traffic is slowing down or speeding up.
+The dynamics of a street or of a neighborhood can be studied in more detail and with greater accuracy.
+The cars density can be extrapolated and linked to the emissions they have and their impact on the city atmosphere.
+The density distribution over time could lead to a better understanding of how the traffic is piling up or down during the day.
+
 \paragraph{Geographical differences}
+
+Another place where those kind of simulation can be helpful is to study would-be differences in the geography of the city.
+A junction can be replaced with a roundabout and based on the extrapolated data from the other roundabouts of the city it can be studied what impact this modification might have.
+There are other modifications which can undergo the same study treatment.
+One would be to add a new lane or, conversely, to remove a lane and make it for public transport only.
+A street can be closed to study its impact on the surrounding streets and asses whether a temporary closing is possible or not.
+Or if it's possible to make it a street only for pedestrians.
+Or to make it one way.
+Those are things which can be studied before actual modifications of them and without disturbing a highly loaded city neighborhood.
 

### 2020-01-21 papers/field trials with microservices/paper.tex
Started to work on paper for field trials with traffic splitting.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Field trials using traffic splitting between microservices}

\author{\IEEEauthorblockN{1\textsuperscript{st} Toma Becea}
    \IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
    \textit{Technical University of Cluj-Napoca}\\
            Cluj-Napoca, Romania \\
            tomabecea@pm.me}
    \and
    \IEEEauthorblockN{2\textsuperscript{nd} Honoriu Valean}
    \IEEEauthorblockA{\textit{Automation and Computer Science Faculty} \\
    \textit{Technical University of Cluj-Napoca}\\
            Cluj-Napoca, Romania \\
            Honoriu.Valean@aut.utcluj.ro}
}

\maketitle

\begin{abstract}
    When deployments of new versions of a software are needed, their introduction into production might create service disruptions for end users. This papers explores the concept of traffic splitting where a new version is released and then gradually deployed to the users. The graduality means that both versions are up at the same time and the new one will initially receive only 10\% of traffic. If the traffic is deemed to have same error rate as the old version then the splitting rules are increased. Finally, the old version will receive no traffic and can be removed from the system.
\end{abstract}

\begin{IEEEkeywords}
    software trials, microservices, traffic splitting, Kubernetes, Linkerd
\end{IEEEkeywords}

\section{Introduction}
    In today's online software solutions the primary service level agreement (SLA) is the uptime. Site reliability engineers (SREs) are striving to achieve perfection and they measure their target in number of nines, i.e. 99.999\%. Altough for ordinary users the small difference is not noticeable due to their various internet connection problems, the achievement is nevertheless an impresive feat. To achieve this, various tehniques and software solutions are put together in clever ways. This paper will explore the use of microservices for serving traffic and the deployment of new versions of the same software in a manner where users are seeing minimum disruption of the service they are using.

\subsection{Terms}
    To lay the foundation of the upcoming discussion few definitions are needed.
\paragraph{Release}
    The process of release means that a new version of a specific piece of software is running into the production system. However, this new version is not yet serving traffic to external end users.
\paragraph{Deploy}
    Once the new version is starting to serve traffic to external end users, it is said that the version has been deployed.
\paragraph{Error rate}
    A service which responds with error codes, on top of any other data it might need to send back, will likely respond from time to time with codes other than successfull ones (e.g. in HTTP error codes the number 200 means the response is succesfull while the well known number 404 means that a certain resource has not been found), from reasons which are outside its inner workings. For example the users accidentally asks for a resource which was previously deleted, case where the service itself cannot do anything but respond with an error code.

\section{Current state}
    Software deployment has become an area of innovation since CD-ROMs weren't anymore the only solution. One of the old (dated back in 1999) and good readings about this subject propose the Software Dock \cite{b1}. This idea splits the process in two areas: producer side and consumer side. Producer side processes are responsible for bridging the development and deployment realms.

    An interesting read is the analysis made by A. Dearle in \cite{b3}. He analyzes few technology names which were widely used to deploy software, despite the fact that, during the period after \cite{b3} has been published few of those technologies are not anymore the defacto deployment solutions. Java Beans, created by Sun (later it was acquired by Oracle) has 4 phases: development, deployment, service availability and undeployment. Linux is another example using Red Hat Package Manager. Altough RPM is widely in use for deploying software or packages, it is not used to deploy it in live or online environments but rather on linux distros, while in "offline" mode.

    Finally, because this paper mentions error rates, it is worth noting the work of Zhang and Pham, in \cite{b2}, who propose a software reliability growth model and adressing the difference seen between test environment and field environment. The model they propose encompasses the fact that errors and faults are not always solvable in an instant manner but they need to be deferred, usually because of the development process or because of using off-the-shelf software solutions.

\section{Solution}

\subsection{Microservices and orchestration}
    Microservices are the state of the art for designing new data and traffic intensive applications, while at the same time taking advantage of few mandatory issues revolving around today's challenges \cite{}: security is paramount and isolation and appropiate permissions for them (e.g. no 'sudo' user running code in a container) are an attractive concept. They enable architecture granularity which solves few challenges around Conway's law ("organizations which design systems ... are constrained to produce designs which are copies of the communication structures of these organizations" \cite{b4}). They advertise and mostly provide identical environments for both development and production. They package code and together with a wise package manager (e.g. Node Package Manager or NuGet) they solve the dependency hell and, sometimes, the tight coupling with the host OS (which still exists but is contained in the container as its definition, not anymore a host OS prerequisite). They also provides easier orchestration and elasticity in demand and configuration. However, as much as they are advertised as the ultimate architectural solution, there are cases when different architecture paradigms are better. One example would be monoliths (altough they are greatly blamed by the proponents of microservices) and another one could be actor model.

    % \begin{figure}
    %     \centering
    %     \includegraphics[width=0.5\textwidth]{column.png}
    %     \caption{Gas separation column}
    %     \label{fig:column}
    % \end{figure}

\subsection{Kubernetes}

\subsection{Service meshes}

\section{Implementation}

\begin{thebibliography}{00}

    \bibitem{b1} Hall, Richard S., Dennis Heimbigner, and Alexander L. Wolf. "A cooperative approach to support software deployment using the software dock." Proceedings of the 21st international conference on Software engineering. 1999.

    \bibitem{b2} Xuemei Zhang, Hoang Pham. "Software field failure rate prediction before software deployment, Journal of Systems and Software", Volume 79, Issue 3, 2006, Pages 291-300, ISSN 0164-1212.

    \bibitem{b3} A. Dearle. "Software Deployment, Past, Present and Future," Future of Software Engineering (FOSE '07), Minneapolis, MN, 2007, pp. 269-284. doi: 10.1109/FOSE.2007.20.

    \bibitem{b4} M. Conway, \href{https://en.wikipedia.org/wiki/Conway%27s_law}{Wikpedia, Conway's law page}

    \bibitem{b5} Thönes, Johannes. "Microservices." IEEE software 32.1 (2015): 116-116.
\end{thebibliography}
\vspace{12pt}

\end{document}

### 2020-01-21 papers/field trials with microservices/paper.tex
Few more words on the microservices traffic splitting.

@@
 \subsection{Microservices and orchestration}
-    Microservices are the state of the art for designing new data and traffic intensive applications, while at the same time taking advantage of few mandatory issues revolving around today's challenges \cite{}: security is paramount and isolation and appropiate permissions for them (e.g.
+    Microservices are the state of the art for designing new data and traffic intensive applications, while at the same time taking advantage of few mandatory issues revolving around today's challenges \cite{b5}: security is paramount and isolation and appropiate permissions for them (e.g.
 no 'sudo' user running code in a container) are an attractive concept.
@@
 However, as much as they are advertised as the ultimate architectural solution, there are cases when different architecture paradigms are better.
-One example would be monoliths (altough they are greatly blamed by the proponents of microservices) and another one could be actor model.
+One example would be monoliths \cite{b6} (altough they are greatly blamed by the proponents of microservices) and another one could be actor model \cite{b7}.
+
+    One can quickly start and use any microservice type approach towards a software solution.
+Any proof of concept is feasible and doable within minutes with all the open source generally available images for various products (databases, APIs, etc.).
+However upon doing a commercially viable product or a solution whose complexity is bigger, the first and uttermost problem is the handling of containers.
+This includes their lifetime, communications between them or with outer entities, persistence, updates, failure resilience, etc.
+The term "orchestration" sums up all those various operations.
+
+    The defacto solution for building and starting containers is Docker \cite{b8}.
+It also offers a simple orchestrator called Docker Compose, meant to run on a single computer.
+Its enterprise offerings contain Docker Swarm, a much more capable orchestrator.
+The current paper, however, will rely on the most ubiquitous solution of today's distributed systems world: Kubernetes \cite{b9}.
+More details on this, in subsection \ref{subsec:kube}.
 
@@
 \subsection{Kubernetes}
+\label{subsec:kube}
+
+    Kubernetes stems from Google and has grown on the experience and challenges Google faced while building and maintaining their services, after two other closed sourced were build and used: Borg and Omega \cite{b9}.
+Kubernetes is open source and is not anymore managed by Google but by a foundation they have spinned up to oversee Kubernetes (and now they have hundreds of open source solutions in their landscape), called Cloud Native Computing Foundation or CNCF \cite{b11}.
 
@@
 
-    \bibitem{b5} Thönes, Johannes. "Microservices." IEEE software 32.1 (2015): 116-116.
+    \bibitem{b5} Thones, Johannes. "Microservices." IEEE software 32.1 (2015): 116-116.
+
+    \bibitem{b6} Villamizar, Mario, et al.
+"Evaluating the monolithic and the microservice architecture pattern to deploy web applications in the cloud." 2015 10th Computing Colombian Conference (10CCC).
+IEEE, 2015.
+
+    \bibitem{b7} Sheikh, Hammad, Ángel Gómez, and Scott Atran. "Empirical evidence for the devoted actor model." Current Anthropology 57.S13 (2016): S204-S209.
+
+    \bibitem{b8} Docker. \href{https://www.docker.com}{Docker website}
+
+    \bibitem{b9} Kubernetes. \href{https://kubernetes.io}{Kubernetes website}
+
+    \bibitem{b10} Burns, Brendan, et al. "Borg, omega, and kubernetes." Queue 14.1 (2016): 70-93.
+
+    \bibitem{b11} CNCF. \href{https://www.cncf.io}{CNCF website}
 \end{thebibliography}

### 2020-01-22 papers/field trials with microservices/paper.tex
Finished paper with traffic splitting for microservices.

@@
 
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=0.5\textwidth]{column.png}
-    %     \caption{Gas separation column}
-    %     \label{fig:column}
-    % \end{figure}
-
 \subsection{Kubernetes}
@@
 
-\subsection{Service meshes}
+    Its offerings are overwhelming, both in their number and in their complexity.
+On top of them the third party offerings are intimidating at least.
+For the current paper only a handful of them will be described.
+
+    Taken in a good order for understanding, the first item to discuss is to asses the microservices concept into Kubernetes context.
+Being made for orchestrating microservices is itself a collection of various microservices.
+(The details of its fabric are out of our scope).
+Therefore whatever it is to run by Kubernetes it has to be an image which became one or more pods (or containers).
+On top of this we need to know there is a service which knows about all pods of a certain type and expose their ports to outer (within same Kubernetes cluster or outside of it) clients.
+This service is also an abstraction on top of replicated pods, where it can load balance the traffic or route only to the pods which are healthy.
+Besides services (and their underpart called deployments) there are many other concepts which are having a play: ingresses, config maps, secrets, replicas, etc.
+
+    The final part for this puzzle we need is the pod mechanics.
+A pod is the base unit in Kubernetes and it might contain one or more containers.
+In practice there is a main container which is serving some business logic functionality and all the others, "supporting" containers, are called sidecars.
+For the current's subject the single piece of interplay between a container and a sidecar, within a pod, is to know the existence of Linux namespaces, specifically of network namespaces \cite{b12}.
+This functionality allows for a sidecar to hijack the communications of a container, using iptables \cite{b13}, in a transparent manner for the container, i.e.
+the container do not know that its communications are proxied via a sidecar.
+In practice this allows for a developer or a devops person to use a self made container with the business functionality it needs, while at the same time to use a third party product for inserting proxy or proxy-based functionalities.
+
+\subsection{Service mesh}
+
+    The high reliance on communications between many microservices has quickly became a friction regarding several aspects.
+The operational cost of them is one aspect.
+Then the many layers which a request has to travel or be split into smaller requests makes it difficult to trace it and debug it.
+Also there are aspects which regard all communications therefore they might benefit of having them centralized and contained.
+Proxy utilities, for both levels, 4 and 7, are another concern which might have a stake into how a service mesh works.
+Latencies, HTTP success rates, requests per second, read/write bytes per second, and many others are metrics which need to be looked after.
+All those aspects and many others, for example see \cite{b15}, come together into an interplay which is generally called a service mesh.
+One off the shelf product of this type will be used in the implementation, section \ref{sec:implementation}.
+
+\subsection{The concept of traffic splitting}
+
+    \begin{figure}
+        \centering
+        \includegraphics[scale=0.25]{traffic100old.png}
+        \caption{Traffic goes to a single version of the service}
+        \label{fig:traffic100old}
+    \end{figure}
+
+    The departure situation can be seen in figure \ref{fig:traffic100old}.
+All clients are talking to the same version of the service.
+It is out of current scope to define the way this traffic goes into a Kubernetes cluster or its ingress, in the its parlance or what kind of traffic is but the assumption is that we have to deal with TCP or UDP traffic, on top of which, usually, there might HTTP calls or RPC ones.
+The basic assumption, though, is that the proxy which is used is a layer 4 one per Open Systems Interconnection model \cite{b14}.
+
+    \begin{figure}
+        \centering
+        \includegraphics[scale=0.25]{traffic9010.png}
+        \caption{Traffic goes to both versions of the service}
+        \label{fig:traffic9010}
+    \end{figure}
+
+    After the deployment of the new version of the service, the service mesh will start to route a part of the traffic to the new service.
+Figure \ref{fig:traffic9010} depicts this situation.
+The factors of routing are a set of initial parameters and they can be manually and progressively modified or an automated tool can be used.
+
+    \begin{figure}
+        \centering
+        \includegraphics[scale=0.25]{traffic100new.png}
+        \caption{Traffic goes to the new version of the service}
+        \label{fig:traffic100new}
+    \end{figure}
+
+    If the traffic to the new service has the same error rate as the old one and same P99 latency then the finally situation is to have the entire traffic routed only to the new service.
+This can be seen in \ref{fig:traffic100new}.
+Once the old service doesn't receive any traffic it can be removed together with all its dependencies.
 
 \section{Implementation}
+\label{sec:implementation}
+
+    For the implementation following products have been used: Kubernetes \cite{b9} for orchestrating microservices and Linkerd \cite{b16} as the service mesh.
+A dummy web server has been setup and an artificial loading pod has been put in place to call the web service and simulate load on it.
+
+    Let's look at the figure \ref{fig:canary}.
+The service used for rolling its versions is named \textit{podinfo}.
+The right rectangle, called \textit{svc/podinfo} is the service which internal clients see and can use.
+External clients may see same service but in order to do this an ingress object has to be set and it has to be instructed to route a certain path towards this service.
+This ingress become an abstraction which can hide the service name for external clients.
+Back to the figure \ref{fig:canary}, the right rectangles, called \textit{svc/podinfo-canary} and \textit{svc/podinfo-primary} are the actual two services which correspond to the two versions implied, the old one, named \textit{primary} and new one which the service is rolling to, called \textit{canary}.
+The same rectangles offers few more informations, in real time:
+
+    \begin{itemize}
+        \item Weight is the percent of traffic which is routed to the respective under-service, primary or canary.
+        \item SR stands for Success Rate and shows how many requests have succesfully finished.
+        \item RPS stands for Requests Per Second
+        \item P99 shows the latency for the 99th percentile
+    \end{itemize}
+
+    The left rectangle, \textit{svc/podinfo}, shows the SR and RPS numbers in total. As expected, the RPS number for both \textit{primary} and \textit{canary} should sum up to the total.
+
+    Finally, the bottom part of the figure \ref{fig:canary} shows similar information with the difference that it also contains latencies for 50 and 95 percentiles.
+
+    \begin{figure}
+        \centering
+        \includegraphics[scale=0.25]{canary.png}
+        \caption{Traffic split between two versions}
+        \label{fig:canary}
+    \end{figure}
+
+    Success rate and latencies are playing a major role into minimizing the errors seen by clients.
+The rolling process can choose to stop the rolling if the success rate of the canary version drops below 99\% or below the primary version.
+
+\section{Conclusions}
+
+    A generally explanation and few introductory terms have been presented.
+Also a bit of history of software has been noted through older published papers so that a comparison can be made between what was happening twenty years ago and today.
+Those have been covered on the \textit{Introduction} and on the \textit{Current state}.
+
+    Theoretical and explanatory subjects have been discussed about the world of microservices and orchestration: isolation, containers, orchestration, linux namespaces, communications, proxies and service meshes.
+Commercial examples have been given so that the concepts discussed can be further studied in the context of how they were implemented.
+
+    The \textit{Solution} section dwells more into how the concepts are choosed to be implemented and what technologies were used, while at the same time offering a complete picture of what traffic splitting is trying to achieve, i.e to minimize the errors seen by the client when new versions of software are deployed and released.
+
+    Finally, the \textit{Implementation} section shows each technology which was choosen and actual screen shots taken during a new software version rolling or releasing, using a service within a Kubernetes cluster.
 
@@
     \bibitem{b11} CNCF. \href{https://www.cncf.io}{CNCF website}
+
+    \bibitem{b12} Linux network namespaces. \href{https://www.systutorials.com/docs/linux/man/7-network_namespaces/}{Linux Man Pages}
+
+    \bibitem{b13} Linux IP Tables. \href{https://www.systutorials.com/docs/linux/man/8-iptables/}{Linux Man Pages}
+
+    \bibitem{b14} OSI Model. \href{https://en.wikipedia.org/wiki/OSI_model}{Wikipedia page for OSI Model}
+
+    \bibitem{b15} Li, Wubin, et al.
+"Service Mesh: Challenges, state of the art, and future research opportunities." 2019 IEEE International Conference on Service-Oriented System Engineering (SOSE).
+IEEE, 2019.
+
+    \bibitem{b16} Linkerd. \href{https://linkerd.io}{Linkerd website}
+
 \end{thebibliography}

### 2020-01-23 papers/field trials with microservices/paper.tex
Corrections made for traffic splitting between microservices paper.

@@
 \usepackage{hyperref}
+\usepackage{balance}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \begin{abstract}
-    When deployments of new versions of a software are needed, their introduction into production might create service disruptions for end users.
+    When deployments of new versions of a piece of software are needed, their introduction into production might create disruptions for end users.
 This papers explores the concept of traffic splitting where a new version is released and then gradually deployed to the users.
-The graduality means that both versions are up at the same time and the new one will initially receive only 10\% of traffic.
+The graduality means that both versions are up at the same time and the new one will initially receive only a small percentage (10\%) of traffic.
 If the traffic is deemed to have same error rate as the old version then the splitting rules are increased.
@@
 \begin{IEEEkeywords}
-    software trials, microservices, traffic splitting, Kubernetes, Linkerd
+    software trials, software deployments, microservices, traffic splitting, Kubernetes, Linkerd
 \end{IEEEkeywords}
@@
 Altough for ordinary users the small difference is not noticeable due to their various internet connection problems, the achievement is nevertheless an impresive feat.
-To achieve this, various tehniques and software solutions are put together in clever ways.
+To achieve this, various techniques and software solutions are put together in clever ways.
 This paper will explore the use of microservices for serving traffic and the deployment of new versions of the same software in a manner where users are seeing minimum disruption of the service they are using.
@@
 in HTTP error codes the number 200 means the response is succesfull while the well known number 404 means that a certain resource has not been found), from reasons which are outside its inner workings.
-For example the users accidentally asks for a resource which was previously deleted, case where the service itself cannot do anything but respond with an error code.
+For example the users accidentally asks for a resource which was previously deleted, this being a case in which the service itself cannot do anything but respond with an error code.
 
@@
 Dearle in \cite{b3}.
-He analyzes few technology names which were widely used to deploy software, despite the fact that, during the period after \cite{b3} has been published few of those technologies are not anymore the defacto deployment solutions.
+He analyzes few technology names which were widely used to deploy software, despite the fact that, in the period after \cite{b3} has been published, few of those technologies are not anymore the defacto deployment solutions.
 Java Beans, created by Sun (later it was acquired by Oracle) has 4 phases: development, deployment, service availability and undeployment.
@@
 
-    Finally, because this paper mentions error rates, it is worth noting the work of Zhang and Pham, in \cite{b2}, who propose a software reliability growth model and adressing the difference seen between test environment and field environment.
+    Finally, because this paper mentions error rates, it is worth noting the work of Zhang and Pham, in \cite{b2}, who propose a software reliability growth model and adress the difference seen between test environment and field environment.
 The model they propose encompasses the fact that errors and faults are not always solvable in an instant manner but they need to be deferred, usually because of the development process or because of using off-the-shelf software solutions.
@@
 no 'sudo' user running code in a container) are an attractive concept.
-They enable architecture granularity which solves few challenges around Conway's law ("organizations which design systems ...
-are constrained to produce designs which are copies of the communication structures of these organizations" \cite{b4}).
-They advertise and mostly provide identical environments for both development and production.
+They enable architecture granularity which solves few challenges around Conway's law:
+
+    \begin{quote}      
+        Organizations which design systems ... are constrained to produce designs which are copies of the communication structures of these organizations. \cite{b4}
+    \end{quote} They advertise and mostly provide identical environments for both development and production.
 They package code and together with a wise package manager (e.g.
@@
     One can quickly start and use any microservice type approach towards a software solution.
-Any proof of concept is feasible and doable within minutes with all the open source generally available images for various products (databases, APIs, etc.).
+Any proof of concept is feasible and doable within minutes with all the open source generally, available images for various products (databases, APIs, etc.).
 However upon doing a commercially viable product or a solution whose complexity is bigger, the first and uttermost problem is the handling of containers.
 This includes their lifetime, communications between them or with outer entities, persistence, updates, failure resilience, etc.
-The term "orchestration" sums up all those various operations.
+The term "orchestration" sums up all those various challenges.
 
@@
 It also offers a simple orchestrator called Docker Compose, meant to run on a single computer.
-Its enterprise offerings contain Docker Swarm, a much more capable orchestrator.
+Its enterprise offerings contains Docker Swarm, a much more capable orchestrator.
 The current paper, however, will rely on the most ubiquitous solution of today's distributed systems world: Kubernetes \cite{b9}.
@@
 On top of them the third party offerings are intimidating at least.
-For the current paper only a handful of them will be described.
+For the current paper, however, only a small number of them will be briefly touched and explained.
 
@@
 Therefore whatever it is to run by Kubernetes it has to be an image which became one or more pods (or containers).
-On top of this we need to know there is a service which knows about all pods of a certain type and expose their ports to outer (within same Kubernetes cluster or outside of it) clients.
+On top of pods, not necessarily the next level, there is a service which knows about all of them who belong to a certain type and expose their ports to outer clients (within same Kubernetes cluster or outside of it).
 This service is also an abstraction on top of replicated pods, where it can load balance the traffic or route only to the pods which are healthy.
@@
 
-    The final part for this puzzle we need is the pod mechanics.
+    The final part for this puzzle is about some of the pod mechanics.
 A pod is the base unit in Kubernetes and it might contain one or more containers.
@@
 For the current's subject the single piece of interplay between a container and a sidecar, within a pod, is to know the existence of Linux namespaces, specifically of network namespaces \cite{b12}.
-This functionality allows for a sidecar to hijack the communications of a container, using iptables \cite{b13}, in a transparent manner for the container, i.e.
+This functionality allows for a sidecar to \textit{hijack} the communications of a container, using iptables \cite{b13}, in a transparent manner for the container, i.e.
 the container do not know that its communications are proxied via a sidecar.
@@
 Then the many layers which a request has to travel or be split into smaller requests makes it difficult to trace it and debug it.
-Also there are aspects which regard all communications therefore they might benefit of having them centralized and contained.
+Also there are metrics and information which regard all communications therefore they might benefit of having them centralized and contained.
 Proxy utilities, for both levels, 4 and 7, are another concern which might have a stake into how a service mesh works.
@@
 All clients are talking to the same version of the service.
-It is out of current scope to define the way this traffic goes into a Kubernetes cluster or its ingress, in the its parlance or what kind of traffic is but the assumption is that we have to deal with TCP or UDP traffic, on top of which, usually, there might HTTP calls or RPC ones.
+It is out of current scope to define the way this traffic goes into a Kubernetes cluster via its ingress, or what kind of traffic is but the assumption is that we have to deal with TCP or UDP traffic, on top of which, usually, there might HTTP calls or RPC ones.
 The basic assumption, though, is that the proxy which is used is a layer 4 one per Open Systems Interconnection model \cite{b14}.
@@
     If the traffic to the new service has the same error rate as the old one and same P99 latency then the finally situation is to have the entire traffic routed only to the new service.
-This can be seen in \ref{fig:traffic100new}.
-Once the old service doesn't receive any traffic it can be removed together with all its dependencies.
+This can be seen in figure \ref{fig:traffic100new}.
+Once the old service doesn't receive any traffic it will be removed together with all its dependencies.
 
@@
 
-    A generally explanation and few introductory terms have been presented.
-Also a bit of history of software has been noted through older published papers so that a comparison can be made between what was happening twenty years ago and today.
-Those have been covered on the \textit{Introduction} and on the \textit{Current state}.
+    A generally explanation and few introductory terms have been presented for laying the grounds of the subject.
+Also a bit of history of software deployments has been noted through older published papers so that a comparison can be made between what was happening twenty years ago and today.
+Those have been covered on the sections \textit{Introduction} and \textit{Current state}.
 
@@
 
-    Finally, the \textit{Implementation} section shows each technology which was choosen and actual screen shots taken during a new software version rolling or releasing, using a service within a Kubernetes cluster.
+    Finally, the \textit{Implementation} section shows each technology which was choosen and actual screen shots taken during a new software version rolling or releasing, using an example service within a Kubernetes cluster.
+
+\balance
 
@@
 \end{thebibliography}
-\vspace{12pt}
 

### 2020-02-18 papers/traffic simulation/paper.tex
Switched png fiels to jpg for a smaller size of pdf.
Traffic simulation paper: incorporated feedback and switched bibliography style.

@@
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{url,apacite}
+\usepackage{url}
 \usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
 \usepackage{graphicx}
@@
 \usepackage{xcolor}
-\usepackage{natbib}
+%\usepackage{natbib}
 \graphicspath{ {.} }
-\bibliographystyle{apacite}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \usepackage{xcolor}
+
+\usepackage{balance}
 
@@
 
-Simulating traffic in a city is a difficult task.
-Current paper is proposing a novel way using the state of the art in today's distributed systems: microservices orchestration.
-The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation.
-The other actor type (multiple instances) is the car which travels across the city.
+The paper proposes a city traffic software simulation based on actors which run independently of one another and have specific characters in their behavior.
+To run indepentently actors are modeled as microservices and they are running within an orchestration framework.
+Their specific behavior is modeled as simple algorithms, embedded in each actor's type code.
+They may act based on the data about all the other actors, data which is gathered together by a single entity called city simulator.
+An orchestration model is proposed and all the actors have a communication protocol to offer data to the city simulator and request data from it.
 
 \end{abstract}
+
+\begin{IEEEkeywords}
+  traffic simulation; microservices; distributed computing
+\end{IEEEkeywords}
 
@@
 
+The solution proposed is aiming to design a traffic simulation software which surpasses the computing limits of a single machine and of existing traffic simulation software and runs as a distributed system.
+In close interplay with the fundamental nature of distributed system (i.e.
+running a cohesive software on multiple machines) the solution defines entities or actors, modeled as separate microservices, such that the advantage is twofold.
+First, an entity being a single microservice (and a microservice containing only one entity) they will be easier and natural scheduled and run across more than one machine.
+Second, the independence of such an entity is also helping in personalize its behavior in randomness and character, offering different ways to model the traffic in a city and allowing more nuanced studies as opposed to macroscopic solutions.
+
 We define a \textbf{(city) actor} as being an independent entity which chose to move between two geographical points within a city.
-It can be a car, a pedestrian or a bike.
+As a character, it can be a car, a pedestrian or a bike.
 We also define the \textbf{city simulator} as being a single entity (subject to distributed and load balacing services) which keeps data about the city (e.g.
@@
 
-This paper is structured in few sections.
-Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulation.
-Section \ref{sec:implementation} discuss the core approach of the current paper in architecting a software solution based on microservices for simulating traffic in a city.
-Section \ref{sec:orchestration} discuss the architectural aspect of orchestrating microservices using popular frameworks and solutions.
-Section \ref{sec:results} presents the results obtained after the implementation of some of the core concepts from the previous sections.
-
-\section{Existing solutions}
+\section{Related work}
 \label{sec:existingsolutions}
 
-Existing solutions have a general distinction of being on a spectrum: macroscopic, mesoscopic or microscopic.
+Existing solutions have a general distinction of being scattered across a spectrum: macroscopic, mesoscopic or microscopic being milestones across it.
 This distinction is caused by how the traffic is modeled: on a macroscopic scale like streets or sections of a highway or on microscopic level, focusing on each car and its relation with neighbours.
@@
 
-\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with its own personality and decisions.
-This idea is an interesting one especially because it is proposed as an enhancement to the solution presented by the current paper.
-In the same area of microscopic simulation we can see another proposal from \cite{7004985}.
+(to rewrite) In the same area of microscopic simulation we can see another proposal from \cite{7004985}.
 Of a great interest is their ability to feed back in the simulation the data gathered from loop detectors across the portion of highway they are particularly interested in.
-
-\cite{6205472} is probably the most viable approach as it is a complete simulation software.
-Based on the Ngel-Schreckenberg \citep{nagel-schreckenberg} automaton model of simulating a freeway traffic, P.
-Gora is advancing this idea into a fully fledged software simulation framework.
 
@@
 The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.
-
-Finally, \cite{6244609} propose optimization computation for intersections (junctions) by using a microscopic approach towards data analysis.
-The proposal use a discrete and stochastic time step of simulation model, combined with Synchro, a signal timing optimization software.
 
@@
 Microservices are not a new concept.
-The idea behind them has existed since Linux kernel has started to be enriched with a concept called namespaces \citep{wiki:linuxns}.
+The idea behind them has existed since Linux kernel has started to be enriched with a concept called namespaces \cite{wiki:linuxns}.
 This allows one set of processes to see one set of resources while another set of processes see another set of resources, where resources might be, but not limited to, process IDs, file names and network resources.
@@
 
-To go to solution for building, manipulating and running images is Docker \citep{docker}.
+To go to solution for building, manipulating and running images is Docker \cite{docker}.
 It allows easy software installation, it works cross platform and it offers a smooth experience most of the time.
@@
 
-Figure \ref{fig:citysimrelations} depicts the city simulator in relation with the other entities.
+Fig.
+\ref{fig:citysimrelations} depicts the city simulator in relation with the other entities.
 A notable exception is the web page.
@@
 The implementation proposed here is aiming to model a single type of actor: a car which moves between two points across a city.
-Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \citep{openstreetmap} and GraphHopper \citep{graphhopper}.
+Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \cite{openstreetmap} and GraphHopper \cite{graphhopper}.
 Open Street Map is an open source licensed map of the entire world.
@@
 A first choice is the programming language to write both the city simulator and the city actor.
-Go programming langauge is born out of Google and it resembles the philosophy they were trying to embedded in it for taming the complexity of their systems \citep{donovan2015go}.
+Go programming langauge is born out of Google and it resembles the philosophy they were trying to embedded in it for taming the complexity of their systems \cite{donovan2015go}.
 Today it has gained a lot of popularity and tools like Kubernetes are wrritten entirely in Go, making it the default language of the cloud technologies.
@@
 
-\begin{figure*}
-  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface.png}
-  \centering
-  \caption{Screenshot of the web interface showing the running simulation}
-  \label{fig:ScreenshotWebInterface}
-\end{figure*}
-
-\begin{figure*}
-  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.png}
-  \centering
-  \caption{Screenshot of the web interface showing the running simulation with a larger number of actors}
-  \label{fig:ScreenshotWebInterface2}
-\end{figure*}
-
 \section{Orchestration}
@@
 
-As noted above the standard in easiness of developing and working with containers is Docker \citep{docker}.
-While it has an offering of orchestrating containers, called \textit{docker-compose}, which is simple to start with, its functionality is limited in comparison with other offerings, the most notable one being Kubernetes \citep{Kubernetes}.
+As noted above the standard in easiness of developing and working with containers is Docker \cite{docker}.
+While it has an offering of orchestrating containers, called \textit{docker-compose}, which is simple to start with, its functionality is limited in comparison with other offerings, the most notable one being Kubernetes \cite{Kubernetes}.
 Those are not the single tools available and many more can be found but they offer a starting point (especially Docker) and Kubernetes, altough it has a steep learning curve, it does offer a comprehensive and complex panoply of details around orchestration.
@@
 Here comes Kubernetes in play.
-Kubernetes \citep{Kubernetes} is an open-source system for automating deployment, scaling and management of containerized applications.
+Kubernetes \cite{Kubernetes} is an open-source system for automating deployment, scaling and management of containerized applications.
 It originates from Google and is their third approach on orchestrating microservices, as described in \cite{burns2016borg}.
@@
 It is by default enriched with certain abilities needed to run in this configuration.
-Made initially to run with LXC and Docker containers \citep{7036275}, it is now a more open system where one can choose the container runtime interface, the container network interface, storage interface, service meshes, etc.
+Made initially to run with LXC and Docker containers \cite{7036275}, it is now a more open system where one can choose the container runtime interface, the container network interface, storage interface, service meshes, etc.
 from a broad range of vendors.
@@
 Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtained IP address.
-The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \citep{risley2001domain}.
+The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \cite{risley2001domain}.
 
@@
 To do this the client needs to know only the service name and a certain "networking" entity should route its request to the appropiate available service.
-Such a entity could be a DNS authoritative server which, based on the load, availability and latency of the services will route the request to the appropiate node \citep{swildens2006scalable}.
+Such a entity could be a DNS authoritative server which, based on the load, availability and latency of the services will route the request to the appropiate node \cite{swildens2006scalable}.
 
@@
 
-\section{Results}
+\begin{figure*}
+  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.jpg}
+  \centering
+  \caption{Screenshot of the web interface showing the running simulation with a larger number of actors}
+  \label{fig:ScreenshotWebInterface2}
+\end{figure*}
+
+\section{Results and evaluation}
 \label{sec:results}
 
-Figure \ref{fig:ScreenshotWebInterface} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a simulation with less than 20 actors.
+Fig.
+\ref{fig:ScreenshotWebInterface2} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a running simulation.
 The web interface is a simple: a html/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}.
@@
 
-The map is sourced from Open Street Map \citep{openstreetmap} and is showing the map zoomed to a specific city, same used by the city actors in their walkings.
+The map is sourced from Open Street Map \cite{openstreetmap} and is showing the map zoomed to a specific city, same used by the city actors in their walkings.
 Thus, whenever a city actor reports that it is traveling across a line and this information arrives at the web interface through the city simulator, the respective line will be immediately drawn into the view.
@@
 For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}).
-Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time throughout the city.
+Fig \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time throughout the city.
 
-\section{Enhancements}
+\balance
+
+\section{Conclusions and future work}
 \label{sec:enhancements}
 
-As with all technicl projects, there are a number of enhancements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
+As with all technical projects, there are a number of enhancements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
 This section will briefly propose few of those details in the following paragraphs.
@@
 To take advantage of the Kubernetes cluster there are similar horizontally scalable databases which can be built upon.
-One option is the distribute key-value store used by Kubernetes itself: Etcd datbase \citep{etcd}.
+One option is the distribute key-value store used by Kubernetes itself: Etcd datbase \cite{etcd}.
 
@@
 
-\bibliography{wiki}
+\bibliographystyle{ieeetran}
+\bibliography{references}
+
 \vspace{12pt}

### 2020-02-19 papers/traffic simulation/paper.tex
Traffic simulation paper: more feedback from reviewers incorporated.

@@
 Second, the independence of such an entity is also helping in personalize its behavior in randomness and character, offering different ways to model the traffic in a city and allowing more nuanced studies as opposed to macroscopic solutions.
+This solution can improve a system where only the global state is computed.
 
@@
 
-\section{Related work}
+\section{Related works}
 \label{sec:existingsolutions}
@@
 
-(to rewrite) In the same area of microscopic simulation we can see another proposal from \cite{7004985}.
-Of a great interest is their ability to feed back in the simulation the data gathered from loop detectors across the portion of highway they are particularly interested in.
+\cite{abm-for-traffic-routing} describes an Agent Based Model (ABM) for improved traffic routing and achieve a system-optimal traffic flow.
+The similarity with the current paper is that the data is generating at the agent level and the decision remains at the same level, altough there are two layers in total.
+First is the microscopic layer which consists of all the intelligent agents.
+The second layer is the macroscopic layer which facilitates the communications between agents.
+However, the main difference is that in \cite{abm-for-traffic-routing} an agent is able communicate only with the surrounding ones through the use of cellular automata approach.
+Agents are communicating a handful of data with their co-participants within a certain range: position, velocity, route and type.
+Using the data they received they are using a transition decision-making model, based on cellular automata, to compute their next move.
 
 In \cite{8569576} the broad idea is to use a discrete event architecture, in which there are logical processes for executing and simulating a number of agents.
-For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them.
+For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them, but the paper avoids presenting the necessary details on how it would work.
 The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.
+
+An approach which is focused on junction modeling is \cite{urban-traffic-sim}.
+As opposed to the generality of various other related works, \cite{urban-traffic-sim} is focused on two specific and real junctions.
+The simulation stems from SUMO, using real data acquired from cameras placed in junctions (and analyzed with image processing technologies) and is relaying traffic data to a Matlab instance.
+The simulation is aiming at evaluating metrics of the traffic which flows through junctions: arrival flow and queue length.
+This idea is closely related to an enhancement proposed in section \ref{sec:enhancements}.
+
+Altough older, \cite{7004985} deserves to be written about because the real data, from the highway portions of interest, is fed back into the simulation.
+Loop detectors, scattered across the highways are read and their readings are used to generate simulation models.
+The model is able to predict car densities across multiple road sections, an ability which the current paper is exploring and is basing its car actor's algorithms onto.
 
@@
 
+\begin{figure*}
+  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.jpg}
+  \centering
+  \caption{Screenshot of the web interface showing the running simulation with a larger number of actors}
+  \label{fig:ScreenshotWebInterface2}
+\end{figure*}
+
 \subsection{Networking design}
@@
 
-\begin{figure*}
-  \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.jpg}
-  \centering
-  \caption{Screenshot of the web interface showing the running simulation with a larger number of actors}
-  \label{fig:ScreenshotWebInterface2}
-\end{figure*}
-
 \section{Results and evaluation}
@@
 Above a certain threshold the bottleneck, surprisingly, is not the memory or the cpu but the way docker handles network interfaces.
-For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}).
+For a larger number of city actors a better tool must be used (see paragraph \ref{subsec:kubernetes}).
 Fig \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time throughout the city.
-
-\balance
 
@@
 
+The concept of modeling or simulating traffic using actors or agents is widespread.
+The advantages it offers, compared to macroscopic simulation techniques, is that more nuanced data can be obtain, usually by simulating real events and situations more easier when the agents can interact and influence each other.
+On top of the actor (or agent, or entity) based model, the current paper, compared with the other noted beforehand, has shown the benefits of using today's concepts of distributed systems: containers and orchestration.
+Those ideas (or concepts or programming frameworks), not novel in their basic traits but novel in the ease of use, enable the scaling of simulation needs across many more physical machines.
+In parallel with this the code being written do not suffer from leaky abstractions of the distributed nature it runs in.
+It is kept slim and focused, while the frameworks used are taking care of all the other networking details.
+
+Microservices concept has been presented and together with it the city simulator and the city actor types have been introduced.
+Their interactions, through a specific protocol, have been presented.
+Moreover, various other interactions, with supporting roles (routing engine and web page) have been presented: the routing engine can be called by any actor via a REST based API and the web page keeps an open WebSocket connection towards the city simulator to be notified of any change.
+Design choices, lead by the programming language (Go language) and container tool (Docker) have been discussed, together with their advantages.
+Networking details and design have been detailed, mostly lead by how the Docker and linux containers landscape works.
+Kubernetes, as the de facto solution for microservices orchestration was also presented.
+Finally, a brief description provided the results of a simulation run on a single machine.
+
+\balance
+
 As with all technical projects, there are a number of enhancements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper.
@@
 
-As the number of the actors are growing a first bottlneck in the current architecture will be the city simulator.
+As the number of the actors are growing, a first bottlneck in the current architecture will be the city simulator.
 It is a single entity and it represents a single source of truth for the actors.
@@
 To take advantage of the Kubernetes cluster there are similar horizontally scalable databases which can be built upon.
-One option is the distribute key-value store used by Kubernetes itself: Etcd datbase \cite{etcd}.
+One option is the distributed key-value store used by Kubernetes itself: Etcd database \cite{etcd}.
 
@@
 Let's take junctions.
-The city simulator can model any junction and then it can permit actors to pass or not through it.
+The city simulator can model any junction and then it can permit actors to pass or not through it, while at the same time incorporating real data as shown in \cite{urban-traffic-sim}.
 Then there are pedestrian crossings where different types of actors can meet each other.

### 2020-02-20 papers/traffic simulation/paper.tex
Fixed paper.tex

@@
 To run indepentently actors are modeled as microservices and they are running within an orchestration framework.
-Their specific behavior is modeled as simple algorithms, embedded in each actor's type code.
+Their behavior is modeled as with specific algorithms for each of their type, embedded in each actor's type code.
 They may act based on the data about all the other actors, data which is gathered together by a single entity called city simulator.
-An orchestration model is proposed and all the actors have a communication protocol to offer data to the city simulator and request data from it.
+An orchestration model is proposed and all the actors use a communication protocol to offer data to the city simulator and request data from it.
 
@@
 In \cite{8569576} the broad idea is to use a discrete event architecture, in which there are logical processes for executing and simulating a number of agents.
-For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them, but the paper avoids presenting the necessary details on how it would work.
+For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them, but the paper avoids presenting the necessary details on how the distributed mechanics would work.
 The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.
@@
 The simulation is aiming at evaluating metrics of the traffic which flows through junctions: arrival flow and queue length.
-This idea is closely related to an enhancement proposed in section \ref{sec:enhancements}.
+This idea is closely related to an enhancement proposed in section \ref{sec:enhancements} around junction modeling.
 
@@
 Design choices, lead by the programming language (Go language) and container tool (Docker) have been discussed, together with their advantages.
-Networking details and design have been detailed, mostly lead by how the Docker and linux containers landscape works.
+Networking details and design have been detailed, mostly lead by how the Docker and Linux containers landscape work.
 Kubernetes, as the de facto solution for microservices orchestration was also presented.

### 2020-10-01 papers/smart contracts for traffic data/paper.tex
Initial check in for the next paper, smart contracts for traffic simulation.

--- 2020-01-21 papers/field trials with microservices/paper.tex
@@
 
-\title{Field trials using traffic splitting between microservices}
+\title{Smart contracts for handling city traffic information}
 
@@
 \section{Introduction}
-    In today's online software solutions the primary service level agreement (SLA) is the uptime.
-Site reliability engineers (SREs) are striving to achieve perfection and they measure their target in number of nines, i.e.
-99.999\%.
-Altough for ordinary users the small difference is not noticeable due to their various internet connection problems, the achievement is nevertheless an impresive feat.
-To achieve this, various tehniques and software solutions are put together in clever ways.
-This paper will explore the use of microservices for serving traffic and the deployment of new versions of the same software in a manner where users are seeing minimum disruption of the service they are using.
 
 \subsection{Terms}
-    To lay the foundation of the upcoming discussion few definitions are needed.
 \paragraph{Release}
-    The process of release means that a new version of a specific piece of software is running into the production system. However, this new version is not yet serving traffic to external end users.
 \paragraph{Deploy}
-    Once the new version is starting to serve traffic to external end users, it is said that the version has been deployed.
 \paragraph{Error rate}
-    A service which responds with error codes, on top of any other data it might need to send back, will likely respond from time to time with codes other than successfull ones (e.g.
-in HTTP error codes the number 200 means the response is succesfull while the well known number 404 means that a certain resource has not been found), from reasons which are outside its inner workings.
-For example the users accidentally asks for a resource which was previously deleted, case where the service itself cannot do anything but respond with an error code.
 
 \section{Current state}
-    Software deployment has become an area of innovation since CD-ROMs weren't anymore the only solution.
-One of the old (dated back in 1999) and good readings about this subject propose the Software Dock \cite{b1}.
-This idea splits the process in two areas: producer side and consumer side.
-Producer side processes are responsible for bridging the development and deployment realms.
-
-    An interesting read is the analysis made by A.
-Dearle in \cite{b3}.
-He analyzes few technology names which were widely used to deploy software, despite the fact that, during the period after \cite{b3} has been published few of those technologies are not anymore the defacto deployment solutions.
-Java Beans, created by Sun (later it was acquired by Oracle) has 4 phases: development, deployment, service availability and undeployment.
-Linux is another example using Red Hat Package Manager.
-Altough RPM is widely in use for deploying software or packages, it is not used to deploy it in live or online environments but rather on linux distros, while in "offline" mode.
-
-    Finally, because this paper mentions error rates, it is worth noting the work of Zhang and Pham, in \cite{b2}, who propose a software reliability growth model and adressing the difference seen between test environment and field environment.
-The model they propose encompasses the fact that errors and faults are not always solvable in an instant manner but they need to be deferred, usually because of the development process or because of using off-the-shelf software solutions.
 
@@
 
-\subsection{Microservices and orchestration}
-    Microservices are the state of the art for designing new data and traffic intensive applications, while at the same time taking advantage of few mandatory issues revolving around today's challenges \cite{b5}: security is paramount and isolation and appropiate permissions for them (e.g.
-no 'sudo' user running code in a container) are an attractive concept.
-They enable architecture granularity which solves few challenges around Conway's law ("organizations which design systems ...
-are constrained to produce designs which are copies of the communication structures of these organizations" \cite{b4}).
-They advertise and mostly provide identical environments for both development and production.
-They package code and together with a wise package manager (e.g.
-Node Package Manager or NuGet) they solve the dependency hell and, sometimes, the tight coupling with the host OS (which still exists but is contained in the container as its definition, not anymore a host OS prerequisite).
-They also provides easier orchestration and elasticity in demand and configuration.
-However, as much as they are advertised as the ultimate architectural solution, there are cases when different architecture paradigms are better.
-One example would be monoliths \cite{b6} (altough they are greatly blamed by the proponents of microservices) and another one could be actor model \cite{b7}.
-
-    One can quickly start and use any microservice type approach towards a software solution.
-Any proof of concept is feasible and doable within minutes with all the open source generally available images for various products (databases, APIs, etc.).
-However upon doing a commercially viable product or a solution whose complexity is bigger, the first and uttermost problem is the handling of containers.
-This includes their lifetime, communications between them or with outer entities, persistence, updates, failure resilience, etc.
-The term "orchestration" sums up all those various operations.
-
-    The defacto solution for building and starting containers is Docker \cite{b8}.
-It also offers a simple orchestrator called Docker Compose, meant to run on a single computer.
-Its enterprise offerings contain Docker Swarm, a much more capable orchestrator.
-The current paper, however, will rely on the most ubiquitous solution of today's distributed systems world: Kubernetes \cite{b9}.
-More details on this, in subsection \ref{subsec:kube}.
-
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=0.5\textwidth]{column.png}
-    %     \caption{Gas separation column}
-    %     \label{fig:column}
-    % \end{figure}
-
-\subsection{Kubernetes}
-\label{subsec:kube}
-
-    Kubernetes stems from Google and has grown on the experience and challenges Google faced while building and maintaining their services, after two other closed sourced were build and used: Borg and Omega \cite{b9}.
-Kubernetes is open source and is not anymore managed by Google but by a foundation they have spinned up to oversee Kubernetes (and now they have hundreds of open source solutions in their landscape), called Cloud Native Computing Foundation or CNCF \cite{b11}.
-
-\subsection{Service meshes}
-
-\section{Implementation}
+\section{Conclusions}
 
@@
     \bibitem{b11} CNCF. \href{https://www.cncf.io}{CNCF website}
+
+    \bibitem{b12} Linux network namespaces. \href{https://www.systutorials.com/docs/linux/man/7-network_namespaces/}{Linux Man Pages}
+
+    \bibitem{b13} Linux IP Tables. \href{https://www.systutorials.com/docs/linux/man/8-iptables/}{Linux Man Pages}
+
+    \bibitem{b14} OSI Model. \href{https://en.wikipedia.org/wiki/OSI_model}{Wikipedia page for OSI Model}
+
+    \bibitem{b15} Li, Wubin, et al.
+"Service Mesh: Challenges, state of the art, and future research opportunities." 2019 IEEE International Conference on Service-Oriented System Engineering (SOSE).
+IEEE, 2019.
+
+    \bibitem{b16} Linkerd. \href{https://linkerd.io}{Linkerd website}
+
 \end{thebibliography}
-\vspace{12pt}
 

### 2020-10-01 papers/smart contracts for traffic data/paper.tex
Abstract written.

@@
 \begin{abstract}
-    When deployments of new versions of a piece of software are needed, their introduction into production might create disruptions for end users.
-This papers explores the concept of traffic splitting where a new version is released and then gradually deployed to the users.
-The graduality means that both versions are up at the same time and the new one will initially receive only a small percentage (10\%) of traffic.
-If the traffic is deemed to have same error rate as the old version then the splitting rules are increased.
-Finally, the old version will receive no traffic and can be removed from the system.
+
+    Blockchain technologies are becoming increasingly complex and thus are tackling more and more various use cases than decentralised finance which was the sole purpose of Bitcoin, first successful blockchain.
+Or they are tackling more and more use cases and thus their inherent complexity increases.
+Either way, one of the untouched places was IoT realm.
+Due to natural IoT constraints, like limited processing power and high volume of data blockchain technologies have stayed separately.
+However, the Ethereum 2.0 and IOTA are two distributed ledger technologies with two different approaches of handling a high volume of transactions without requiring specialized hardware.
+The current paper is aiming to explore the possibilites of those two technologies through the lens of a concrete use case: transfering city traffic information between traffic participants and reaching consensus towards optimizing the traffic.
+The data to be transferred has specific needs and formats.
+The traffic optimization consensus has to take into account various metrics and goals.
+Those two aspects need to be modeled on the exact blockchain technology via smart contracts.
+
 \end{abstract}
@@
 \begin{IEEEkeywords}
-    software trials, software deployments, microservices, traffic splitting, Kubernetes, Linkerd
+
 \end{IEEEkeywords}
@@
 
-    \bibitem{b2} Xuemei Zhang, Hoang Pham.
-"Software field failure rate prediction before software deployment, Journal of Systems and Software", Volume 79, Issue 3, 2006, Pages 291-300, ISSN 0164-1212.
-
-    \bibitem{b3} A. Dearle. "Software Deployment, Past, Present and Future," Future of Software Engineering (FOSE '07), Minneapolis, MN, 2007, pp. 269-284. doi: 10.1109/FOSE.2007.20.
-
-    \bibitem{b4} M. Conway, \href{https://en.wikipedia.org/wiki/Conway%27s_law}{Wikpedia, Conway's law page}
-
-    \bibitem{b5} Thones, Johannes. "Microservices." IEEE software 32.1 (2015): 116-116.
-
-    \bibitem{b6} Villamizar, Mario, et al.
-"Evaluating the monolithic and the microservice architecture pattern to deploy web applications in the cloud." 2015 10th Computing Colombian Conference (10CCC).
-IEEE, 2015.
-
-    \bibitem{b7} Sheikh, Hammad, Ángel Gómez, and Scott Atran. "Empirical evidence for the devoted actor model." Current Anthropology 57.S13 (2016): S204-S209.
-
-    \bibitem{b8} Docker. \href{https://www.docker.com}{Docker website}
-
-    \bibitem{b9} Kubernetes. \href{https://kubernetes.io}{Kubernetes website}
-
-    \bibitem{b10} Burns, Brendan, et al. "Borg, omega, and kubernetes." Queue 14.1 (2016): 70-93.
-
-    \bibitem{b11} CNCF. \href{https://www.cncf.io}{CNCF website}
-
-    \bibitem{b12} Linux network namespaces. \href{https://www.systutorials.com/docs/linux/man/7-network_namespaces/}{Linux Man Pages}
-
-    \bibitem{b13} Linux IP Tables. \href{https://www.systutorials.com/docs/linux/man/8-iptables/}{Linux Man Pages}
-
-    \bibitem{b14} OSI Model. \href{https://en.wikipedia.org/wiki/OSI_model}{Wikipedia page for OSI Model}
-
-    \bibitem{b15} Li, Wubin, et al.
-"Service Mesh: Challenges, state of the art, and future research opportunities." 2019 IEEE International Conference on Service-Oriented System Engineering (SOSE).
-IEEE, 2019.
-
-    \bibitem{b16} Linkerd. \href{https://linkerd.io}{Linkerd website}
-
 \end{thebibliography}

### 2020-10-04 papers/smart contracts for traffic data/paper.tex
Some progress on the smart contracts traffic data.

@@
 \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+
 \usepackage{cite}
@@
 \usepackage{textcomp}
-\usepackage{xcolor}
 \usepackage{hyperref}
 \usepackage{balance}
+\usepackage{listings}
+\usepackage{xcolor}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    keepspaces=true,
+    numbers=left,
+    numbersep=5pt,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    tabsize=2
+}
+
+\lstset{style=mystyle}
+
 \begin{document}
@@
 
-    Blockchain technologies are becoming increasingly complex and thus are tackling more and more various use cases than decentralised finance which was the sole purpose of Bitcoin, first successful blockchain.
-Or they are tackling more and more use cases and thus their inherent complexity increases.
-Either way, one of the untouched places was IoT realm.
-Due to natural IoT constraints, like limited processing power and high volume of data blockchain technologies have stayed separately.
-However, the Ethereum 2.0 and IOTA are two distributed ledger technologies with two different approaches of handling a high volume of transactions without requiring specialized hardware.
-The current paper is aiming to explore the possibilites of those two technologies through the lens of a concrete use case: transfering city traffic information between traffic participants and reaching consensus towards optimizing the traffic.
-The data to be transferred has specific needs and formats.
-The traffic optimization consensus has to take into account various metrics and goals.
-Those two aspects need to be modeled on the exact blockchain technology via smart contracts.
+    Blockchain technologies have been initially useful for decentralised finance transactions.
+Due to limitations in scaling, both for number of transcations and magnitude of stored data, other use cases have remained outside of their reach.
+However, recent developments of Ethereum 2.0 and IoT focused ledger technology IOTA are tackling those two limitations.
+We are analyzing the suitability of their smart contracts platform for the concrete use case of a decentralized system where participants within a city traffic are sharing traffic data among themselves and are reaching consensus for optimizing the traffic.
+The particular criteria for this analysis is the language support for structures of grouped variables,
 
@@
 
+    blockhain, ethereum, iota
+
 \end{IEEEkeywords}
@@
 
-\subsection{Terms}
-\paragraph{Release}
+    Blockchain technologies are becoming increasingly complex and thus are tackling more and more various use cases than decentralized finance which was the sole purpose of Bitcoin, first successful blockchain.
+Or they are tackling more and more use cases and thus their inherent complexity increases.
+Either way, one of the difficult for them places was IoT realm.
+Due to ineherent IoT constraints, like limited processing power and high volume of data, blockchain technologies have stayed separately.
+However, the Ethereum 2.0 and IOTA are two distributed ledger technologies with two different approaches of handling a high volume of transactions without requiring specialized hardware like GPU or ASIC.
+The current paper is aiming to explore the possibilites of those two technologies through the lens of a concrete use case: transfering city traffic information between traffic participants and reaching consensus towards optimizing the traffic.
+The data to be transferred has specific needs and formats.
+The traffic optimization consensus has to take into account various metrics and goals.
+Those two aspects need to be modeled on the blockchain technology via smart contracts.
+
+\subsection{Criteria}
+\paragraph{Data format}
+
+Following the 
+\begin{lstlisting}[caption=Go enumerations for messaging, label=lst:goactorreports]
+    const (   
+        // ReportOnTheLine is the report sent by one agent to
+        // notify the city that he is currently advancing
+        // through one line.
+        ReportOnTheLine = iota
+
+        // ReportOffFromLine is the report sent by one agent to
+        // notify the city that he has finished advancing through
+        // one line and has departed from it.
+        ReportOffFromLine = iota
+    )
+
+    const (
+        // SendReport is a message passed from an actor to the city
+        // indicating its status (e.g. location).
+        SendReport = iota
+
+        // AskForLine is a message passed from an actor to the city.
+        // A response is awaited.
+        AskForLine = iota
+
+        // RespondWithLine is a message passed from the city to
+        // an actor and it contains line data.
+        RespondWithLine = iota
+    )
+    \end{lstlisting}
 \paragraph{Deploy}
@@
 
-    \bibitem{b1} Hall, Richard S., Dennis Heimbigner, and Alexander L.
-Wolf.
-"A cooperative approach to support software deployment using the software dock." Proceedings of the 21st international conference on Software engineering.
-1999.
+    \bibitem{b1} Toma Becea and Honoriu Valean.
+“A Microservices based Approach for City Traffic Simulation”.
+\\ International Journal of Advanced Computer Science and Applications(IJACSA), 11(2), 2020.
+\\ http://dx.doi.org/10.14569/IJACSA.2020.0110209
 

### 2020-10-11 thesis/report2.tex
Initial checkin for report number 2.

--- 2020-01-19 thesis/report1.tex
@@
 
-\title{Analyzing opportunities arround city traffic optimization }
+\title{ City traffic, IoT and Blockchain technologies }
 \author{Toma Becea}
@@
 
-The traffic in a city is a very complex phenomenon.
-The parties involved are of many types: cars, pedestrians, buses, trams, taxi cabs (they are arugably cars, too but there might be small differences, e.g.
-they are allowed to drive on public transport lanes), bicycles, e-scooters (again, they are arguably same as bicycles), etc.
-Then there are rules on how they move: traffic lights for cars in junctions, across roads for pedestrian crossings, different traffic lights for buses and trams, etc.
-Within this rule-bounded context is the free will of every intelligent decision maker (it might be a human or it might be a self driving car with algorithms which we can deem as being intelligent from the perspective of behaving and moving within the traffic), the individual randomness unified into a collective orderness.
-
-But the devil is in the details: the orderness we see when looking across the entire city uncovers many opportunities.
-The architecture of many old cities is not meant for current traffic needs.
-The behavior of residents and dwellers of a city is not optimized for collective good but for personal comfort.
-The heterogenous and continous development of a city might add pain to the picture.
-Highly-accelerated developing areas or neighborhoods pose problems which city administrations are slow to understand and tackle with.
-Those, so far, are problems on the macro scale of a spectrum.
-On the other side of the spectrum, called micro, we can find another set of opportunities which can be modeled and analyzed through computer based simulations.
-Junction design, traffic lights programming and synchronization, pedestrian crossings and their influence on the surrounding driving lanes, bus stops and their influence, side walks design, lanes formation and one way drives, etc.
-Those are localized problems, focused on small geographical areas.
-
-The current chapter aim to find and define technical means on exploring intricacies within the traffic of a city.
-Software simulation is a broad term which encompasses any idea or way of using a computer to mimic as closely as possible a real phenomenon, with or without the help of a mathematical apparatus.
-This entire chapter is centered around various tehniques of using a computer for traffic simulation.
-There are a number of existing solutions to simulating traffic using software abilites, on various places on the macro-micro spectrum, as well as on other spectrums like discrete event vs continous simulation.
-Those will be explored and briefly categorized.
-
 \section{Abstract}
 
-The current chapter is aiming to propose a novel solution for simulating traffic in a city and exploring the opportunities which this solution makes possible to address.
-To describe it in few words, before going in full details on it during section \ref{sec:microservicesbasedapproachtrafficsim}, it is a discretes event based simulation, using a distributed microservices architecture where actors (defined as independent and intelligent participants in a traffic) are moving across a city and interacting with one another.
-They are a number of different instances running in isolation as containers.
-Their interaction is made via a single instance, called city simulator, which keep various statistics about the actors which are live and moving.
-On top of those there are few other supporting utilities, not essential for the current subject but mandatory for the simulation to run.
-They are a map and routing service which offers a route between two points of a city to those actors which need it and a service which renders a web page with the map of the city where the simulation takes place and the gradual occupation of its streets.
-
-The second area of this chapter is to analyze opportunities for optimizing the traffic within a city using the aforementioned simulation solution.
-Those ideas will revolve around the core concepts which form the foundation of the microservices based approach for traffic simulation: discrete and independent actors (also randomness is included into their behavior), decoupled in their logic and choice, interacting via discrete events based on their selfish interest (their behiavor which this simulation is modeled for is their interest in navigating their route as fast as they can, using the shortest path between two points, departure and arrival ones).
-
-Following concepts will be discussed in order to explore the opportunities for optimizing the traffic in a city.
-They are to be found in section \ref{sec:citytrafficoptimizations} and a brief list of them is presented here:
-
-\begin{itemize}
-    \item Modeling various actors with very different behaviors. For example a pedestrian and a city bus. Or a bicycle and a garbage truck.
-    \item Modeling various relations between different actors. An example of such a relation can be contained actors: a pedestrian can take a tram.
-    \item Introducing exception actors and their influence in the surrounding traffic. Firefighters truck moving throughout the city is one example. Cars stopped on the road is another.
-    \item Localize hot areas or bottlenecks of a city: junctions, streets, bridges.
-    \item Computation of various numbers and statistics. Few examples: average speed on street intervals, cars density per area and/or time.
-    \item Introduce geographical differences in simulation and study their influence: a roundabout instead of a junction, a new lane on a street.
-\end{itemize}
-
-\section{Current state of traffic simulation}
-
-\section{A microservices based approach for traffic simulation}
+\section{}
 \label{sec:microservicesbasedapproachtrafficsim}
 
-\subsection{Prerequisites}
-\label{subsec:prerequisites}
-In order to discuss how a discrete-based, microservices architectured simulation works, few prerequisites are needed to set the base.
-
-First definition we need is for setting the fact that we are using geographical coordinate system.
-It doesn't matter which one, what it matters is that all actors are using the same set of geographical coordinates and same projection (or to know if there should be a conversion).
-For example, they could be EPSG:4326 or EPSG:3857.
-To define this in an appropiate way, equation \ref{eq:coordinates} will be used.
-Its loosness allow any implementation to treat coordinates using a built in language data type (e.g.
-\textit{double} in C\#).
-
-\begin{equation}
-\label{eq:coordinates}
-    C = \{ c | c \in \Re, c \geq 0 \}
-\end{equation}
-
-Simulation wise, actors do not have any \textit{intent}.
-They are just \textit{dumb} programs which need to be programmed to have any small \textit{intention}.
-To give them a \textit{sense} of selfishness they are programmed to navigate between two random points.
-Therefore points need to be defined as a pair of coordinates.
-The implementation of this will have to take care of their order or of knowing which one represents latitude and which one longitude.
-Equation \ref{eq:coordinatespair} defines such a pair of coordinates.
-Furthermore we define a way which introduce randomness into an actor's behavior: the random choice of two points.
-The actor needs to have two points predefined, $Pd_1$ and $Pd_2$.
-Those are treated as an aproximative rectangular area, out of which other two points will be randomly chosen: $Pr_1$ and $Pr_2$.
-Equation \ref{eq:funcrandom} defines this function, without entering into the details of how such a random function would work.
-
-\begin{equation}
-\label{eq:coordinatespair}
-    P = \{ (p1, p2) | p1 \in C, p2 \in C \}
-\end{equation}
-
-\begin{equation}
-\label{eq:funcrandom}
-\begin{split}
-    & f_{rand}: P \rightarrow P, \\
-    & f_{rand}(Pd_1, Pd_2) = rand (Pd_1, Pd_2) \Rightarrow (Pr_1, Pr_2)
-\end{split}
-\end{equation}
-
-So far an actor has two random points and now it needs a route.
-For this the final function we need to define as a prerquisite is in equation \ref{eq:funcroute}.
-Essentially this function takes two points, in this case the two random points obtained earlier, in equation \ref{eq:funcrandom}: $Pr_1$ and $Pr_2$ and outputs a series or a list of points where first point is the first random point, $Pr_1$ and last point is the second random point, $Pr_2$.
-Between them a number of points represents the path which the actor should follow in order to travel from its first (randomly choosen) point to its second (randomly choosen) point.
-Translated into a more intuitive description, those points are the instruction which any maps application outputs when asked for a route.
-For example OpenStreetMap, \cite{openstreetmap} is a map provider and Graphhopper, \cite{graphhopper} is a route provider which uses OpenStreetMap as the default map.
-
-\begin{equation}
-\label{eq:funcroute}
-\begin{split}
-    & f_{route}: P \rightarrow [P], \\
-    & f_{route}(Pr_1, Pr_2) = route (Pr_1, Pr_2) \Rightarrow [Pr_1, Pc_1, Pc_2, ..., Pc_n, Pr_2]
-\end{split}
-\end{equation}
-
-\subsection{Microservices - city actor}
-\label{subsec:cityactor}
-
-Having defined the prerequisites from section \ref{subsec:prerequisites} we move on to present the first type of microservice from our architecture: the (city) actor.
-A city actor is meant to represent a specific behavior of an entity which moves across the city.
-To create such an actor it must be infused with specific behaviors based on the reality it wants to mimic.
-The prevalent example of this chapter will be a car.
-Therefore the prerequisites from section \ref{subsec:prerequisites} will suit a car's purpose to obtain a route using a predefined area where the car is allowed to wander.
-To further isolate its movements and decisions an algorithm is supposed to be used at its core.
-Listing \ref{lst:caractoralgorithm} presents it.
-
-\begin{lstlisting}[caption=Basic algorithm for a car actor, label=lst:caractoralgorithm]
-\\ Get two random points
-(pr1, pr2) = rand (pd1, pd2)
-\\ Get the route
-R = route (pr1, pr2)
-\\ Set the position to the first point
-P my_position = pr1
-int index = 0
-while (index < R.length) {
-    int speed_coeff = get_info(R[index])
-    my_position = advance(speed_coeff, R, index)
-    report(my_position)
-}
-\\ Report that the route has been finished
-report(no_position)
-\end{lstlisting}
-
-Few remarks about the core algorithm showed in listing \ref{lst:caractoralgorithm}.
-There are few pieces which are missing from the prerequisites but are discussed into section \ref{subsec:citysimulator} because they are related to the communications between a car actor and the city simulator.
-Those pieces are the calls to following methods: \textit{get\_info} and \textit{report}.
-The third one, \textit{advance} is the place where each actor is going to be personalized with own rules and behavior.
-Those aspects will be discussed into the subsection \ref{subsec:actordiffbehaviors}.
-
-The core algorithm needs to be run on continuosly basis, with a standard frequency (e.g.
-500 ms).
-Varying this number, also called simulation step can be an option for speeding or slowing the simulation running time.
-The \textit{advance} method is supposed to compute where to go next and it is responsible for incorporating the \textit{speed\_coeff} coefficient obtained from the city simulator.
-This coefficient is meant to express the density of cars on the street (if the actor which needs it is a car.
-In the case where it is a pedestrian the coefficient will be different.
-This way it is intuitively described the fact that a pedestrian walking on the sidewalk has nothing to do with the cars which are driving on near lanes and viceversa, a car driving on some lane doesn't care how many pedestrians are on the sidewalk near it) and it varies from expressing a totally free road to a road blocked by the numbers of cars waiting to pass through a junction.
-
-A written, more expansive description of this algorithm is as follows.
-For easier following of the simulation and for the current paper all actors are bounded to a certain area, expressed as two defined points $Pd_1$ and $Pd_2$.
-Those points are treated as the opposing corners of a rectangle and another two random points are chosen such that they are wihtin the rectangle: $Pr_1$ and $Pr_2$.
-The actor will then proceed to obtain a route, from another service, in order to know how to travel from $Pr_1$ to $Pr_2$.
-The route obtained is a series of another points, $Pc_1$ tp $Pc_n$, starting from $Pr_1$, called departure point, right to the finish point: $Pr_2$.
-They can be organized into subsets.
-A subset is a number of consecutive points which are treated as one instruction.
-Those instructions are meant to offer an easier guide, vocal or visual, for a human traveler.
-However, there are no human travelers into the simulation so the instructions have found another useful role.
-Usually an instruction is a limited length of a street during which the traveler doesn't need to do any change of direction.
-As soon as a junction must be traveled, there will be another instruction.
-Therefore this subset of the entire vector of points will become the interval which the actor reports as being traveling on to the city simulator.
-The city simulator, having such reports from all actors, will respond with the a number called speed coefficient, which reflects how many other actors are on the same interval and it will influence the traveling speed.
-The $advance$ method is meant to compute how the actor will advance and is the main place, though not the only one, where the behavior of different types of actors will be implemented.
-This method also takes care of handling the speed coefficient.
-Once the method computes the next position where the actor will be, that position will be reported back to the city simulator.
-This logic repeats itself as long as the actor has points and instructions which haven't been traveled yet.
-At last, the actor will report that it has been finished its route so that the city simulator will take into account one less actor.
-
-\subsection{Microservices - city simulator}
-\label{subsec:citysimulator}
-
-In this first chapter the city simulator is a separate entity which takes care of all the reports the actors send to it and also respond with every query about a certain route.
-Thus, it will need a way to keep this data.
-There are few assumptions around this which need to be taken into consideration:
-
-\begin{itemize}
-    \item The set of points $P$ represents routes whose interpretation are portions of streets;
-    \item A report sent by an actor contains a set of points and an indication if the actor is currently travelling on that route or if it went off it;
-    \item A set of points which was received from an actor can be a subset of another set of points received earlier from another actor;
-    \item A set of points which was received from an actor can be a superset of another set of points received earlier from another actor;
-    \item An actor can ask various information for any set of points it needs;
-    \item Any information which will be conveyed to an actor must take into account any subset and superset which might already exist;
-    \item An actor will send in each report an piece of information representing its type (car, person, etc.);
-\end{itemize}
-
-To illustrate above points the speed coefficient will be discussed.
-The first and uttermost information which an actor needs is representing the concentration of other actors of its type which around it or on a certain piece of road.
-For example, a car needs to know how many cars are around it, to know how fast it can advance.
-Or it might need to know how many cars are on a certain piece of road to asses if it can find a faster route.
-While the latter example is an intuitive one the former needs a little more explanation.
-Being a simulation, an actor doesn't know how many other actors are around it.
-In a real scenario the driver just have to look around but here there are no drivers and no looking around.
-One solution is for an actor to ask the city simulator how many other actors of its type are on the same street it's currently traveling on.
-Therefore, this information will need to be baked in the each actor type.
-For example a car will need to adjust its nominal speed with the coefficient which it received from the city simulator.
-
-\subsection{Implementation}
-
-The following implementation briefly presents the solution chosen by the authors in \cite{microservicestrafficsimulation}.
-
-First, a brief and high level view for few of the terms which will be used henceforth.
-A \textbf{microservice} is small and contained application, embedded into a tar file which contains everything it needs to run by itself.
-Moving tar files (called images) and running an image multiple times on the same machine gives a greater flexibility in developing and deploying application where the nature of distributin and scaling matters.
-A \textbf{container} is a running image.
-\textbf{Docker} \citep{docker} is a commercial (with free tier offers) for building, manipulating and running containers with limited functionality on the running side.
-
-Each actor type and the city simulators are an image.
-When they run, they become containers.
-Thus a simulation with hundreds of actors of car type needs only one image, that of a car actor.
-On top of this the second image comes into play: this is different and its called city simulator.
-It will run only as one instance or one container.
-The running environment, called Docker, will take care of running those two images and any other which are needed.
-Docker also offers a running environment where the containers can see each other through assigned names.
-When defining a running instance of type city simulator it can be specified that its name is \textit{citysimulator}.
-Docker will take care of providing a DNS-like experience for finding where this instance is running.
-Any other container which wants to talk to the \textit{citysimulator} will just use its name into the URL it wants to access (provided that the type of communications are REST APIs).
-Thus the easiness in use for container to container communications removes many frictions and smooths activities such as architecting a software application or developing it.
-
-Beyond the two images, those of a city car actor and a city simulator, two others are needed for the simulation to actually take place.
-One of them is a routing engine.
-It uses an underlying map, taken from Open Street Map \citep{openstreetmap}.
-The routing engine itself is called Graphhoper \citep{graphhopper} and offers routing services like finding the best (shortest) route between two points while at the same time providing useful options like computing the routing with the avoidance of a certain point (useful for actors which would like to avoid a too crowded street).
-The other image used is a web server which sources its data from the city simulator instance and presents a map where a human user can see the live simulation of the city.
-
-\begin{figure*}
-    \includegraphics[width=0.9\linewidth]{ScreenshotWebInterface2.jpg}
-    \centering
-    \caption{Screenshot of the web interface showing the running simulation}
-    \label{fig:ScreenshotWebInterface2}
-  \end{figure*}
-
-\section{City traffic optimizations}
+\section{}
 \label{sec:citytrafficoptimizations}
-
-\subsection{Actors with different behaviors}
-\label{subsec:actordiffbehaviors}
-
-One good opportunity for studying the traffic in a city is to model different actor types, each with different behaviors and travel rules.
-They all share same common goal: travel a route in as little time as possible.
-But the means of traveling (car, bus, pedestrian, etc.) and the rules (junctions, sidewalks, stop signs, bus lanes, bicycle lanes, etc.) are different for each of them.
-Therefore the interaction between them start to be complex in modeling via macro-mathematical means.
-Three examples are discussed in more detail:
-
-\paragraph{Cars} They are the default example of the current chapter.
-Their behavior is described in subsection \ref{subsec:citysimulator} and is presented here only as a comparison member for the other examples.
-\paragraph{Pedestrians} \label{par:pedestrians} They differ drastically from the cars in few aspects.
-First is the fact that they do not share same space of a street.
-From the perspective of a city simulator this means that it needs to know the actor type and it needs to maintain a separate zone of density information.
-The second aspect is that, altough, the pedestrian are by default walking on bare foots along streets, they might take public transport or taxis to travel between their points of interests.
-This has to be reflected into their routing algorithms.
-The other part of this aspect is discussed in subsection \ref{subsec:relationsbetweenactors}.
-\paragraph{Buses} Those are the same as cars but they have a fixed travel route to move across.
-Also they might share same lanes with cars or they might have different lanes.
-Thus the city simulator needs to know this actor type too and it needs to know where on the map they have special public transport lanes, to keep appropiate density numbers for them.
-\paragraph{Bicycles} The interesting particularity of those are the fact that they can limit the speed of other actors (cars or buses) to their speed on the places where overtaking is not possible.
-Again, this is one more burden for the city simulator because it needs to know where bicycles are and it needs to know which streets are too narrow.
-
-\subsection{Relations between actors}
-\label{subsec:relationsbetweenactors}
-
-\paragraph{Driving rules} Those are probably the most common interactions between actors.
-They interact with one another, regardless of their type, through the way a city works.
-Such a relation might be modeled by a junction or roundabout.
-Or it might be a crosswalking.
-Therefore the city simulator, once again, need to take into account those details.
-Modeling a junction might not be an easy feat and most probably here lies one of the places where a discrete-computed approach is more difficult to define and implement than an analogic aproximation.
-The difference, altough stated in loosely defined terms is that the definition of density can be a function of the number of actors on a given street.
-This can be computed at any time (or, better yet, it is computed on every event which occurs.
-i.e.
-an actor reporting its state on or off a street).
-A junction, to the contrary, might be too dynamic to be easy modeled the same way.
-An aproximation of it is a coefficient based upon an actor is allowed to go through that junction and continue its journey.
-Whenever an actor reaches a junction its algorithm will have to compute its advance speed based not only on the density (or speed coefficient) number but also based on the junction coefficient.
-And, of course, the city simulator might choose to vary that very junction coefficient based on a number of factors such as how many cars are awaiting on a joined street (intuitively, as more and more cars are piling up around a junction, harder it is to pass through it).
-
-\paragraph{Contained actors}
-The paragraph \ref{subsec:actordiffbehaviors} discussed a type of actor called pedestrians.
-They might independent but also they might choose to take a bus from one point to another.
-We call this a contained actor.
-Thus its reporting will be enriched with one more detail: if is walking by itself or if its inside a bus or a subway.
-The city simulator will have to compute the densities accordingly for the street or sidewalk but also it might need to know how many pedestrians is a bus allowed to carry.
-
-\paragraph{Close-follow actors}
-The bicycle which pedals onto a bus lane might be holding up a bus behind it.
-Or a number of cars if it's on a common lane.
-Therefore the speed they have might not be the one which they have without a bike.
-
-\subsection{Exception situations}
-Various situations, altough exceptional, might have a big influence on the surround traffic.
-If a portion of a street is going under few heavy modifications then the traffic might not be allowed to flow through there.
-Or if a car is stopped on the lane with hazard lights on.
-Therefore the city simulator will have the opportunity to modify the speed coefficient.
-It might choose to make it difficult for traffic to pass through that area (if out of 4 lanes, two for each direction, only 2 are used, one for each direction) or it might choose to completely block the portion.
-It would be of great interest to see, in this case, how the traffic adapts itself to flow through other streets and how much pressure it puts on them, on a different situation than the usual one.
-The situation of a completely blocked street might also pose the idea of making it a pedestrian-only street and study the effect it would have on the car traffic.
-
-\subsection{Analyzers}
-
-The final part of this traffic simulation chapter is about the possibilities of analyzing pieces of the big picture.
-It is the supreme purpose of a traffic simulation, to be able to analyze how it works, what specific behaviors it exhibits, in order to find new means and ways of improving it.
-
-The central piece of those analyzers is the city simulator.
-It contains all the centralized data, from all its actors and it has a minimal API (Application Programming Interface) to serve this data.
-Actors are the first to consume it as they might need various pieces of information to wisely choose their track.
-A separate microservice may be the one which takes this data and analyzes it further.
-The reason for a separate microservice is the separation of concerns.
-The city simulator is responsible only for keeping together the data and serving it.
-Any other duty of transform, extrapolate, etc.
-is not its task to do.
-
-Therefore few ways of taking advantage of this data are described below. They are by no means exhaustive and represent only few places to start.
-
-\paragraph{Hot areas}
-
-Hot areas are defined as places or streets where traffic is pilling up more than in other areas. Intuitively
-
-\paragraph{Statistics numbers}
-
-\paragraph{Geographical differences}
 

### 2020-10-15 thesis/report2.tex
Advance on the report 2

@@
 \usepackage{natbib}
+\usepackage{url}
+\usepackage{amsmath,amssymb,amsfonts}
+
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+
+\graphicspath{ {.} }
+
 \usepackage{listings}
-\usepackage{graphicx}
+\usepackage{xcolor}
+
+\usepackage{balance}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+    }
+
+    \lstset{style=mystyle}
 
@@
 \pagestyle{myheadings}
-
 \title{ City traffic, IoT and Blockchain technologies }
@@
 
+    Following previous chapter, about how traffic in a city can be modeled and analyzed, we step forward into the realm of IoT and Blockchain technologies.
+Those two terms, IoT and Blockchain, are only buzz words.
+Or they are so general, so loosely defined that we can treat them as marketing or advertising terms.
+Hence, we have to define them for our specific purpose.
+
+    Internet of Things is term coined to express the multitude of devices or physical objects which are connected, in various ways, to the public internet.
+They are transmitting data, are requesting data from a central server, one from another or in collaboration with a mesh of IoT devices.
+An IoT device has something to "offer".
+It is usually a specific job it has to do by being provided with sensors and a alogrithmical method of sending data.
+On the other hand it can be also used for automation jobs where not only it is able to send data to other systems but it can act based upon the data it acquires or it can receive various commands to execute from other IoT devices or systems.
+In the context of the present thesis an IoT device can be considered an actor which is travelling thorugh the city and it reports data or it uses other actor's data.
+
+    Blockchain are a set of emerging technologies which were first aimed at decentralizing financial transactions.
+Bitcoin cryptocurrency\citep{SatoshiNakamoto}, first of its kind and still the most popular, created a global distributed and decentralized peer to peer system where financial transactions (sending money from one account or wallet to another one) are taking place without relying to a central party or a bank for shielding against double spend \citep{DoubleSpend}.
+It also sparked a cornucopia of other cryptocurrencies, each with its own distinctive feature.
+Few of them, worth to be noted because they will be analyzed further in this thesis, are Ethereum, Hyperledger Fabric and IOTA.
+
 \section{Abstract}
 
-\section{}
-\label{sec:microservicesbasedapproachtrafficsim}
+    This chapter is aiming to lay the ground on how we intend to bring the city traffic infrastructure from previous chapter and its needs into a blockchain technology.
+The two worlds, two very different paradigms, can be bridged together: IoT and Blockchain.
+The traffic of a city is an inherent IoT application.
+The actors, regardless of their type (cars, bikes, pedestrians), might have limited power consumption, might have intermittent internet conectivity and can be offline for long periods of time.
+Those constraints led to design an IoT solution for their communications.
+On the other side a Blockchain technology has needs which, at first sight, are in contrast with what an IoT device can do: proof of work is the first reason as it requires a great amount of power.
+Then the continuous online presence is needed, to participate in the blockchain network and stay up to date.
+Last, but not the least, the smart contracts nature of each Blockchain requires a different paradigm, a shift in how business logic are written.
+This comes with its own set of opportunities and limitations.
+To bring those areas together gaps must be identified in detail and bridges must be built over them.
+The target is to discuss those areas for two blockchains networks: Ethereum 2.0 and IOTA.
 
-\section{}
-\label{sec:citytrafficoptimizations}
+\section{Introduction}
+\label{sec:sec}
+
+    The journey of this chapter follows the items below.
+We first establish what IoT means for the city traffic simulation from chapter one.
+Several concepts from it will be repeated but from a different perspective.
+This happens in section \ref{sec:IoTNature}.
+Afterwards few aspects from Ethereum 2.0 blockchain will be discussed, in section \ref{sec:EthereumNature}.
+Same perspective but with a different blockcahin will be in section \ref{sec:IOTABlockchainNature}.
+Both of them have various proposals around smart contracts executioon and this will be discussed in section \ref{sec:SmartContracts}.
+A big part of every distributed ledger is the way it has designed its Proof of Work or Proof of Stake.
+Those will be discussed in section \ref{sec:PoWandPoS}.
+Finally, the two worlds will be connected using few proposals in section \ref{sec:GapsBridges}.
+
+    \begin{itemize}
+        \item The IoT nature of the actors from a city traffic
+        \item The Blockchain nature of Ethereum 2.0
+        \item The Blockchain nature of IOTA
+        \item Smart contracts
+        \item Proof of Work and Proof of Stake
+        \item Gaps and bridges
+        \item Conclusion
+    \end{itemize}
+
+\section{The IoT nature of the actors from a city traffic}
+\label{sec:IoTNature}
+
+    IoT is a very general concept with many dimensions.
+Across the type of applications dimension there are several domains like, but not limited to, consumer applications (e.g.
+smart home), medical and healthcare, industrial (e.g.
+agriculture), military, etc.
+On the dimension of physical connection, they can use bluetooth, near-field communication or wifi.
+Therefore the effort of defining the general nature of IoT is quite extensive and doesn't fit into current subject.
+
+    What we can achieve here, however, is to define the nature of the actors participating in a traffic city (or in a simulation, see previous chapter).
+The walkthrough of this has to reach few points:
+
+    \begin{itemize}
+        \item Hardware
+        \item Communication layer
+        \item Communication style
+        \item Security
+        \item Anonimity
+        \item Interoperability
+    \end{itemize}
+
+    \paragraph{Hardware} Traditionally a person which travels within a city is expected to have a standalone GPS or a smartphone with GPS abilities and a suitable application (e.g.
+Waze).
+For our purposes the hardware is a less important choice.
+What matters is that the software running is able to send data, receive data and act accordingly for informing the user.
+Given the decentralized architecture this thesis is going towards, a classic navigation application (like Waze, which depends on a server to be fully functionally) is less likely to fit.
+Therefore a custom software stack needs to run.
+With respect to how much powerfull the navigation device is, there is a plethora of options, ranging from low energy to high performing CPUs and GPUs.
+This enables the navigation device to be able to run smart contracts and proof of work activities.
+\ref{sec:PoWandPoS} will talk more on this subject.
+
+    \paragraph{Communication layer} The first assumption is that the device is able to send and receive data via a TCP based protocol.
+The second assumption is that the bandwith is enough to send and receive packets which have JSON format.
+
+    \paragraph{Communication style} To explain the style of communications for an actor the best way is to illustrate it using two JSON examples.
+Following \cite{MicroservicesCityTrafficSimulation} paper about a microservices approach for city traffic simulation, we take and use the data defined there for defining the communication style.
+Listings \ref{lst:jsonClientEnums} shows the JSON schema for a report which an actor sends (in the cited paper it sends this message to a central server but the intention of the thesis is to have a decentralized architecture and no central servers).
+\textit{ReportOnTheLine} is the report sent by one agent to notify the city that he is currently advancing through one line.
+\textit{ReportOffFromLine} is the report sent by one agent to notify the city that he has finished advancing through one line and has departed from it.
+\textit{SendReport} is a message passed from an actor to the city indicating its status (e.g.
+location).
+\textit{AskForLine} is a message passed from an actor to the city.
+A response is awaited.
+\textit{RespondWithLine} is a message passed from the city to an actor and it contains line data.
+
+    \begin{lstlisting}[caption=JSON schema with enumerations for messaging, label=lst:jsonClientEnums]
+    "ReportType": {
+        "type": "array",
+        "items": {
+            "type": "string",
+            "enum": ["ReportOnTheLine", 
+                        "ReportOffFromLine"]
+        }
+    },
+
+    "MessageType": {
+        "type": "array",
+        "items": {
+            "type": "string",
+            "enum": ["SendReport",
+                        "AskForLine", 
+                        "RespondWithLine"]
+        }
+    }            
+    \end{lstlisting}
+
+    \begin{lstlisting}[caption=JSON top level struct (envelope), label=lst:jsonEnvelope]        
+    "Envelope": {
+        "MessageType": "SendReport
+        "Payload": {
+            "ReportType": "ReportOffFromLine",
+            "Line": [ { 45.64354, 34.6768 }, 
+                        { 45.66543, 34.6722 }
+            ]
+        }
+    }
+    \end{lstlisting}
+
+    The global format of such a JSON needs the ability to transport any type of message. Therefore the listing \ref{lst:jsonEnvelope} represents such a message.
+
+\section{The Blockchain nature of Ethereum 2.0}
+\label{sec:EthereumNature}
+
+\section{The Blockchain nature of IOTA}
+\label{sec:IOTABlockchainNature}
+
+\section{Smart contracts}
+\label{sec:SmartContracts}
+
+\section{Proof of Work and Proof of Stake}
+\label{sec:PoWandPoS}
+
+\section{Gaps and Bridges}
+\label{sec:GapsBridges}
+
+\section{Conclusion}
+\label{sec:Conclusion}
 

### 2020-10-20 thesis/report2.tex
One more section completed on report 2.

@@
 \usepackage{xcolor}
+\usepackage{dirtytalk}
 
@@
 
-    The global format of such a JSON needs the ability to transport any type of message. Therefore the listing \ref{lst:jsonEnvelope} represents such a message.
+    The global format of such a JSON has the ability to transport any type of message.
+Therefore the listing \ref{lst:jsonEnvelope} represents such a message.
+The envelope, on its turn, contains another message, of different formats.
+Listing \ref{lst:jsonClientEnums} shows one example of such a contained message.
+It is the way a client reports one line it is travelling onto or one line it has just finished travelling.
+Both listings, \ref{lst:jsonEnvelope} and \ref{lst:jsonClientEnums}, show json schema messages.
+
+    Another aspect of those messages is that an actor can send a message of type \textit{AskForLine} and afterwards it is expecting a message back with details about that line (e.g.
+how many other actors are travelling on it).
+The question of uttermost importance is: to whom does it send this message?
+In \cite{MicroservicesCityTrafficSimulation} the authors use a \textit{central server} and each actor is communicating only with it.
+However, having already set a foot into the world of decentralized systems, we do not have anymore a \textit{central server}.
+A decentralized ledger or blockchain or system, soon to be discussed in \ref{sec:EthereumNature}, has no such concept of \textit{server}.
+For the moment of this paragraph which is concerned only with the way an actor is communicating, we can assume that there are two components: the \textit{frontend} one and the \textit{backend} one.
+The \textit{frontend} one is concerned with showing information to the user (graphical, text, etc.) and also computing (via classic algorithms or Artifical Intelligence algorithms) the things which a user needs (e.g.
+shortest or fastest path from one point to another).
+To obtain various benefits along those computations there are certain types of information which have to be obtained.
+There is static information like the map of the city where the actor is in and there is dynamic information about what is currently, at the time of the computation or at the certain and precisely known time in the future, happening in the city.
+One reiterated example is how many actors are already on a certain line in the city.
+Having this information for the lines which would represent one of the paths the querying actor can run onto, the computations can compute more accurately a shorter (time wise) path or it can optimize various other outcomes.
+So far this has been the \textit{frontend}.
+The other part is the \textit{backend}.
+This is concerned less with how the data is presented to the user and how the computations are working.
+Its concerns are how to obtain the data needed by the \textit{frontend} and how to serve it.
+Thus it can be a server where all the actors are calling into.
+But it also can be one of the peers which are part of a decentralized network like Ethereum or IOTA.
+It can take care of all the nitty-gritty parts.
+It can compile various types of data, it can run necessary validations which are mandatory in such types of networks and so on.
+More details about how it does this are not part of this paragraph or of this chapter but they will be discussed later on on this thesis.
+Such distinctions between \textit{frontend} and \textit{backend} can also been seen in \cite{ApproachesToFrontEnd} where, among few options, there are standalone nodes responsible for participating into the network and a \say{local device} which offers \say{local} functionalities for various applications.
 

### 2020-10-22 thesis/report2.tex
Small advancements on report 2.

@@
 
+    \paragraph{Security} The security is not a concern at this stage.
+The commmunications between the \textit{frontend} and the \textit{backend} have been described above only with respect to the form of communication.
+Any other aspects like encryption, certificates will be part of next sections.
+
+    \paragraph{Anonimity} In the client server architecture 
+    \citep{MicroservicesCityTrafficSimulation} the server does not need to maintain an identity of the clients which it has to communicate with.
+In order to keep an accurate list of how various lines and their load any client is required to send a report where it signals that it finished the traveling through a line.
+This can be seen in listing \ref{lst:jsonClientEnums}, first message type called \textit{ReportOffFromLine}.
+The momentarily assumption is that the clients are correctly behaving and are not actively trying to introduce unwanted effects into the wider network, nor to mislead the server about their position.
+Such enforcments, of mandatory good behavior, are modeled through mechanisms like Proof of Work and Proof of Stake and will be discussed in section \ref{sec:PoWandPoS}.
+The other aspect which can naturally occur is when a client, altough it has good intentions to send a \textit{ReportOffFromLine} message, it is not able to due to dropped connections or any other reason outside its control.
+A server might need, in this case, to implement a mechanism of timeouts for \textit{ReportOnTheLine}.
+In order to know which message to timeout it should keep an identification token for each client.
+This is one of the solutions.
+Knowing the average time to travel the respective line it can safely asssume that if there is no sign from that client within a reasonable and configurable time, it can dequeue its load.
+
+    \paragraph{Interoperability} The plethora of programming languages, frameworks, runtimes and operating systems give a multitude of choices of where and how to run such clients and servers.
+So far nothing has been discussed regarding this aspect.
+The only place where we can see an implementation of a client server architecture, as presented in this section, is \cite{MicroservicesCityTrafficSimulation}.
+Programming language is Golang, the UI is made using web technologies like HTMl and JavaScript, together with various frameworks or libraries for ease of development (a library called Gorilla for working with WebSockets in Golang, OpenSourceMap and OpenLayers for displaying and manipulating maps in the browser, etc.).
+However, despite all those fine grained choices, the one which assures the interoperability is the ubiquity of JSON formats.
+Every language supports it and every language knows how to encode its internal data structures into a JSON format.
+Or how to decode them.
+Thus listings \ref{lst:jsonClientEnums} and \ref{lst:jsonEnvelope} are presented in as JSON messages schema.
+They also represents a small part of wannabe protocol.
+
 \section{The Blockchain nature of Ethereum 2.0}
 \label{sec:EthereumNature}
+
+    \paragraph{Security} Security in Ethereum is a complex process.
+If an account, being a user's account or a contract account, has successfully secured the private signing key, it is assured that no one can simulate it and no one can execute transactions on its behalf.
+Known breaches have always occured in the same way: private keys have been leaked and malicious actors were able to execute transactions with them.
+    An approach in securing IoT networks is described in \cite{SecureCommunicatingThingsNetworkFramework}.
+The aim is to ensure secure product shipment, as well as connecting manufacturing locations from different geographical locations.
+They have a front end for human interactions and a back end for the internal communication of the blockchain network.
+In the blockchain networks there are two types of associates, the authentication nodes also known is miners and execution nodes.
+The authentication nodes take care of validating new transactions and keeping up to date with the ever growing blockchain in which they participate.
+The executing nodes are watching accumulating values and verify their validity.
+The authors are considering various scenarios like traceability of product records, illegal activities made by workers or shippments.
+Transactions like those are always embedding the previous block's hash, providing a secure way around manipulating
 
@@
 
+% good behavior
+
 \section{Gaps and Bridges}

### 2020-10-25 thesis/report2.tex
Some report 2 paragraphs about security in blockchains.

@@
 Known breaches have always occured in the same way: private keys have been leaked and malicious actors were able to execute transactions with them.
+
     An approach in securing IoT networks is described in \cite{SecureCommunicatingThingsNetworkFramework}.
@@
 They have a front end for human interactions and a back end for the internal communication of the blockchain network.
-In the blockchain networks there are two types of associates, the authentication nodes also known is miners and execution nodes.
+In the blockchain networks there are two types of associates, the authentication nodes also known as miners and execution nodes.
 The authentication nodes take care of validating new transactions and keeping up to date with the ever growing blockchain in which they participate.
@@
 The authors are considering various scenarios like traceability of product records, illegal activities made by workers or shippments.
-Transactions like those are always embedding the previous block's hash, providing a secure way around manipulating
+Transactions like those are always embedding the previous block's hash, providing a secure way around manipulating or tampering records of shipped items, for example.
+
+    Another proposal of security can be found in \cite{SecuringUserIdentity}.
+The authors address a more basic need of having IoT devices participating in a M2M (machine-to-machine) network, for extended periods of time.
+Extended periods of time implies that certificates must be rotated and the old ones shouldn't be exploitable by malicious actors if they are able to compromise those private keys (certificates, in this case, are made of public and private keys pairs where the private key is used by an entity to prove its identity and various details which are part of its identity).
+Another part of their proposal is to replace raw and uncertified public keys from Bitcoin or Ethereum transcations with Enrollment certificates and transaction certificates.
+The latter are included in usual transactions while the former are for proving the attributes like identities, identifiers, etc.
+Transcations certificates are used to hide sensitive information in the public blocks while at the same time to prove the ownership of the entity which made them.
+To sum up, in author's word, the proposal is a permissioned blockchain technology with \say{privacy-preserving,  traffic-analysis-  resistant methodology  that  leverages  external  trust  relationships  so  as to establish an auditable identity- and attributes- management authorization  framework}.
 

### 2020-10-25 thesis/report2.tex
Another small paragraph.

@@
 
+    Data assurance and resilience are two themes within the big concept of Blockchain security.
+In \cite{TowardsdatAassuranceAndResilience} the authors are describing a mechanism for drones that achieves few objectives like trusted data origin (allows to know for sure which drone send a part of data), instant and permanent data integrity ()
+
 \section{The Blockchain nature of IOTA}

### 2020-10-25 thesis/report2.tex
Another paragraph.

--- 2020-10-25 thesis/report2.tex
@@
 
+    Data assurance and resilience are two themes within the wide concept of Blockchain security.
+In \cite{TowardsdatAassuranceAndResilience} the authors are describing a communication mechanism for drones that achieves few objectives like trusted data origin (allows to know for sure which drone send a part of data), instant and permanent data integrity (data which the drones are reporting), trusted accountability (each command is recorded and malicious activities can be tracked down to the entities which started them) and resilient backend (nodes can go down or can start to behave weirdly with minimum distruption to others).
+The paper is especially interesting because it proposes a system which resembles ours.
+A swarm of devices (drones in the paper's case, cars in ours; cluster of drones and cluster or cities of cars) is a similar concept.
+A database separated from the blockchain, each part with its specific roles, is another similar concept.
+Nevertheless there are few differences.
+First and foremost the the Control System give drones a path where they should go on while the Cloud Server can augment the path with intelligent algorithms.
+On this thesis the application models the control as being exclusive the role of the actor.
+No central server, being a cloud or an intermediary block chain node, can give instructions to an actor.
+This doesn't mean the actor has the freedom and flexibility in doing everything it wants.
+But more on this in the next chapters.
+
 \section{The Blockchain nature of IOTA}

### 2020-11-02 thesis/report2.tex
Continued on the report 2.

--- 2020-10-25 thesis/report2.tex
@@
 
-    \paragraph{Security} Security in Ethereum is a complex process.
+    Security in Ethereum is a complex process with many faces.
+It drives the entire reason of why ht nature of Ethereum is so different than a centralized solutin.
 If an account, being a user's account or a contract account, has successfully secured the private signing key, it is assured that no one can simulate it and no one can execute transactions on its behalf.
 Known breaches have always occured in the same way: private keys have been leaked and malicious actors were able to execute transactions with them.
+But for this mechanism to work the rest of the blockchain infrastructure is complex enough to not fit into this paragraph.
+It is worth, though, to be noted few aspects.
+The extensive use of private/public key pair is the first one.
+Second, the Keccak256 hashing algorithm are used to generate proof of works.
+Those proof of works will be discussed widely in the section \ref{sec:PoWandPoS} but for this section it is important to mention them as they have an immense value in providing the security of mining a transaction.
+All miners are racing up to find a solution using the Keccak256 hashing algorithm and all of them are obliged to so because of consensus rules.
+Any of them which would try to trick this step will lose money.
+Thus the security of mining transactions and making sure the transaction is made by the true owner and the money are received by the true intended recipient are baked into the rules used for finding the proof of work and reaching consensus among the enire blockchain.
 
@@
 
+    Data assurance and resilience are two themes within the big concept of Blockchain security.
+In \cite{TowardsdatAassuranceAndResilience} the authors are describing a mechanism for drones that achieves few objectives like trusted data origin (allows to know for sure which drone send a part of data), instant and permanent data integrity (data cannot be tampered with on during its entire travel from drones to the blockchain or to the database).
+The mechanism involved here is generaaly known as Public Key Infrastructure (PKI).
+PKI encompasses all the operations, storage, key generation, keys transmission, key rotation, etc.
+needed for ensuring a crucial aspect for the entities which are communicating in a system: a message can be verified that it was sent by the entity which claims it sent it while the message cannot be tampered with by a man-in-the-middle attack.
+
 \section{The Blockchain nature of IOTA}
 \label{sec:IOTABlockchainNature}
+
+    IOTA is a distributed ledger technology, but not a blockchain, that allows devices in a IOTA network to transact in micropayments and to send each other immutable data.
+One crucial difference between IOTA and a blockchain is the how the transactions are linked together.
+In blockchain-type technologies, as the name implies, a transaction references only the one behind it and is referenced only by the one which comes next.
+In IOTA, being a distributed ledger technology (DLT), a transaction can reference multiple transactions which were made earlier and can be referenced by multiple transactions which comes after it.
+This is known as direct acyclic graph (DAG).
+Another crucial difference between a blockchain and IOTA DLT is the consensus mechanism.
+In the blockchain the consensus is made through the proof of work by the miners.
+In IOTA there are no miners and there is no proof of work.
+Instead there is an entity called The Coordinator or more recently Coordicide.
+Their aim is to validate transactions without requiring a proof of work mechanism.
+The main reason for this is to give the ability of participating to many other types of devices which otherwise do not have the power of doing proof of work tasks.
 

### 2020-11-02 thesis/report2.tex

@@
     Data assurance and resilience are two themes within the big concept of Blockchain security.
-In \cite{TowardsdatAassuranceAndResilience} the authors are describing a mechanism for drones that achieves few objectives like trusted data origin (allows to know for sure which drone send a part of data), instant and permanent data integrity (data cannot be tampered with on during its entire travel from drones to the blockchain or to the database).
-The mechanism involved here is generaaly known as Public Key Infrastructure (PKI).
+In \cite{TowardsdatAassuranceAndResilience} the authors are describing a mechanism for drones that achieves few objectives like trusted data origin (allows to know for sure which drone send a part of data), instant and permanent data integrity (data cannot be tampered with on during its entire travel from drones to the blockchain or to the database), trusted accountability (each command is recorded and malicious activities can be tracked down to the entities which started them) and resilient backend (nodes can go down or can start to behave weirdly with minimum distruption to others).
+The paper is especially interesting because it proposes a system which resembles ours.
+A swarm of devices (drones in the paper's case, cars in ours; cluster of drones and cluster or cities of cars) is a similar concept.
+A database separated from the blockchain, each part with its specific roles, is another similar concept.
+Nevertheless there are few differences.
+First and foremost the the Control System give drones a path where they should go on while the Cloud Server can augment the path with intelligent algorithms.
+On this thesis the application models the control as being exclusive the role of the actor.
+No central server, being a cloud or an intermediary block chain node, can give instructions to an actor.
+This doesn't mean the actor has the freedom and flexibility in doing everything it wants.
+But more on this in the next chapters.
+The mechanism involved here is generaly known as Public Key Infrastructure (PKI).
 PKI encompasses all the operations, storage, key generation, keys transmission, key rotation, etc.

### 2020-11-03 thesis/report2.tex
Another paragraph on report 2.

@@
 In blockchain-type technologies, as the name implies, a transaction references only the one behind it and is referenced only by the one which comes next.
+This gives a chain of blocks, each block being related only to two other blocks.
 In IOTA, being a distributed ledger technology (DLT), a transaction can reference multiple transactions which were made earlier and can be referenced by multiple transactions which comes after it.
@@
 In IOTA there are no miners and there is no proof of work.
-Instead there is an entity called The Coordinator or more recently Coordicide.
+Instead there is an entity called The Coordinator or more recently a post coordinator named Coordicide.
 Their aim is to validate transactions without requiring a proof of work mechanism.
 The main reason for this is to give the ability of participating to many other types of devices which otherwise do not have the power of doing proof of work tasks.
+
+    Coordicide has few specific goals.
+Known as the scalabilty trilemma, both in Ethereum and IOTA worlds, the guarantee of current IOTA solutions offers at most two (scalability and security) out of three characteristics from this list: scalability, security, decentralization.
+Coordicide aim to offers all three of them.
+Scalability means that the rate of transcations are not limited by the Coordicide protocols.
+The can scale as much as needed.
+Security means that the entire network cannot be attached by malicious actors.
+They cannot influence the consensus.
+Decentralization means that all nodes can be part of the consensus process and that there is no central authority to oversee it.
+
+    \cite{EnhancingIoTSecurityIOTA} is discussing a practical approach (a smart utility meter system and a smart car transaction system) towards IoT using blockchains (Ethereum and IOTA).
+Their first finding is the unsuitability of Ethereum blockchain: mining is computationally intensive, poorly scaling for the raising number of IoT devices, overhead traffic because of the Ethereum protocols.
+Their second finding, however, is the IOTA DLT readiness for IoT devices.
+IOTA reference implementation specifies two types of nodes: a full node and a light node.
+The full node runs locally (on a ordinary computer) and can be accessed via a GUI.
+The light node is a local interface from the device itself to the full node it belongs to.
+The solution they have designed has scalability out-of-the-box, due to the nature of IOTA, partition tolerance, security through small fees which would prohibit an attacker to spam the network and the lack of peer discovery.
+
+    % Privacy aware IOTA ledger: Decentralized mixing and unlinkable IOTA transactions
 

### 2020-11-04 thesis/report2.tex
More and more work on report 2.

@@
 
-    % Privacy aware IOTA ledger: Decentralized mixing and unlinkable IOTA transactions
+    \cite{PrivacyAwareIOTALedger} tackles another area: privacy or anonimity.
+Various techinques, the authors are claiming, offers the ability for a Bictoin account, or an IOTA account to be identified and tracked such that two informations can be obtain: total balance and transaction history.
+They summarize various related works of mixing services where tokens are exchanged, based on fees both for Bitcoin and for IOTA.
+Then they proceed to define the problem with few design goals: anonimity, correctness, compatibility, no mixing fees and efficiency.
+Their proposal have three phases.
+Phase 1 is called settlement and its aim is to take a pledge from the participants.
+Phase 2 is called outputs shuffling and it randomizes input and output addresses.
+Thus no participant can learn what relation was originally between any input address and any output address.
+Phase 3 is called transaction and it assures that the funds are correctly transferred from the inputs to the outputs.
+This mixiing protocol assures the properties discussed earlier.
 
@@
 \label{sec:SmartContracts}
+
+    Bitcoin is capable only of a limited, non Turing-complete ability of executing smart contracts.
+Therefore its main purpose remained transfer of money.
+Many other blockchain technologies, most notable being Ethereum, has been designed from the ground up to have support for smart contracts.
+Ethereum has a virtual machine, called Ethereum Virtual Machine (EVM).
+The EVM takes an input from the current Ethereum state and produces an output or a transaction.
+It is a stack machine with a depth of 1024 items, where each item is a 256-bit word.
+Every client implementation has an EVM implementation because each client has to be able to execute smart contracts.
+EVM have those smart contracts being \say{on-chain}.
+This means they are part of the core protocol.
+Execution and validation are being made by all nodes in the protocol.
+This offers a good security architecture and as much as the network grows it will improve the security even more.
+Smart contracts, which are one of the two types of Ethereum accounts (the other being the usual user accounts), are able to transfer tokens from their account without providing the usual signature of an user account.
+But there also disadvantages.
+All nodes have to execute a smart contracts and this leads to a lack of scalability.
+Their execution cost money (gas in Ethereum paralance).
+Thus the second type of smart contracts appear in sight: \say{off-chain}.
+The main difference is that not all nodes are mandated to execute them and thus the consensus is not part of the core protocol.
+The advantages are first the scalability because not all the nodes need to take part into the consensus.
+The greatest disadvantage is the lack of privacy (hence the work of \cite{PrivacyAwareIOTALedger} to improve it).
+
+    What are the roles of smart contracts, though?
+Ever since the blockchain gained wider adoption the insufficiency of Bitcoin, first one of its kind and still the most successfull, has been an obstacle for implementing other types of transactions than the simple money transfer.
+Therefore other blockchain have started to be designed from the ground up to allow more complex logic.
+Examples are plenty: decentralized voting, assets transfer and ownership, etc.
 

### 2020-11-05 thesis/report2.tex
One more paragraph.

@@
 Those constraints led to design an IoT solution for their communications.
-On the other side a Blockchain technology has needs which, at first sight, are in contrast with what an IoT device can do: proof of work is the first reason as it requires a great amount of power.
+On the other side, a Blockchain technology has needs which, at first sight, are in contrast with what an IoT device can do: proof of work is the first reason as it requires a great amount of power.
 Then the continuous online presence is needed, to participate in the blockchain network and stay up to date.
@@
 The target is to discuss those areas for two blockchains networks: Ethereum 2.0 and IOTA.
+Section \ref{sec:GapsBridges} will go into more detail.s
 
-\section{Introduction}
-\label{sec:sec}
+\section{Contents}
+\label{sec:Contents}
 
@@
 Examples are plenty: decentralized voting, assets transfer and ownership, etc.
+Ethereum is the most popular offering with a smart contracts (see section \ref{sec:EthereumNature}).
+IOTA is another one and is tailored for the IoT world.
 
-\section{Proof of Work and Proof of Stake}
-\label{sec:PoWandPoS}
+    % some papers from 'papers 2' folder can be added here
 
-% good behavior
+% \section{Proof of Work and Proof of Stake}
+% \label{sec:PoWandPoS}
+
+    % some other papers from 'papers 2' folder
 
@@
 \label{sec:GapsBridges}
+
+    \paragraph{Serverless} We found out, already, that there is no central server into the sight of a blockchain or a distributed ledger technology.
+Therefore, our solution designed in chapter one, will need to undergo drastic modifications.
+There was already o similar idea which could help us on this matter, proposed by \cite{ApproachesToFrontEnd}.
+The frontend side will remain the actor's interaction with the user.
+An example can be an Android application running on the phone or an Android application running on a car's dashboard.
+The backend might have different configurations:
+
+    \begin{itemize}
+        \item Backend is running on the same machine with the frontend.
+For a mobile phone it might not be a good idea if the resources needed for it will drain batteries and consume data plans.
+For an application running on a car's dashboard, this can be an option.
+        \item Backend is running on other machine than the one where the frontend is running.
+It can be a user's machine, running somwhere at home or in the cloud or it can be a paid one, a Software as a Service offering, run by a third party on behalf of the user.
+The core principle is that heavy computations, if necessary for Ethereum, for example, are handed over to a more powerful system.
+        \item Backend is not running at all anywhere.
+Instead, all the interactions an actor need are sent to miners of a Blockchain.
+Miners will be responsible for gathering data, formulating responses and sending answers back to the users.
+All the interactions may be recorded on the ledger.
+The frontend will get its data from the ledger as soon as a miner adds it and is validated by the network.
+    \end{itemize}
+
+    \paragraph{Data on ledger} Having so many interactions between an actor's need and the ledger, there is the legitimate question of which data should be recorded forever on the ledger and which should not.
+
+    \paragraph{Miners' decisions} The miner will decide the path the actor will travel
+
+    \paragraph{Local ledgers} Coordinators, local 'servers' might exist for each city or region
+
+    \paragraph{Actor role} What is an actor's role on the blockchain
+
+    \paragraph{Third party observers} Participants on the network, only for observing
 
